### 2010-06-07 ms/main.tex
Add a template for MS

\documentclass{report}
\usepackage{graphicx}
%\usepackage{fullpage}

\begin{document}

\input{titelblatt}

\tableofcontents
\chapter{Introduction}
\label{sec:intro}
To test bibliography~\cite{Monaghan2005}

\chapter{Method}
\label{sec:method}

\chapter{Results}
\label{sec:results}

\chapter{Conclusion}
\label{sec:conclusion}

\bibliography{bibdata}
\bibliographystyle{plain}

\end{document}