### 2010-06-07 ms/main.tex
Add a program listiong example

\documentclass{report}
\usepackage{graphicx}
\usepackage{listings}
%\usepackage{fullpage}

\begin{document}

\input{titelblatt}

\tableofcontents
\chapter{Introduction}
\label{sec:intro}
To test bibliography~\cite{Monaghan2005}

\chapter{Method}
\label{sec:method}

\lstinputlisting[language=c++,caption=An example of program listing,captionpos=b]{code/hello.cpp}

\chapter{Results}
\label{sec:results}

In the Figure~\ref{fig:1dshock} one can see \ldots

\begin{figure}[h]
  \centering
  \includegraphics[width=0.95\textwidth]{img/1dshock.eps}
  \caption{An example of a figure}
  \label{fig:1dshock}
\end{figure}

\chapter{Conclusion}
\label{sec:conclusion}

\listoffigures

\bibliography{bibdata}
\bibliographystyle{plain}

\end{document}