### 2010-06-17 src/latex/annotated.tex
compressible branch

\section{Class List}
Here are the classes, structs, unions and interfaces with brief descriptions:\begin{CompactList}
\item\contentsline{section}{\hyperlink{classBetaSpline}{BetaSpline} (A concrete kernel class )}{\pageref{classBetaSpline}}{}
\item\contentsline{section}{\hyperlink{classBoundary}{Boundary} (\hyperlink{classBoundary}{Boundary} conditions )}{\pageref{classBoundary}}{}
\item\contentsline{section}{\hyperlink{classDiagnose}{Diagnose} (\hyperlink{classOutput}{Output} diagnosal )}{\pageref{classDiagnose}}{}
\item\contentsline{section}{\hyperlink{classForce}{Force} (The class defining force on or between particles )}{\pageref{classForce}}{}
\item\contentsline{section}{\hyperlink{classHydrodynamics}{Hydrodynamics} (Definition of materials and their hydrodynamical interactions )}{\pageref{classHydrodynamics}}{}
\item\contentsline{section}{\hyperlink{classInitiation}{Initiation} (Initiates the simulation )}{\pageref{classInitiation}}{}
\item\contentsline{section}{\hyperlink{classInteraction}{Interaction} (Defines interaction between particles )}{\pageref{classInteraction}}{}
\item\contentsline{section}{\hyperlink{classKernel}{Kernel} (\hyperlink{classKernel}{Kernel} abstract base )}{\pageref{classKernel}}{}
\item\contentsline{section}{\hyperlink{classLlist}{Llist$<$ Ldata $>$} (Double linked list implementation )}{\pageref{classLlist}}{}
\item\contentsline{section}{\hyperlink{classLlistNode}{LlistNode$<$ Ldata $>$} (A template node in the linked list )}{\pageref{classLlistNode}}{}
\item\contentsline{section}{\hyperlink{classMaterial}{Material} (\hyperlink{classMaterial}{Material} )}{\pageref{classMaterial}}{}
\item\contentsline{section}{\hyperlink{classMLS}{MLS} (Moving Least Squares (\hyperlink{classMLS}{MLS}) class )}{\pageref{classMLS}}{}
\item\contentsline{section}{\hyperlink{classOutput}{Output} (\hyperlink{classOutput}{Output} class )}{\pageref{classOutput}}{}
\item\contentsline{section}{\hyperlink{classParticle}{Particle} (\hyperlink{classParticle}{Particle} class )}{\pageref{classParticle}}{}
\item\contentsline{section}{\hyperlink{classParticleManager}{ParticleManager} (\hyperlink{classParticle}{Particle} manager class )}{\pageref{classParticleManager}}{}
\item\contentsline{section}{\hyperlink{classQuinticSpline}{QuinticSpline} (Quintic spline class )}{\pageref{classQuinticSpline}}{}
\item\contentsline{section}{\hyperlink{classTimeSolver}{TimeSolver} (Time solver class )}{\pageref{classTimeSolver}}{}
\item\contentsline{section}{\hyperlink{classWiener}{Wiener} (\hyperlink{classWiener}{Wiener} process )}{\pageref{classWiener}}{}
\end{CompactList}