### 2010-06-17 src/latex/classBetaSpline.tex

\hypertarget{classBetaSpline}{
\section{BetaSpline Class Reference}
\label{classBetaSpline}\index{BetaSpline@{BetaSpline}}
}
A concrete kernel class.

{\tt \#include $<$betaspline.h$>$}

Inheritance diagram for BetaSpline:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=106pt]{classBetaSpline__inherit__graph}
\end{center}
\end{figure}
Collaboration diagram for BetaSpline:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=106pt]{classBetaSpline__coll__graph}
\end{center}
\end{figure}
\subsection*{Public Member Functions}
\begin{CompactItemize}
\item 
\hyperlink{classBetaSpline_00e744e584e6e7291813ab5e59d801ec}{BetaSpline} (double \hyperlink{classKernel_f088821dba4a00d1d320b27ebcb71258}{smoothingLength})
\begin{CompactList}\small\item\em constructor to initialize the data members and \item\end{CompactList}\item 
virtual double \hyperlink{classBetaSpline_797016b36239b0be98c3aa0715bb045a}{w} (double distance) const 
\begin{CompactList}\small\item\em Calculates the kernel value for the given distance of two particles. \item\end{CompactList}\item 
virtual Vec2d \hyperlink{classBetaSpline_409d22f7b4165a612ffbb62400e55db7}{gradW} (double distance, const Vec2d \&distanceVector) const 
\begin{CompactList}\small\item\em Calculates the kernel derivation for the given distance of two particles. \item\end{CompactList}\item 
\hypertarget{classBetaSpline_5b01a3303b3d45e74cdbff21419a07ff}{
double \hyperlink{classBetaSpline_5b01a3303b3d45e74cdbff21419a07ff}{F} (double distance) const }
\label{classBetaSpline_5b01a3303b3d45e74cdbff21419a07ff}

\begin{CompactList}\small\item\em Calculates the kernel derivation to distance. \item\end{CompactList}\end{CompactItemize}
\subsection*{Private Attributes}
\begin{CompactItemize}
\item 
\hypertarget{classBetaSpline_3cc93fb702c723e280f4c55aa7422def}{
double \hyperlink{classBetaSpline_3cc93fb702c723e280f4c55aa7422def}{norm}}
\label{classBetaSpline_3cc93fb702c723e280f4c55aa7422def}

\begin{CompactList}\small\item\em Normalization factor. \item\end{CompactList}\item 
\hypertarget{classBetaSpline_ca1e84910928d93dc2b7f4586f1102ac}{
double \hyperlink{classBetaSpline_ca1e84910928d93dc2b7f4586f1102ac}{reciprocH}}
\label{classBetaSpline_ca1e84910928d93dc2b7f4586f1102ac}

\begin{CompactList}\small\item\em Auxiliary factors for intermediate results: The inverse smoothing length $\ast$/. \item\end{CompactList}\item 
\hypertarget{classBetaSpline_06df50a142cacc03bfd3168cc6ae1f57}{
double \hyperlink{classBetaSpline_06df50a142cacc03bfd3168cc6ae1f57}{factorW}}
\label{classBetaSpline_06df50a142cacc03bfd3168cc6ae1f57}

\begin{CompactList}\small\item\em Auxiliary factors for intermediate results: A pre-factor for w $\ast$/. \item\end{CompactList}\item 
\hypertarget{classBetaSpline_0b285867c4f1d4b9fced5847650a47f8}{
double \hyperlink{classBetaSpline_0b285867c4f1d4b9fced5847650a47f8}{factorGradW}}
\label{classBetaSpline_0b285867c4f1d4b9fced5847650a47f8}

\begin{CompactList}\small\item\em Auxiliary factors for intermediate results: A pre-factor for grad w $\ast$/. \item\end{CompactList}\end{CompactItemize}

\subsection{Detailed Description}
A concrete kernel class.

see Monaghan \& Lattanzio (1985) most often used kernel implemented in \hyperlink{bskernel_8cpp}{bskernel.cpp}

\subsection{Constructor \& Destructor Documentation}
\hypertarget{classBetaSpline_00e744e584e6e7291813ab5e59d801ec}{
\index{BetaSpline@{BetaSpline}!BetaSpline@{BetaSpline}}
\index{BetaSpline@{BetaSpline}!BetaSpline@{BetaSpline}}
\subsubsection[{BetaSpline}]{\setlength{\rightskip}{0pt plus 5cm}BetaSpline::BetaSpline (double {\em smoothingLength})}}
\label{classBetaSpline_00e744e584e6e7291813ab5e59d801ec}

constructor to initialize the data members and

\begin{itemize}
\item initialize the auxiliary factors\end{itemize}

\begin{itemize}
\item normalize to 1.0 \end{itemize}

\subsection{Member Function Documentation}
\hypertarget{classBetaSpline_409d22f7b4165a612ffbb62400e55db7}{
\index{BetaSpline@{BetaSpline}!gradW@{gradW}}
\index{gradW@{gradW}!BetaSpline@{BetaSpline}}
\subsubsection[{gradW}]{\setlength{\rightskip}{0pt plus 5cm}Vec2d BetaSpline::gradW (double {\em distance}, \/  const Vec2d \& {\em distanceVector}) const\hspace{0.3cm}{\tt  \mbox{[}virtual\mbox{]}}}}
\label{classBetaSpline_409d22f7b4165a612ffbb62400e55db7}

Calculates the kernel derivation for the given distance of two particles.

We take this from Monaghan \& Lattenzio (1985) but used a doubled smoothing length for the definition of the interaction radius.

Implements \hyperlink{classKernel_545e61b98db05db6dd02b88e2ad3da23}{Kernel}.\hypertarget{classBetaSpline_797016b36239b0be98c3aa0715bb045a}{
\index{BetaSpline@{BetaSpline}!w@{w}}
\index{w@{w}!BetaSpline@{BetaSpline}}
\subsubsection[{w}]{\setlength{\rightskip}{0pt plus 5cm}double BetaSpline::w (double {\em distance}) const\hspace{0.3cm}{\tt  \mbox{[}virtual\mbox{]}}}}
\label{classBetaSpline_797016b36239b0be98c3aa0715bb045a}

Calculates the kernel value for the given distance of two particles.

We take this from Monaghan \& Lattenzio (1985) but used a doubled smoothing length for the definition of the interaction radius.

Implements \hyperlink{classKernel_6b4d26ba99457a4acbbbca1454dbdc8a}{Kernel}.

The documentation for this class was generated from the following files:\begin{CompactItemize}
\item 
\hyperlink{betaspline_8h}{betaspline.h}\item 
\hyperlink{bskernel_8cpp}{bskernel.cpp}\end{CompactItemize}