### 2010-06-17 src/latex/classBoundary.tex

\hypertarget{classBoundary}{
\section{Boundary Class Reference}
\label{classBoundary}\index{Boundary@{Boundary}}
}
\hyperlink{classBoundary}{Boundary} conditions.

{\tt \#include $<$boundary.h$>$}

Collaboration diagram for Boundary:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=192pt]{classBoundary__coll__graph}
\end{center}
\end{figure}
\subsection*{Public Member Functions}
\begin{CompactItemize}
\item 
\hyperlink{classBoundary_8fe88473063d74e67a5932f2e4d2e517}{Boundary} (\hyperlink{classInitiation}{Initiation} \&ini, \hyperlink{classHydrodynamics}{Hydrodynamics} \&hydro, \hyperlink{classParticleManager}{ParticleManager} \&particles)
\begin{CompactList}\small\item\em boundary particle list for all boundray particles \item\end{CompactList}\item 
void \hyperlink{classBoundary_8a40f99b73f3622cd5f1387fcbbdb824}{BuildBoundaryParticles} (\hyperlink{classParticleManager}{ParticleManager} \&particles, \hyperlink{classHydrodynamics}{Hydrodynamics} \&hydro)
\begin{CompactList}\small\item\em build boundary particles \item\end{CompactList}\item 
void \hyperlink{classBoundary_bedd190ee91482e4e59b2a613ab25d57}{BoundaryCondition} (\hyperlink{classParticleManager}{ParticleManager} \&particles)
\begin{CompactList}\small\item\em boundary conditions \item\end{CompactList}\item 
void \hyperlink{classBoundary_ac37e18aaf60503a66173d9428ed9e54}{RunAwayCheck} (\hyperlink{classHydrodynamics}{Hydrodynamics} \&hydro)
\begin{CompactList}\small\item\em check particle if particle run out of the computational domain \item\end{CompactList}\end{CompactItemize}
\subsection*{Public Attributes}
\begin{CompactItemize}
\item 
int \hyperlink{classBoundary_10dd58ba0715412968c4902fe80b1318}{xBl}
\begin{CompactList}\small\item\em boundary condition indicator left hand side \item\end{CompactList}\item 
int \hyperlink{classBoundary_4a582b489cb580b8a6f3a7123f411a53}{xBr}
\begin{CompactList}\small\item\em boundary condition indicator right hand side \item\end{CompactList}\item 
int \hyperlink{classBoundary_0329bf6b954b3001418391cc16a9194e}{yBd}
\begin{CompactList}\small\item\em boundary condition indicator bottom side \item\end{CompactList}\item 
int \hyperlink{classBoundary_6ccff2a8c5a10104d202504034af9e84}{yBu}
\begin{CompactList}\small\item\em boundary condition indicator upper side \item\end{CompactList}\item 
\hypertarget{classBoundary_238a4e6e7358e1f9a2375d8df08a5c7a}{
Vec2d \hyperlink{classBoundary_238a4e6e7358e1f9a2375d8df08a5c7a}{UxBl}}
\label{classBoundary_238a4e6e7358e1f9a2375d8df08a5c7a}

\begin{CompactList}\small\item\em left hand side boundary velocity \item\end{CompactList}\item 
\hypertarget{classBoundary_375a949ddac9f633af007daa7d323099}{
Vec2d \hyperlink{classBoundary_375a949ddac9f633af007daa7d323099}{UxBr}}
\label{classBoundary_375a949ddac9f633af007daa7d323099}

\begin{CompactList}\small\item\em right hand side boundary velocity \item\end{CompactList}\item 
\hypertarget{classBoundary_cf4a2665428f1516626bf683d1eeb4ba}{
Vec2d \hyperlink{classBoundary_cf4a2665428f1516626bf683d1eeb4ba}{UyBd}}
\label{classBoundary_cf4a2665428f1516626bf683d1eeb4ba}

\begin{CompactList}\small\item\em bottom side boundary velocity \item\end{CompactList}\item 
\hypertarget{classBoundary_47eeae29ae83c1ec9ed3a0fdd96bec23}{
Vec2d \hyperlink{classBoundary_47eeae29ae83c1ec9ed3a0fdd96bec23}{UyBu}}
\label{classBoundary_47eeae29ae83c1ec9ed3a0fdd96bec23}

\begin{CompactList}\small\item\em upper side boundary velocity \item\end{CompactList}\item 
\hypertarget{classBoundary_03c656384691f3575170447ab267b8a5}{
\hyperlink{classLlist}{Llist}$<$ \hyperlink{classParticle}{Particle} $>$ \hyperlink{classBoundary_03c656384691f3575170447ab267b8a5}{boundary\_\-particle\_\-list}}
\label{classBoundary_03c656384691f3575170447ab267b8a5}

\begin{CompactList}\small\item\em boundary particle lists \item\end{CompactList}\end{CompactItemize}
\subsection*{Private Member Functions}
\begin{CompactItemize}
\item 
\hypertarget{classBoundary_bdce0051ff1e2651413242877a4fc507}{
void \hyperlink{classBoundary_bdce0051ff1e2651413242877a4fc507}{non\_\-dimensionalize} (\hyperlink{classInitiation}{Initiation} \&ini)}
\label{classBoundary_bdce0051ff1e2651413242877a4fc507}

\begin{CompactList}\small\item\em non-dimensionalize \item\end{CompactList}\item 
void \hyperlink{classBoundary_116df4b717184d962f40bdc88de26763}{show\_\-information} (\hyperlink{classInitiation}{Initiation} \&ini)
\begin{CompactList}\small\item\em show information on screen \item\end{CompactList}\item 
\hypertarget{classBoundary_8879fae60c663d0bdf55acf44516cd42}{
void \hyperlink{classBoundary_8879fae60c663d0bdf55acf44516cd42}{Boundary\_\-W} (\hyperlink{classParticle}{Particle} $\ast$prtl)}
\label{classBoundary_8879fae60c663d0bdf55acf44516cd42}

\begin{CompactList}\small\item\em implement west side boundary by modifying particle states \item\end{CompactList}\item 
\hypertarget{classBoundary_e721f2c23112698b3cee9a3848f3deb1}{
void \hyperlink{classBoundary_e721f2c23112698b3cee9a3848f3deb1}{Boundary\_\-E} (\hyperlink{classParticle}{Particle} $\ast$prtl)}
\label{classBoundary_e721f2c23112698b3cee9a3848f3deb1}

\begin{CompactList}\small\item\em implement east side boundary by modifying particle states \item\end{CompactList}\item 
\hypertarget{classBoundary_50c82df28f8e11186b2f3e6f0bd971b0}{
void \hyperlink{classBoundary_50c82df28f8e11186b2f3e6f0bd971b0}{Boundary\_\-S} (\hyperlink{classParticle}{Particle} $\ast$prtl)}
\label{classBoundary_50c82df28f8e11186b2f3e6f0bd971b0}

\begin{CompactList}\small\item\em implement south side boundary by modifying particle states \item\end{CompactList}\item 
\hypertarget{classBoundary_ec89b3e4212c60f0b6a42199a844d032}{
void \hyperlink{classBoundary_ec89b3e4212c60f0b6a42199a844d032}{Boundary\_\-N} (\hyperlink{classParticle}{Particle} $\ast$prtl)}
\label{classBoundary_ec89b3e4212c60f0b6a42199a844d032}

\begin{CompactList}\small\item\em implement north side boundary by modifying particle states \item\end{CompactList}\item 
\hypertarget{classBoundary_5ec3b43fbe487290b3b3c584ffdfbd54}{
void \hyperlink{classBoundary_5ec3b43fbe487290b3b3c584ffdfbd54}{Boundary\_\-SW} (\hyperlink{classParticle}{Particle} $\ast$prtl)}
\label{classBoundary_5ec3b43fbe487290b3b3c584ffdfbd54}

\begin{CompactList}\small\item\em implement south-west corner boundary by modifying particle states \item\end{CompactList}\item 
\hypertarget{classBoundary_8a334d5f6950a17261504751a706ec78}{
void \hyperlink{classBoundary_8a334d5f6950a17261504751a706ec78}{Boundary\_\-SE} (\hyperlink{classParticle}{Particle} $\ast$prtl)}
\label{classBoundary_8a334d5f6950a17261504751a706ec78}

\begin{CompactList}\small\item\em implement south-east corner boundary by modifying particle states \item\end{CompactList}\item 
\hypertarget{classBoundary_75bfad6f3ef080a83ca3c0746f94ca4e}{
void \hyperlink{classBoundary_75bfad6f3ef080a83ca3c0746f94ca4e}{Boundary\_\-NW} (\hyperlink{classParticle}{Particle} $\ast$prtl)}
\label{classBoundary_75bfad6f3ef080a83ca3c0746f94ca4e}

\begin{CompactList}\small\item\em implement north-west corner boundary by modifying particle states \item\end{CompactList}\item 
\hypertarget{classBoundary_7ce0aad430a89265e89b76c79ab09dc9}{
void \hyperlink{classBoundary_7ce0aad430a89265e89b76c79ab09dc9}{Boundary\_\-NE} (\hyperlink{classParticle}{Particle} $\ast$prtl)}
\label{classBoundary_7ce0aad430a89265e89b76c79ab09dc9}

\begin{CompactList}\small\item\em implement north-east corner boundary by modifying particle states \item\end{CompactList}\end{CompactItemize}
\subsection*{Private Attributes}
\begin{CompactItemize}
\item 
\hypertarget{classBoundary_0033d4ccb76b4fa29d2addda930dd572}{
int \hyperlink{classBoundary_0033d4ccb76b4fa29d2addda930dd572}{wall\_\-file}}
\label{classBoundary_0033d4ccb76b4fa29d2addda930dd572}

\begin{CompactList}\small\item\em if read wall particle from file .wll \item\end{CompactList}\item 
\hypertarget{classBoundary_e9c27eae4f9ac0853a089aa4d0c76904}{
Vec2d \hyperlink{classBoundary_e9c27eae4f9ac0853a089aa4d0c76904}{box\_\-size}}
\label{classBoundary_e9c27eae4f9ac0853a089aa4d0c76904}

\begin{CompactList}\small\item\em computational domain size \item\end{CompactList}\item 
\hypertarget{classBoundary_d6bdf67458e91f78f5679f450eee8e29}{
int \hyperlink{classBoundary_d6bdf67458e91f78f5679f450eee8e29}{x\_\-clls}}
\label{classBoundary_d6bdf67458e91f78f5679f450eee8e29}

\begin{CompactList}\small\item\em cell matrix size \item\end{CompactList}\item 
\hypertarget{classBoundary_9066d15cdf92a6d522344e037ed51503}{
int \textbf{y\_\-clls}}
\label{classBoundary_9066d15cdf92a6d522344e037ed51503}

\item 
\hypertarget{classBoundary_ad5dca4ff83933fafee8a858011ab56a}{
int \textbf{number\_\-of\_\-materials}}
\label{classBoundary_ad5dca4ff83933fafee8a858011ab56a}

\end{CompactItemize}

\subsection{Detailed Description}
\hyperlink{classBoundary}{Boundary} conditions.

\subsection{Constructor \& Destructor Documentation}
\hypertarget{classBoundary_8fe88473063d74e67a5932f2e4d2e517}{
\index{Boundary@{Boundary}!Boundary@{Boundary}}
\index{Boundary@{Boundary}!Boundary@{Boundary}}
\subsubsection[{Boundary}]{\setlength{\rightskip}{0pt plus 5cm}Boundary::Boundary ({\bf Initiation} \& {\em ini}, \/  {\bf Hydrodynamics} \& {\em hydro}, \/  {\bf ParticleManager} \& {\em particles})}}
\label{classBoundary_8fe88473063d74e67a5932f2e4d2e517}

boundary particle list for all boundray particles

constructor

copy global properties from initiation class

\begin{itemize}
\item reading key words and configuration data from input file\end{itemize}

\begin{itemize}
\item compare the key words for left, right, upper and lower boundary condition type and their velocities\end{itemize}

\begin{itemize}
\item build boundary particles \end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=262pt]{classBoundary_8fe88473063d74e67a5932f2e4d2e517_cgraph}
\end{center}
\end{figure}

\subsection{Member Function Documentation}
\hypertarget{classBoundary_bedd190ee91482e4e59b2a613ab25d57}{
\index{Boundary@{Boundary}!BoundaryCondition@{BoundaryCondition}}
\index{BoundaryCondition@{BoundaryCondition}!Boundary@{Boundary}}
\subsubsection[{BoundaryCondition}]{\setlength{\rightskip}{0pt plus 5cm}void Boundary::BoundaryCondition ({\bf ParticleManager} \& {\em particles})}}
\label{classBoundary_bedd190ee91482e4e59b2a613ab25d57}

boundary conditions

\begin{itemize}
\item west side

\begin{itemize}
\item the rigid wall conditions\end{itemize}

\begin{itemize}
\item the perodic or symmetry conditions \end{itemize}

\item east side \begin{itemize}
\item the rigid wall conditions

\item the perodic or symmetry conditions\end{itemize}

\item south side

\begin{itemize}
\item the rigid wall conditions

\item the perodic or symmetry conditions\end{itemize}

\item north side

\begin{itemize}
\item the rigid wall conditions

\item the perodic or symmetry conditions\end{itemize}
\end{itemize}

considering the coner cells

\begin{itemize}
\item south-west corner \begin{itemize}
\item the rigid wall conditions

\item the perodic or symmetry conditions\end{itemize}

\item north-west corner \begin{itemize}
\item the rigid wall conditions

\item the perodic or symmetry conditions\end{itemize}

\item north-east corner \begin{itemize}
\item the rigid wall conditions

\item the perodic or symmetry conditions\end{itemize}

\item south-east corner \begin{itemize}
\item the rigid wall conditions

\item the perodic or symmetry conditions\end{itemize}
\end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=176pt]{classBoundary_bedd190ee91482e4e59b2a613ab25d57_cgraph}
\end{center}
\end{figure}

Here is the caller graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=238pt]{classBoundary_bedd190ee91482e4e59b2a613ab25d57_icgraph}
\end{center}
\end{figure}
\hypertarget{classBoundary_8a40f99b73f3622cd5f1387fcbbdb824}{
\index{Boundary@{Boundary}!BuildBoundaryParticles@{BuildBoundaryParticles}}
\index{BuildBoundaryParticles@{BuildBoundaryParticles}!Boundary@{Boundary}}
\subsubsection[{BuildBoundaryParticles}]{\setlength{\rightskip}{0pt plus 5cm}void Boundary::BuildBoundaryParticles ({\bf ParticleManager} \& {\em particles}, \/  {\bf Hydrodynamics} \& {\em hydro})}}
\label{classBoundary_8a40f99b73f3622cd5f1387fcbbdb824}

build boundary particles

\begin{itemize}
\item clear boundary particles list

\item boundary condition for each side

\begin{itemize}
\item west side: test boundary parameter for one of the following cases and then build the appropriate boundary particles

\begin{itemize}
\item the rigid wall conditions

\item the symmetry conditions

\item the perodic conditions \end{itemize}

\item east side

\begin{itemize}
\item the rigid wall conditions

\item the symmetry conditions

\item the perodic conditions\end{itemize}

\item south side

\begin{itemize}
\item the rigid wall conditions

\item the symmetry conditions

\item the perodic conditions\end{itemize}

\item north side

\begin{itemize}
\item the rigid wall conditions

\item the symmetry conditions

\item the perodic conditions\end{itemize}
\end{itemize}

\item considering the coner cells

\begin{itemize}
\item south-west corner \begin{itemize}
\item the rigid wall conditions

\item the symmetry conditions

\item the perodic conditions \end{itemize}

\item north-west corner

\begin{itemize}
\item the rigid wall conditions

\item the symmetry conditions

\item the perodic conditions\end{itemize}

\item north-east corner

\begin{itemize}
\item the rigid wall conditions

\item the symmetry conditions

\item the perodic conditions\end{itemize}

\item south-east corner

\begin{itemize}
\item the rigid wall conditions

\item the symmetry conditions

\item the perodic conditions\end{itemize}
\end{itemize}
\end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=186pt]{classBoundary_8a40f99b73f3622cd5f1387fcbbdb824_cgraph}
\end{center}
\end{figure}

Here is the caller graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=248pt]{classBoundary_8a40f99b73f3622cd5f1387fcbbdb824_icgraph}
\end{center}
\end{figure}
\hypertarget{classBoundary_ac37e18aaf60503a66173d9428ed9e54}{
\index{Boundary@{Boundary}!RunAwayCheck@{RunAwayCheck}}
\index{RunAwayCheck@{RunAwayCheck}!Boundary@{Boundary}}
\subsubsection[{RunAwayCheck}]{\setlength{\rightskip}{0pt plus 5cm}void Boundary::RunAwayCheck ({\bf Hydrodynamics} \& {\em hydro})}}
\label{classBoundary_ac37e18aaf60503a66173d9428ed9e54}

check particle if particle run out of the computational domain

\begin{itemize}
\item iterate the partilce list\end{itemize}

\begin{itemize}
\item only checking real particles (for all boundaries) \end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=142pt]{classBoundary_ac37e18aaf60503a66173d9428ed9e54_cgraph}
\end{center}
\end{figure}

Here is the caller graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=231pt]{classBoundary_ac37e18aaf60503a66173d9428ed9e54_icgraph}
\end{center}
\end{figure}
\hypertarget{classBoundary_116df4b717184d962f40bdc88de26763}{
\index{Boundary@{Boundary}!show\_\-information@{show\_\-information}}
\index{show\_\-information@{show\_\-information}!Boundary@{Boundary}}
\subsubsection[{show\_\-information}]{\setlength{\rightskip}{0pt plus 5cm}void Boundary::show\_\-information ({\bf Initiation} \& {\em ini})\hspace{0.3cm}{\tt  \mbox{[}private\mbox{]}}}}
\label{classBoundary_116df4b717184d962f40bdc88de26763}

show information on screen

\begin{itemize}
\item output the property parameters to the screen \end{itemize}

Here is the caller graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=162pt]{classBoundary_116df4b717184d962f40bdc88de26763_icgraph}
\end{center}
\end{figure}

\subsection{Member Data Documentation}
\hypertarget{classBoundary_10dd58ba0715412968c4902fe80b1318}{
\index{Boundary@{Boundary}!xBl@{xBl}}
\index{xBl@{xBl}!Boundary@{Boundary}}
\subsubsection[{xBl}]{\setlength{\rightskip}{0pt plus 5cm}int {\bf Boundary::xBl}}}
\label{classBoundary_10dd58ba0715412968c4902fe80b1318}

boundary condition indicator left hand side

\begin{itemize}
\item 0: wall boundary condition\item 1: perodic boundary condition\item 2: free slip wall boundary condition\item 3: symmetry boundary condition \end{itemize}
\hypertarget{classBoundary_4a582b489cb580b8a6f3a7123f411a53}{
\index{Boundary@{Boundary}!xBr@{xBr}}
\index{xBr@{xBr}!Boundary@{Boundary}}
\subsubsection[{xBr}]{\setlength{\rightskip}{0pt plus 5cm}int {\bf Boundary::xBr}}}
\label{classBoundary_4a582b489cb580b8a6f3a7123f411a53}

boundary condition indicator right hand side

\begin{itemize}
\item 0: wall boundary condition\item 1: perodic boundary condition\item 2: free slip wall boundary condition\item 3: symmetry boundary condition \end{itemize}
\hypertarget{classBoundary_0329bf6b954b3001418391cc16a9194e}{
\index{Boundary@{Boundary}!yBd@{yBd}}
\index{yBd@{yBd}!Boundary@{Boundary}}
\subsubsection[{yBd}]{\setlength{\rightskip}{0pt plus 5cm}int {\bf Boundary::yBd}}}
\label{classBoundary_0329bf6b954b3001418391cc16a9194e}

boundary condition indicator bottom side

\begin{itemize}
\item 0: wall boundary condition\item 1: perodic boundary condition\item 2: free slip wall boundary condition\item 3: symmetry boundary condition \end{itemize}
\hypertarget{classBoundary_6ccff2a8c5a10104d202504034af9e84}{
\index{Boundary@{Boundary}!yBu@{yBu}}
\index{yBu@{yBu}!Boundary@{Boundary}}
\subsubsection[{yBu}]{\setlength{\rightskip}{0pt plus 5cm}int {\bf Boundary::yBu}}}
\label{classBoundary_6ccff2a8c5a10104d202504034af9e84}

boundary condition indicator upper side

\begin{itemize}
\item 0: wall boundary condition\item 1: perodic boundary condition\item 2: free slip wall boundary condition\item 3: symmetry boundary condition \end{itemize}

The documentation for this class was generated from the following files:\begin{CompactItemize}
\item 
\hyperlink{boundary_8h}{boundary.h}\item 
\hyperlink{boundary_8cpp}{boundary.cpp}\end{CompactItemize}