### 2010-06-17 src/latex/classDiagnose.tex

\hypertarget{classDiagnose}{
\section{Diagnose Class Reference}
\label{classDiagnose}\index{Diagnose@{Diagnose}}
}
\hyperlink{classOutput}{Output} diagnosal.

{\tt \#include $<$diagnose.h$>$}

Collaboration diagram for Diagnose:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=183pt]{classDiagnose__coll__graph}
\end{center}
\end{figure}
\subsection*{Public Member Functions}
\begin{CompactItemize}
\item 
\hyperlink{classDiagnose_915a3d571504c843748bf5e9a9ec801c}{Diagnose} (\hyperlink{classInitiation}{Initiation} \&ini, \hyperlink{classHydrodynamics}{Hydrodynamics} \&hydro)
\begin{CompactList}\small\item\em constructor \item\end{CompactList}\item 
void \hyperlink{classDiagnose_bf6d65f3716a9d637186f97c62f7218b}{SaveStates} (\hyperlink{classHydrodynamics}{Hydrodynamics} \&hydro)
\begin{CompactList}\small\item\em save the states of a particle in particle states vectors (vx\_\-list, vy\_\-list, rho\_\-list) \item\end{CompactList}\item 
void \hyperlink{classDiagnose_db2f9991031fb8c301f44518c1d851ea}{OutputProfile} (double Time, \hyperlink{classInitiation}{Initiation} \&ini)
\begin{CompactList}\small\item\em output distribution up to the time \item\end{CompactList}\item 
void \hyperlink{classDiagnose_8c14b3fa58083f64be2018bd0462604d}{Average} (\hyperlink{classParticleManager}{ParticleManager} \&particles, \hyperlink{classMLS}{MLS} \&mls, \hyperlink{classQuinticSpline}{QuinticSpline} \&weight\_\-function, \hyperlink{classInitiation}{Initiation} \&ini)
\begin{CompactList}\small\item\em calculate the average values \item\end{CompactList}\item 
void \hyperlink{classDiagnose_dc46dd26362f94e09b956b3affa247c9}{OutputAverage} (double Time, \hyperlink{classInitiation}{Initiation} \&ini)
\begin{CompactList}\small\item\em output the average values \item\end{CompactList}\item 
void \hyperlink{classDiagnose_9c73b32dc8facce488617780250c14bf}{KineticInformation} (double Time, \hyperlink{classInitiation}{Initiation} \&ini, \hyperlink{classHydrodynamics}{Hydrodynamics} \&hydro)
\begin{CompactList}\small\item\em track the globle average kinetic energy, weight center position and velocity \item\end{CompactList}\end{CompactItemize}
\subsection*{Private Member Functions}
\begin{CompactItemize}
\item 
\hypertarget{classDiagnose_ef3922caa185f5d587edad0eaaf655d3}{
void \hyperlink{classDiagnose_ef3922caa185f5d587edad0eaaf655d3}{BuildDistribution} (\hyperlink{classLlist}{Llist}$<$ double $>$ \&list, double dstrb\mbox{[}2\mbox{]}\mbox{[}101\mbox{]})}
\label{classDiagnose_ef3922caa185f5d587edad0eaaf655d3}

\begin{CompactList}\small\item\em build distribution \item\end{CompactList}\end{CompactItemize}
\subsection*{Private Attributes}
\begin{CompactItemize}
\item 
\hypertarget{classDiagnose_ff5505506d498c9a3d41a15b848f4fed}{
double \hyperlink{classDiagnose_ff5505506d498c9a3d41a15b848f4fed}{delta}}
\label{classDiagnose_ff5505506d498c9a3d41a15b848f4fed}

\begin{CompactList}\small\item\em the inital particle distance \item\end{CompactList}\item 
\hypertarget{classDiagnose_392e23f84c8d6c19a62ff729dabc4b08}{
int \hyperlink{classDiagnose_392e23f84c8d6c19a62ff729dabc4b08}{x\_\-cells}}
\label{classDiagnose_392e23f84c8d6c19a62ff729dabc4b08}

\begin{CompactList}\small\item\em cells matrix for real particles \item\end{CompactList}\item 
\hypertarget{classDiagnose_b99ebe8786c68b5ef9477c49e32227ad}{
int \textbf{y\_\-cells}}
\label{classDiagnose_b99ebe8786c68b5ef9477c49e32227ad}

\item 
\hypertarget{classDiagnose_85b126f2b079c38847e1b8a42fe39604}{
int \hyperlink{classDiagnose_85b126f2b079c38847e1b8a42fe39604}{hdelta}}
\label{classDiagnose_85b126f2b079c38847e1b8a42fe39604}

\begin{CompactList}\small\item\em the ration between smoothing length and inital particle distance \item\end{CompactList}\item 
\hypertarget{classDiagnose_ea27d34fbb662ee394b81802ddeca495}{
char \hyperlink{classDiagnose_ea27d34fbb662ee394b81802ddeca495}{Project\_\-name} \mbox{[}25\mbox{]}}
\label{classDiagnose_ea27d34fbb662ee394b81802ddeca495}

\begin{CompactList}\small\item\em the project name \item\end{CompactList}\item 
\hypertarget{classDiagnose_e91dcce776e5182100a0d813c7a490c9}{
int \textbf{number\_\-of\_\-materials}}
\label{classDiagnose_e91dcce776e5182100a0d813c7a490c9}

\item 
\hypertarget{classDiagnose_f67fdb70cb79df3d31d151630c63cef8}{
double \hyperlink{classDiagnose_f67fdb70cb79df3d31d151630c63cef8}{vx\_\-dstrb} \mbox{[}2\mbox{]}\mbox{[}101\mbox{]}}
\label{classDiagnose_f67fdb70cb79df3d31d151630c63cef8}

\begin{CompactList}\small\item\em x-velocity distribution \item\end{CompactList}\item 
\hypertarget{classDiagnose_0ec20a8f26b147df7f6dd74f054518fb}{
double \hyperlink{classDiagnose_0ec20a8f26b147df7f6dd74f054518fb}{vy\_\-dstrb} \mbox{[}2\mbox{]}\mbox{[}101\mbox{]}}
\label{classDiagnose_0ec20a8f26b147df7f6dd74f054518fb}

\begin{CompactList}\small\item\em y-velocity distribution \item\end{CompactList}\item 
\hypertarget{classDiagnose_c3b8eb576691e06bc18f30e1e23b053c}{
double \hyperlink{classDiagnose_c3b8eb576691e06bc18f30e1e23b053c}{rho\_\-dstrb} \mbox{[}2\mbox{]}\mbox{[}101\mbox{]}}
\label{classDiagnose_c3b8eb576691e06bc18f30e1e23b053c}

\begin{CompactList}\small\item\em density distribution \item\end{CompactList}\item 
\hypertarget{classDiagnose_145722f84ce6e10d9e6fc08104824996}{
\hyperlink{classLlist}{Llist}$<$ double $>$ \hyperlink{classDiagnose_145722f84ce6e10d9e6fc08104824996}{vx\_\-list}}
\label{classDiagnose_145722f84ce6e10d9e6fc08104824996}

\begin{CompactList}\small\item\em list for vx states of a given particle \item\end{CompactList}\item 
\hypertarget{classDiagnose_7b7a236d8ac6f9b7a901670a8f7703e0}{
\hyperlink{classLlist}{Llist}$<$ double $>$ \hyperlink{classDiagnose_7b7a236d8ac6f9b7a901670a8f7703e0}{vy\_\-list}}
\label{classDiagnose_7b7a236d8ac6f9b7a901670a8f7703e0}

\begin{CompactList}\small\item\em list for vy states of a given particle \item\end{CompactList}\item 
\hypertarget{classDiagnose_d10e7325edc977979a7db9fdcd34d8fd}{
\hyperlink{classLlist}{Llist}$<$ double $>$ \hyperlink{classDiagnose_d10e7325edc977979a7db9fdcd34d8fd}{rho\_\-list}}
\label{classDiagnose_d10e7325edc977979a7db9fdcd34d8fd}

\begin{CompactList}\small\item\em list for rho states of a given particle \item\end{CompactList}\item 
\hypertarget{classDiagnose_63dce0bdefd95c49730ab011321ad984}{
int \hyperlink{classDiagnose_63dce0bdefd95c49730ab011321ad984}{gridx}}
\label{classDiagnose_63dce0bdefd95c49730ab011321ad984}

\begin{CompactList}\small\item\em average profile mesh size in x-direction \item\end{CompactList}\item 
\hypertarget{classDiagnose_3cda015218278477cbbe85b6e55b4398}{
int \hyperlink{classDiagnose_3cda015218278477cbbe85b6e55b4398}{gridy}}
\label{classDiagnose_3cda015218278477cbbe85b6e55b4398}

\begin{CompactList}\small\item\em average profile mesh size in y-direction \item\end{CompactList}\item 
\hypertarget{classDiagnose_8fbc4859156ad72a70b471edd8326552}{
double $\ast$$\ast$$\ast$ \textbf{U}}
\label{classDiagnose_8fbc4859156ad72a70b471edd8326552}

\item 
\hypertarget{classDiagnose_a08ef64241663163214a68507d578e27}{
int \hyperlink{classDiagnose_a08ef64241663163214a68507d578e27}{n\_\-average}}
\label{classDiagnose_a08ef64241663163214a68507d578e27}

\begin{CompactList}\small\item\em the times of average \item\end{CompactList}\item 
double \hyperlink{classDiagnose_31e87c9b084b73a78c1e2758149800ed}{ttl\_\-m}
\begin{CompactList}\small\item\em total mass, global average kinetic energy, \item\end{CompactList}\item 
\hypertarget{classDiagnose_fb8a8b1acd170da00404455ee0e674df}{
double $\ast$ \hyperlink{classDiagnose_fb8a8b1acd170da00404455ee0e674df}{mtl\_\-m}}
\label{classDiagnose_fb8a8b1acd170da00404455ee0e674df}

\begin{CompactList}\small\item\em ??????{\bf (\hyperlink{diagnose_8h}{diagnose.h}, line 36)} \item\end{CompactList}\item 
\hypertarget{classDiagnose_1784f6a7677b271e8968147df019e8b0}{
double \hyperlink{classDiagnose_1784f6a7677b271e8968147df019e8b0}{glb\_\-ave\_\-Ek}}
\label{classDiagnose_1784f6a7677b271e8968147df019e8b0}

\begin{CompactList}\small\item\em global average kinetic energy \item\end{CompactList}\item 
\hypertarget{classDiagnose_52f9cc1859fecd569c05de78243e45bd}{
Vec2d $\ast$ \hyperlink{classDiagnose_52f9cc1859fecd569c05de78243e45bd}{wght\_\-cntr}}
\label{classDiagnose_52f9cc1859fecd569c05de78243e45bd}

\begin{CompactList}\small\item\em material weight center position \item\end{CompactList}\item 
\hypertarget{classDiagnose_8b02748469646274034f793aab57193b}{
Vec2d $\ast$ \hyperlink{classDiagnose_8b02748469646274034f793aab57193b}{wght\_\-v}}
\label{classDiagnose_8b02748469646274034f793aab57193b}

\begin{CompactList}\small\item\em material weight center velocity \item\end{CompactList}\end{CompactItemize}

\subsection{Detailed Description}
\hyperlink{classOutput}{Output} diagnosal.

\subsection{Constructor \& Destructor Documentation}
\hypertarget{classDiagnose_915a3d571504c843748bf5e9a9ec801c}{
\index{Diagnose@{Diagnose}!Diagnose@{Diagnose}}
\index{Diagnose@{Diagnose}!Diagnose@{Diagnose}}
\subsubsection[{Diagnose}]{\setlength{\rightskip}{0pt plus 5cm}Diagnose::Diagnose ({\bf Initiation} \& {\em ini}, \/  {\bf Hydrodynamics} \& {\em hydro})}}
\label{classDiagnose_915a3d571504c843748bf5e9a9ec801c}

constructor

\begin{itemize}
\item copy parameters from \hyperlink{classInitiation}{Initiation} class

\item copy boundary properties from initiation class

\item calculate total mass

\item produce output file name (\char`\"{}kinetic\_\-info.dat\char`\"{})and file header \end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=128pt]{classDiagnose_915a3d571504c843748bf5e9a9ec801c_cgraph}
\end{center}
\end{figure}

\subsection{Member Function Documentation}
\hypertarget{classDiagnose_8c14b3fa58083f64be2018bd0462604d}{
\index{Diagnose@{Diagnose}!Average@{Average}}
\index{Average@{Average}!Diagnose@{Diagnose}}
\subsubsection[{Average}]{\setlength{\rightskip}{0pt plus 5cm}void Diagnose::Average ({\bf ParticleManager} \& {\em particles}, \/  {\bf MLS} \& {\em mls}, \/  {\bf QuinticSpline} \& {\em weight\_\-function}, \/  {\bf Initiation} \& {\em ini})}}
\label{classDiagnose_8c14b3fa58083f64be2018bd0462604d}

calculate the average values

\begin{itemize}
\item loop the grid points

\begin{itemize}
\item build the NNP\_\-list

\item if the NNP list is not empty run \hyperlink{classMLS}{MLS} approximation

\item iterate this Nearest Neighbor \hyperlink{classParticle}{Particle} list

\begin{itemize}
\item get particle data \end{itemize}

\item clear the NNP\_\-list

\item calculating the averages \end{itemize}
\end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=331pt]{classDiagnose_8c14b3fa58083f64be2018bd0462604d_cgraph}
\end{center}
\end{figure}

Here is the caller graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=214pt]{classDiagnose_8c14b3fa58083f64be2018bd0462604d_icgraph}
\end{center}
\end{figure}
\hypertarget{classDiagnose_9c73b32dc8facce488617780250c14bf}{
\index{Diagnose@{Diagnose}!KineticInformation@{KineticInformation}}
\index{KineticInformation@{KineticInformation}!Diagnose@{Diagnose}}
\subsubsection[{KineticInformation}]{\setlength{\rightskip}{0pt plus 5cm}void Diagnose::KineticInformation (double {\em Time}, \/  {\bf Initiation} \& {\em ini}, \/  {\bf Hydrodynamics} \& {\em hydro})}}
\label{classDiagnose_9c73b32dc8facce488617780250c14bf}

track the globle average kinetic energy, weight center position and velocity

\begin{itemize}
\item produce output file name

\item iterate the partilce list and calculate data

\item output calculated data in \char`\"{}kinetic\_\-info.dat\char`\"{} \end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=166pt]{classDiagnose_9c73b32dc8facce488617780250c14bf_cgraph}
\end{center}
\end{figure}

Here is the caller graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=235pt]{classDiagnose_9c73b32dc8facce488617780250c14bf_icgraph}
\end{center}
\end{figure}
\hypertarget{classDiagnose_dc46dd26362f94e09b956b3affa247c9}{
\index{Diagnose@{Diagnose}!OutputAverage@{OutputAverage}}
\index{OutputAverage@{OutputAverage}!Diagnose@{Diagnose}}
\subsubsection[{OutputAverage}]{\setlength{\rightskip}{0pt plus 5cm}void Diagnose::OutputAverage (double {\em Time}, \/  {\bf Initiation} \& {\em ini})}}
\label{classDiagnose_dc46dd26362f94e09b956b3affa247c9}

output the average values

\begin{itemize}
\item produce output file name\end{itemize}

\begin{itemize}
\item defining header for tecplot(plot software)\end{itemize}

\begin{itemize}
\item loop the grid points and output data \end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=160pt]{classDiagnose_dc46dd26362f94e09b956b3affa247c9_cgraph}
\end{center}
\end{figure}

Here is the caller graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=122pt]{classDiagnose_dc46dd26362f94e09b956b3affa247c9_icgraph}
\end{center}
\end{figure}
\hypertarget{classDiagnose_db2f9991031fb8c301f44518c1d851ea}{
\index{Diagnose@{Diagnose}!OutputProfile@{OutputProfile}}
\index{OutputProfile@{OutputProfile}!Diagnose@{Diagnose}}
\subsubsection[{OutputProfile}]{\setlength{\rightskip}{0pt plus 5cm}void Diagnose::OutputProfile (double {\em Time}, \/  {\bf Initiation} \& {\em ini})}}
\label{classDiagnose_db2f9991031fb8c301f44518c1d851ea}

output distribution up to the time

\begin{itemize}
\item produce output file name \char`\"{}dstrXXXX.dat\char`\"{}\end{itemize}

\begin{itemize}
\item defining file header for tecplot(plot software)\end{itemize}

-build the distribution

\begin{itemize}
\item normalize to unit\end{itemize}

\begin{itemize}
\item output distribution in file \end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=221pt]{classDiagnose_db2f9991031fb8c301f44518c1d851ea_cgraph}
\end{center}
\end{figure}

Here is the caller graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=119pt]{classDiagnose_db2f9991031fb8c301f44518c1d851ea_icgraph}
\end{center}
\end{figure}
\hypertarget{classDiagnose_bf6d65f3716a9d637186f97c62f7218b}{
\index{Diagnose@{Diagnose}!SaveStates@{SaveStates}}
\index{SaveStates@{SaveStates}!Diagnose@{Diagnose}}
\subsubsection[{SaveStates}]{\setlength{\rightskip}{0pt plus 5cm}void Diagnose::SaveStates ({\bf Hydrodynamics} \& {\em hydro})}}
\label{classDiagnose_bf6d65f3716a9d637186f97c62f7218b}

save the states of a particle in particle states vectors (vx\_\-list, vy\_\-list, rho\_\-list)

\begin{itemize}
\item select a given node on the particle list\end{itemize}

\begin{itemize}
\item Insert the states of the particles in the states vectores (vx\_\-list, vy\_\-list, rho\_\-list) \end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=132pt]{classDiagnose_bf6d65f3716a9d637186f97c62f7218b_cgraph}
\end{center}
\end{figure}

Here is the caller graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=221pt]{classDiagnose_bf6d65f3716a9d637186f97c62f7218b_icgraph}
\end{center}
\end{figure}

\subsection{Member Data Documentation}
\hypertarget{classDiagnose_31e87c9b084b73a78c1e2758149800ed}{
\index{Diagnose@{Diagnose}!ttl\_\-m@{ttl\_\-m}}
\index{ttl\_\-m@{ttl\_\-m}!Diagnose@{Diagnose}}
\subsubsection[{ttl\_\-m}]{\setlength{\rightskip}{0pt plus 5cm}double {\bf Diagnose::ttl\_\-m}\hspace{0.3cm}{\tt  \mbox{[}private\mbox{]}}}}
\label{classDiagnose_31e87c9b084b73a78c1e2758149800ed}

total mass, global average kinetic energy,

total mass

The documentation for this class was generated from the following files:\begin{CompactItemize}
\item 
\hyperlink{diagnose_8h}{diagnose.h}\item 
\hyperlink{diagose_8cpp}{diagose.cpp}\end{CompactItemize}