### 2010-06-17 src/latex/classForce.tex

\hypertarget{classForce}{
\section{Force Class Reference}
\label{classForce}\index{Force@{Force}}
}
The class defining force on or between particles.

{\tt \#include $<$force.h$>$}

\subsection*{Public Member Functions}
\begin{CompactItemize}
\item 
\hyperlink{classForce_00983e3bbc206a00bb9253deafc4e424}{Force} ()
\begin{CompactList}\small\item\em shear and bulk slip length \item\end{CompactList}\item 
\hypertarget{classForce_6ce6bddb4bbe52af88b12827324a818d}{
\textbf{Force} (\hyperlink{classInitiation}{Initiation} \&ini)}
\label{classForce_6ce6bddb4bbe52af88b12827324a818d}

\item 
\hypertarget{classForce_75f2e675a6d57e08127d47b6bc31d3d9}{
void \hyperlink{classForce_75f2e675a6d57e08127d47b6bc31d3d9}{non\_\-dimensionalize} (\hyperlink{classInitiation}{Initiation} \&ini)}
\label{classForce_75f2e675a6d57e08127d47b6bc31d3d9}

\begin{CompactList}\small\item\em non-dimensionalize \item\end{CompactList}\end{CompactItemize}
\subsection*{Public Attributes}
\begin{CompactItemize}
\item 
\hypertarget{classForce_7142842359bf1e78499b5f59cff7a600}{
double \hyperlink{classForce_7142842359bf1e78499b5f59cff7a600}{sigma}}
\label{classForce_7142842359bf1e78499b5f59cff7a600}

\begin{CompactList}\small\item\em heat conduction slip length \item\end{CompactList}\item 
\hypertarget{classForce_5b371059110b1baee895f30c24e209f3}{
double \hyperlink{classForce_5b371059110b1baee895f30c24e209f3}{shear\_\-slip}}
\label{classForce_5b371059110b1baee895f30c24e209f3}

\begin{CompactList}\small\item\em surface tension parameters, its dimension is rho$\ast$u$^\wedge$2$\ast$L \item\end{CompactList}\item 
\hypertarget{classForce_707dd63d90d41bd1c0b345f5ba27a047}{
double \textbf{bulk\_\-slip}}
\label{classForce_707dd63d90d41bd1c0b345f5ba27a047}

\end{CompactItemize}
\subsection*{Private Attributes}
\begin{CompactItemize}
\item 
\hypertarget{classForce_ff0b02789964024867fd633fd0e5722a}{
double \hyperlink{classForce_ff0b02789964024867fd633fd0e5722a}{epsilon}}
\label{classForce_ff0b02789964024867fd633fd0e5722a}

\begin{CompactList}\small\item\em interactive force parameters \item\end{CompactList}\item 
\hypertarget{classForce_499d60a627ddd2b9a2909891cd45b78f}{
double \textbf{heat\_\-slip}}
\label{classForce_499d60a627ddd2b9a2909891cd45b78f}

\end{CompactItemize}
\subsection*{Static Private Attributes}
\begin{CompactItemize}
\item 
\hypertarget{classForce_84055a493186144b98ab363bd958095f}{
static int \hyperlink{classForce_84055a493186144b98ab363bd958095f}{number\_\-of\_\-materials} = 0}
\label{classForce_84055a493186144b98ab363bd958095f}

\begin{CompactList}\small\item\em total number of materials \item\end{CompactList}\item 
\hypertarget{classForce_bcab616ac69fe78d87567506dedd8a56}{
static double \hyperlink{classForce_bcab616ac69fe78d87567506dedd8a56}{smoothinglength} = 0.0}
\label{classForce_bcab616ac69fe78d87567506dedd8a56}

\begin{CompactList}\small\item\em smoothinglenth \item\end{CompactList}\end{CompactItemize}
\subsection*{Friends}
\begin{CompactItemize}
\item 
\hypertarget{classForce_652e6d35f07e2091db6c1d01e86e2286}{
class \hyperlink{classForce_652e6d35f07e2091db6c1d01e86e2286}{Hydrodynamics}}
\label{classForce_652e6d35f07e2091db6c1d01e86e2286}

\end{CompactItemize}

\subsection{Detailed Description}
The class defining force on or between particles.

\subsection{Constructor \& Destructor Documentation}
\hypertarget{classForce_00983e3bbc206a00bb9253deafc4e424}{
\index{Force@{Force}!Force@{Force}}
\index{Force@{Force}!Force@{Force}}
\subsubsection[{Force}]{\setlength{\rightskip}{0pt plus 5cm}Force::Force ()}}
\label{classForce_00983e3bbc206a00bb9253deafc4e424}

shear and bulk slip length

constructor

The documentation for this class was generated from the following files:\begin{CompactItemize}
\item 
\hyperlink{force_8h}{force.h}\item 
force.cpp\end{CompactItemize}