### 2010-06-17 src/latex/classHydrodynamics.tex

\hypertarget{classHydrodynamics}{
\section{Hydrodynamics Class Reference}
\label{classHydrodynamics}\index{Hydrodynamics@{Hydrodynamics}}
}
Definition of materials and their hydrodynamical interactions.

{\tt \#include $<$hydrodynamics.h$>$}

Collaboration diagram for Hydrodynamics:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=400pt]{classHydrodynamics__coll__graph}
\end{center}
\end{figure}
\subsection*{Public Member Functions}
\begin{CompactItemize}
\item 
\hyperlink{classHydrodynamics_837d2adb43e7d4ea4fa5a1c89bb59313}{Hydrodynamics} (\hyperlink{classParticleManager}{ParticleManager} \&particles, \hyperlink{classInitiation}{Initiation} \&ini)
\begin{CompactList}\small\item\em constructor \item\end{CompactList}\item 
\hypertarget{classHydrodynamics_281acddb6d4dd1efa9970f24a6bf57cd}{
double \hyperlink{classHydrodynamics_281acddb6d4dd1efa9970f24a6bf57cd}{GetTimestep} ()}
\label{classHydrodynamics_281acddb6d4dd1efa9970f24a6bf57cd}

\begin{CompactList}\small\item\em get the time step \item\end{CompactList}\item 
void \hyperlink{classHydrodynamics_41e1527e65d6f93d81c30467871021f5}{BuildPair} (\hyperlink{classParticleManager}{ParticleManager} \&particles, \hyperlink{classQuinticSpline}{QuinticSpline} \&weight\_\-function)
\begin{CompactList}\small\item\em build new pairs \item\end{CompactList}\item 
void \hyperlink{classHydrodynamics_cf0d749b11e8b8474dd204c6ea080235}{UpdatePair} (\hyperlink{classQuinticSpline}{QuinticSpline} \&weight\_\-function)
\begin{CompactList}\small\item\em update new parameters in pairs \item\end{CompactList}\item 
void \hyperlink{classHydrodynamics_67e2d72d69156a086d43168f5cd58bf7}{ZeroChangeRate} ()
\begin{CompactList}\small\item\em initiate particle change rate \item\end{CompactList}\item 
void \hyperlink{classHydrodynamics_abdebac769f07f2500c21689fafa2981}{AddGravity} ()
\begin{CompactList}\small\item\em add the gravity effects \item\end{CompactList}\item 
void \hyperlink{classHydrodynamics_22df8569d81c4c363029efac143ddb26}{UpdateChangeRate} (\hyperlink{classParticleManager}{ParticleManager} \&particles, \hyperlink{classQuinticSpline}{QuinticSpline} \&weight\_\-function)
\begin{CompactList}\small\item\em calculate interaction with updating interaction list \item\end{CompactList}\item 
void \hyperlink{classHydrodynamics_f744ee07f3f3b511d0db18949160c20a}{UpdateChangeRate} ()
\begin{CompactList}\small\item\em calculate interaction without updating interaction list \item\end{CompactList}\item 
void \hyperlink{classHydrodynamics_13810b3b4110a2b931bf7426af2f9bcf}{Zero\_\-density} ()
\begin{CompactList}\small\item\em initiate particle density to zero \item\end{CompactList}\item 
void \hyperlink{classHydrodynamics_049f88c2f97863edae4489618dc9e853}{Zero\_\-ShearRate} ()
\item 
void \hyperlink{classHydrodynamics_1818eb1f9044ae980e92c70146e51c5b}{UpdateDensity} (\hyperlink{classParticleManager}{ParticleManager} \&particles, \hyperlink{classQuinticSpline}{QuinticSpline} \&weight\_\-function)
\begin{CompactList}\small\item\em summation for particles density (with updating interaction list) \item\end{CompactList}\item 
void \hyperlink{classHydrodynamics_7ae4f1005699f13414ec6c487bac0ee0}{UpdateShearRate} (\hyperlink{classParticleManager}{ParticleManager} \&particles, \hyperlink{classQuinticSpline}{QuinticSpline} \&weight\_\-function)
\begin{CompactList}\small\item\em summation for shear rates (with updating interaction list) \item\end{CompactList}\item 
void \hyperlink{classHydrodynamics_b6cab35a4d7adf70657ef16b9ec4dafd}{UpdateDensity} ()
\begin{CompactList}\small\item\em currently no shear rate calculated without updating interaction list \item\end{CompactList}\item 
void \hyperlink{classHydrodynamics_49570aaeedea32d7a2baf0532880b2e8}{UpdateShearRate} ()
\begin{CompactList}\small\item\em ??? \item\end{CompactList}\item 
void \hyperlink{classHydrodynamics_939e9b2ec26b3e4a2eaf2683e5f0343d}{UpdatePhaseGradient} (\hyperlink{classBoundary}{Boundary} \&boundary)
\begin{CompactList}\small\item\em not independant with UpdateDensity \item\end{CompactList}\item 
void \hyperlink{classHydrodynamics_3c9019f19ccac5370b7b0ccda343b5d2}{Zero\_\-PhaseGradient} (\hyperlink{classBoundary}{Boundary} \&boundary)
\item 
void \hyperlink{classHydrodynamics_c87ff6437d540454ed7de2f726858f4a}{UpdatePhaseField} (\hyperlink{classBoundary}{Boundary} \&boundary)
\item 
void \hyperlink{classHydrodynamics_dedab91b9a62eb7e26931cbbe73f79cc}{Zero\_\-PhaseField} (\hyperlink{classBoundary}{Boundary} \&boundary)
\item 
void \hyperlink{classHydrodynamics_f5b2c4ad14824e4b6192faa6f60f4014}{UpdateSurfaceStress} (\hyperlink{classBoundary}{Boundary} \&boundary)
\item 
void \hyperlink{classHydrodynamics_dbbe86fd4eac49912a9ffb5e284883ee}{UpdatePhaseLaplacian} (\hyperlink{classBoundary}{Boundary} \&boundary)
\item 
void \hyperlink{classHydrodynamics_d1c5fd5b500eacb886a72f2dab13e91b}{Zero\_\-PhaseLaplacian} (\hyperlink{classBoundary}{Boundary} \&boundary)
\item 
double \hyperlink{classHydrodynamics_e6983cf4d86bf7b33b88c8f089caf3a8}{SurfaceTensionCoefficient} ()
\begin{CompactList}\small\item\em calculate surface tension coefficient \item\end{CompactList}\item 
void \hyperlink{classHydrodynamics_a755853b07cbbc5b484d78a3e787d1ef}{UpdatePahseMatrix} (\hyperlink{classBoundary}{Boundary} \&boundary)
\begin{CompactList}\small\item\em this method currently does {\bf NOTHING} \item\end{CompactList}\item 
void \hyperlink{classHydrodynamics_1698dbe8ecc0e730319de5d7eca7a891}{UpdateState} ()
\begin{CompactList}\small\item\em calculate states from conservatives \item\end{CompactList}\item 
void \hyperlink{classHydrodynamics_09a3f5a5c055e0efc822fdf80f40cd7d}{UpdateVolume} (\hyperlink{classParticleManager}{ParticleManager} \&particles, \hyperlink{classQuinticSpline}{QuinticSpline} \&weight\_\-function)
\begin{CompactList}\small\item\em calculate partilce volume \item\end{CompactList}\item 
void \hyperlink{classHydrodynamics_316a6079bf22102d9b911e2632cd2680}{Predictor} (double dt)
\begin{CompactList}\small\item\em predictor method, density evaluated directly \item\end{CompactList}\item 
void \hyperlink{classHydrodynamics_928a3fb7752d458026ed06dca1fbf137}{Corrector} (double dt)
\begin{CompactList}\small\item\em corrector method, density evaluated directly:{\bf  corrector advances p, rho, U} \item\end{CompactList}\item 
void \hyperlink{classHydrodynamics_911fe25b94fec398fad1f2def036d755}{Predictor\_\-summation} (double dt)
\begin{CompactList}\small\item\em for predictor method, density evaluated with summation (that means: no density update within this method) \item\end{CompactList}\item 
void \hyperlink{classHydrodynamics_6a62dad5b8c33b504481a8e4e5b023cb}{Corrector\_\-summation} (double dt)
\begin{CompactList}\small\item\em for corrector method, density evaluated with summation (that means: no density update within this method) \item\end{CompactList}\item 
void \hyperlink{classHydrodynamics_f67579320cbdccaf49bbdf3dfb13f23c}{Zero\_\-Random} ()
\begin{CompactList}\small\item\em initiate random force (DPD simulation) \item\end{CompactList}\item 
\hypertarget{classHydrodynamics_2ea6824bef3da552942cbf0c0653a4c0}{
void \hyperlink{classHydrodynamics_2ea6824bef3da552942cbf0c0653a4c0}{UpdateRandom} (double sqrtdt)}
\label{classHydrodynamics_2ea6824bef3da552942cbf0c0653a4c0}

\begin{CompactList}\small\item\em calculate random interaction without updating interaction list (DPD simulation) \item\end{CompactList}\item 
void \hyperlink{classHydrodynamics_8f826d2003ca7a4782b028b761a26421}{RandomEffects} ()
\begin{CompactList}\small\item\em including random effects (DPD simulation) \item\end{CompactList}\item 
\hypertarget{classHydrodynamics_a737dcae7a8b53a7829da52a6eee593d}{
void \textbf{MovingTest} (\hyperlink{classInitiation}{Initiation} \&ini)}
\label{classHydrodynamics_a737dcae7a8b53a7829da52a6eee593d}

\item 
\hypertarget{classHydrodynamics_1baf3a0ed4c17a3758b50c223b321f8a}{
double \hyperlink{classHydrodynamics_1baf3a0ed4c17a3758b50c223b321f8a}{ConservationTest} ()}
\label{classHydrodynamics_1baf3a0ed4c17a3758b50c223b321f8a}

\begin{CompactList}\small\item\em test for debug \item\end{CompactList}\item 
void \hyperlink{classHydrodynamics_4f65196eb3002238371d0e04dd76c02a}{Zero\_\-Velocity} ()
\begin{CompactList}\small\item\em special uitilities \item\end{CompactList}\end{CompactItemize}
\subsection*{Public Attributes}
\begin{CompactItemize}
\item 
\hypertarget{classHydrodynamics_fddd4c6d09a8be4c90fd914913303331}{
\hyperlink{classMaterial}{Material} $\ast$ \hyperlink{classHydrodynamics_fddd4c6d09a8be4c90fd914913303331}{materials}}
\label{classHydrodynamics_fddd4c6d09a8be4c90fd914913303331}

\begin{CompactList}\small\item\em the materials used \item\end{CompactList}\item 
\hypertarget{classHydrodynamics_7a478cfc3248a8ea35d1819f3169c46f}{
\hyperlink{classForce}{Force} $\ast$$\ast$ \hyperlink{classHydrodynamics_7a478cfc3248a8ea35d1819f3169c46f}{forces}}
\label{classHydrodynamics_7a478cfc3248a8ea35d1819f3169c46f}

\begin{CompactList}\small\item\em the interaction force used \item\end{CompactList}\item 
\hypertarget{classHydrodynamics_75c09c83c42e2af4521db7685b4bfdf6}{
\hyperlink{classLlist}{Llist}$<$ \hyperlink{classParticle}{Particle} $>$ \hyperlink{classHydrodynamics_75c09c83c42e2af4521db7685b4bfdf6}{particle\_\-list}}
\label{classHydrodynamics_75c09c83c42e2af4521db7685b4bfdf6}

\begin{CompactList}\small\item\em particle list for all particles \item\end{CompactList}\item 
\hypertarget{classHydrodynamics_01480327db244d4bd8b5ce7f109f7884}{
\hyperlink{classWiener}{Wiener} \hyperlink{classHydrodynamics_01480327db244d4bd8b5ce7f109f7884}{wiener}}
\label{classHydrodynamics_01480327db244d4bd8b5ce7f109f7884}

\begin{CompactList}\small\item\em \hyperlink{classWiener}{Wiener} process. \item\end{CompactList}\end{CompactItemize}
\subsection*{Private Attributes}
\begin{CompactItemize}
\item 
\hypertarget{classHydrodynamics_74497f44758886505ffdea8972919a6b}{
int \textbf{number\_\-of\_\-materials}}
\label{classHydrodynamics_74497f44758886505ffdea8972919a6b}

\item 
\hypertarget{classHydrodynamics_6f7660b8e3c861315c421a7baac8e81e}{
Vec2d \textbf{gravity}}
\label{classHydrodynamics_6f7660b8e3c861315c421a7baac8e81e}

\item 
\hypertarget{classHydrodynamics_5d05d80cca5039ade949fd272261ed57}{
double \textbf{smoothinglength}}
\label{classHydrodynamics_5d05d80cca5039ade949fd272261ed57}

\item 
\hypertarget{classHydrodynamics_d70d17be4ff45e1b01aaa779771a27e0}{
double \textbf{delta}}
\label{classHydrodynamics_d70d17be4ff45e1b01aaa779771a27e0}

\item 
\hypertarget{classHydrodynamics_74ed6f3fd3d1a615fceea4aa926add37}{
double \textbf{delta2}}
\label{classHydrodynamics_74ed6f3fd3d1a615fceea4aa926add37}

\item 
\hypertarget{classHydrodynamics_d39804d0ef02e7bbced89f22f6980867}{
double \textbf{delta3}}
\label{classHydrodynamics_d39804d0ef02e7bbced89f22f6980867}

\item 
\hypertarget{classHydrodynamics_50784f838fa0e34eca594e9eb99a2bb0}{
double \textbf{dt\_\-g\_\-vis}}
\label{classHydrodynamics_50784f838fa0e34eca594e9eb99a2bb0}

\item 
\hypertarget{classHydrodynamics_666c0e9c93f1b3fbc0d6b8dd0e697f3e}{
double \textbf{dt\_\-surf}}
\label{classHydrodynamics_666c0e9c93f1b3fbc0d6b8dd0e697f3e}

\item 
\hypertarget{classHydrodynamics_5b155497d0ab985266f7dcf4448c2253}{
\hyperlink{classLlist}{Llist}$<$ \hyperlink{classInteraction}{Interaction} $>$ \hyperlink{classHydrodynamics_5b155497d0ab985266f7dcf4448c2253}{interaction\_\-list}}
\label{classHydrodynamics_5b155497d0ab985266f7dcf4448c2253}

\begin{CompactList}\small\item\em the interaction (particle pair) list \item\end{CompactList}\item 
\hypertarget{classHydrodynamics_9f6ba4a1659ccd79a80b5a0291c7c5a0}{
double \hyperlink{classHydrodynamics_9f6ba4a1659ccd79a80b5a0291c7c5a0}{viscosity\_\-max}}
\label{classHydrodynamics_9f6ba4a1659ccd79a80b5a0291c7c5a0}

\begin{CompactList}\small\item\em for first time step \item\end{CompactList}\item 
\hypertarget{classHydrodynamics_a1ead264a05af3f201e712d5d17abccb}{
double \hyperlink{classHydrodynamics_a1ead264a05af3f201e712d5d17abccb}{surface\_\-max}}
\label{classHydrodynamics_a1ead264a05af3f201e712d5d17abccb}

\begin{CompactList}\small\item\em for first time step \item\end{CompactList}\end{CompactItemize}

\subsection{Detailed Description}
Definition of materials and their hydrodynamical interactions.

\subsection{Constructor \& Destructor Documentation}
\hypertarget{classHydrodynamics_837d2adb43e7d4ea4fa5a1c89bb59313}{
\index{Hydrodynamics@{Hydrodynamics}!Hydrodynamics@{Hydrodynamics}}
\index{Hydrodynamics@{Hydrodynamics}!Hydrodynamics@{Hydrodynamics}}
\subsubsection[{Hydrodynamics}]{\setlength{\rightskip}{0pt plus 5cm}Hydrodynamics::Hydrodynamics ({\bf ParticleManager} \& {\em particles}, \/  {\bf Initiation} \& {\em ini})}}
\label{classHydrodynamics_837d2adb43e7d4ea4fa5a1c89bb59313}

constructor

\begin{itemize}
\item copy properties from initiation class

\item create material matrix

\item create the force matrix

\item check if inputfile exists

\item reading all key words and configuration data

\item if key word material: read all materials (from .cfg file)

\begin{itemize}
\item save each one of them in materials matrix

\item output the material property parameters to the screen

\item non-dimensionalize\end{itemize}

\item if key word forces: read all forces for all materials (from .cfg file)

\begin{itemize}
\item save eahc one of them in forces matrix

\item copy smoothing length fro initiation

\item and non-dimensionalize\end{itemize}

\item initialize parameters for time step and the artificial compressiblity

\item determine the artificial compressiblity

\item biuld the real particles \end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=310pt]{classHydrodynamics_837d2adb43e7d4ea4fa5a1c89bb59313_cgraph}
\end{center}
\end{figure}

\subsection{Member Function Documentation}
\hypertarget{classHydrodynamics_abdebac769f07f2500c21689fafa2981}{
\index{Hydrodynamics@{Hydrodynamics}!AddGravity@{AddGravity}}
\index{AddGravity@{AddGravity}!Hydrodynamics@{Hydrodynamics}}
\subsubsection[{AddGravity}]{\setlength{\rightskip}{0pt plus 5cm}void Hydrodynamics::AddGravity ()}}
\label{classHydrodynamics_abdebac769f07f2500c21689fafa2981}

add the gravity effects

\begin{itemize}
\item iterate particles on the real particle list\end{itemize}

\begin{itemize}
\item to each particles dUdt: add the gravity effects \end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=145pt]{classHydrodynamics_abdebac769f07f2500c21689fafa2981_cgraph}
\end{center}
\end{figure}

Here is the caller graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=339pt]{classHydrodynamics_abdebac769f07f2500c21689fafa2981_icgraph}
\end{center}
\end{figure}
\hypertarget{classHydrodynamics_41e1527e65d6f93d81c30467871021f5}{
\index{Hydrodynamics@{Hydrodynamics}!BuildPair@{BuildPair}}
\index{BuildPair@{BuildPair}!Hydrodynamics@{Hydrodynamics}}
\subsubsection[{BuildPair}]{\setlength{\rightskip}{0pt plus 5cm}void Hydrodynamics::BuildPair ({\bf ParticleManager} \& {\em particles}, \/  {\bf QuinticSpline} \& {\em weight\_\-function})}}
\label{classHydrodynamics_41e1527e65d6f93d81c30467871021f5}

build new pairs

\begin{itemize}
\item obtain the interaction pairs by just calling the particles BuildInteraction method \end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=340pt]{classHydrodynamics_41e1527e65d6f93d81c30467871021f5_cgraph}
\end{center}
\end{figure}

Here is the caller graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=230pt]{classHydrodynamics_41e1527e65d6f93d81c30467871021f5_icgraph}
\end{center}
\end{figure}
\hypertarget{classHydrodynamics_928a3fb7752d458026ed06dca1fbf137}{
\index{Hydrodynamics@{Hydrodynamics}!Corrector@{Corrector}}
\index{Corrector@{Corrector}!Hydrodynamics@{Hydrodynamics}}
\subsubsection[{Corrector}]{\setlength{\rightskip}{0pt plus 5cm}void Hydrodynamics::Corrector (double {\em dt})}}
\label{classHydrodynamics_928a3fb7752d458026ed06dca1fbf137}

corrector method, density evaluated directly:{\bf  corrector advances p, rho, U}

\begin{itemize}
\item iterate the real partilce list\end{itemize}

\begin{itemize}
\item for each particle: correction based on values on n step and change rate at n+1/2 \end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=141pt]{classHydrodynamics_928a3fb7752d458026ed06dca1fbf137_cgraph}
\end{center}
\end{figure}

Here is the caller graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=166pt]{classHydrodynamics_928a3fb7752d458026ed06dca1fbf137_icgraph}
\end{center}
\end{figure}
\hypertarget{classHydrodynamics_6a62dad5b8c33b504481a8e4e5b023cb}{
\index{Hydrodynamics@{Hydrodynamics}!Corrector\_\-summation@{Corrector\_\-summation}}
\index{Corrector\_\-summation@{Corrector\_\-summation}!Hydrodynamics@{Hydrodynamics}}
\subsubsection[{Corrector\_\-summation}]{\setlength{\rightskip}{0pt plus 5cm}void Hydrodynamics::Corrector\_\-summation (double {\em dt})}}
\label{classHydrodynamics_6a62dad5b8c33b504481a8e4e5b023cb}

for corrector method, density evaluated with summation (that means: no density update within this method)

\begin{itemize}
\item iterate the real partilce list\end{itemize}

\begin{itemize}
\item for each particle: correction (advances R,U) based on values on n step and change rate at n+1/2 \end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=168pt]{classHydrodynamics_6a62dad5b8c33b504481a8e4e5b023cb_cgraph}
\end{center}
\end{figure}

Here is the caller graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=257pt]{classHydrodynamics_6a62dad5b8c33b504481a8e4e5b023cb_icgraph}
\end{center}
\end{figure}
\hypertarget{classHydrodynamics_316a6079bf22102d9b911e2632cd2680}{
\index{Hydrodynamics@{Hydrodynamics}!Predictor@{Predictor}}
\index{Predictor@{Predictor}!Hydrodynamics@{Hydrodynamics}}
\subsubsection[{Predictor}]{\setlength{\rightskip}{0pt plus 5cm}void Hydrodynamics::Predictor (double {\em dt})}}
\label{classHydrodynamics_316a6079bf22102d9b911e2632cd2680}

predictor method, density evaluated directly

\begin{itemize}
\item iterate the real partilce list

\begin{itemize}
\item save values at step n

\item predict values at step n+1

\item calculate the middle values at step n+1/2\end{itemize}
\end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=141pt]{classHydrodynamics_316a6079bf22102d9b911e2632cd2680_cgraph}
\end{center}
\end{figure}

Here is the caller graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=166pt]{classHydrodynamics_316a6079bf22102d9b911e2632cd2680_icgraph}
\end{center}
\end{figure}
\hypertarget{classHydrodynamics_911fe25b94fec398fad1f2def036d755}{
\index{Hydrodynamics@{Hydrodynamics}!Predictor\_\-summation@{Predictor\_\-summation}}
\index{Predictor\_\-summation@{Predictor\_\-summation}!Hydrodynamics@{Hydrodynamics}}
\subsubsection[{Predictor\_\-summation}]{\setlength{\rightskip}{0pt plus 5cm}void Hydrodynamics::Predictor\_\-summation (double {\em dt})}}
\label{classHydrodynamics_911fe25b94fec398fad1f2def036d755}

for predictor method, density evaluated with summation (that means: no density update within this method)

\begin{itemize}
\item iterate the real partilce list

\begin{itemize}
\item save values (R,U) at step n in intermediate variables .\_\-I

\item predict values at step n+1

\item calculate the middle values at step n+1/2 and save them in \hyperlink{classParticle}{Particle} objects prtl\end{itemize}
\end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=167pt]{classHydrodynamics_911fe25b94fec398fad1f2def036d755_cgraph}
\end{center}
\end{figure}

Here is the caller graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=256pt]{classHydrodynamics_911fe25b94fec398fad1f2def036d755_icgraph}
\end{center}
\end{figure}
\hypertarget{classHydrodynamics_8f826d2003ca7a4782b028b761a26421}{
\index{Hydrodynamics@{Hydrodynamics}!RandomEffects@{RandomEffects}}
\index{RandomEffects@{RandomEffects}!Hydrodynamics@{Hydrodynamics}}
\subsubsection[{RandomEffects}]{\setlength{\rightskip}{0pt plus 5cm}void Hydrodynamics::RandomEffects ()}}
\label{classHydrodynamics_8f826d2003ca7a4782b028b761a26421}

including random effects (DPD simulation)

\begin{itemize}
\item iterate the real partilce list\end{itemize}

\begin{itemize}
\item for each particle: add random velocity \_\-dU \end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=154pt]{classHydrodynamics_8f826d2003ca7a4782b028b761a26421_cgraph}
\end{center}
\end{figure}

Here is the caller graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=179pt]{classHydrodynamics_8f826d2003ca7a4782b028b761a26421_icgraph}
\end{center}
\end{figure}
\hypertarget{classHydrodynamics_e6983cf4d86bf7b33b88c8f089caf3a8}{
\index{Hydrodynamics@{Hydrodynamics}!SurfaceTensionCoefficient@{SurfaceTensionCoefficient}}
\index{SurfaceTensionCoefficient@{SurfaceTensionCoefficient}!Hydrodynamics@{Hydrodynamics}}
\subsubsection[{SurfaceTensionCoefficient}]{\setlength{\rightskip}{0pt plus 5cm}double Hydrodynamics::SurfaceTensionCoefficient ()}}
\label{classHydrodynamics_e6983cf4d86bf7b33b88c8f089caf3a8}

calculate surface tension coefficient

\begin{itemize}
\item iterate particles on the real particle list\end{itemize}

\begin{itemize}
\item calculate phase surface stress for all particles \end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=177pt]{classHydrodynamics_e6983cf4d86bf7b33b88c8f089caf3a8_cgraph}
\end{center}
\end{figure}
\hypertarget{classHydrodynamics_f744ee07f3f3b511d0db18949160c20a}{
\index{Hydrodynamics@{Hydrodynamics}!UpdateChangeRate@{UpdateChangeRate}}
\index{UpdateChangeRate@{UpdateChangeRate}!Hydrodynamics@{Hydrodynamics}}
\subsubsection[{UpdateChangeRate}]{\setlength{\rightskip}{0pt plus 5cm}void Hydrodynamics::UpdateChangeRate ()}}
\label{classHydrodynamics_f744ee07f3f3b511d0db18949160c20a}

calculate interaction without updating interaction list

\begin{itemize}
\item initiate the change rate of each real particle by calling \hyperlink{classHydrodynamics_67e2d72d69156a086d43168f5cd58bf7}{ZeroChangeRate()}\end{itemize}

\begin{itemize}
\item iterate the interaction list\end{itemize}

\begin{itemize}
\item calculate for each pair the pair forces or change rate\end{itemize}

\begin{itemize}
\item include the gravity effects \end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=262pt]{classHydrodynamics_f744ee07f3f3b511d0db18949160c20a_cgraph}
\end{center}
\end{figure}
\hypertarget{classHydrodynamics_22df8569d81c4c363029efac143ddb26}{
\index{Hydrodynamics@{Hydrodynamics}!UpdateChangeRate@{UpdateChangeRate}}
\index{UpdateChangeRate@{UpdateChangeRate}!Hydrodynamics@{Hydrodynamics}}
\subsubsection[{UpdateChangeRate}]{\setlength{\rightskip}{0pt plus 5cm}void Hydrodynamics::UpdateChangeRate ({\bf ParticleManager} \& {\em particles}, \/  {\bf QuinticSpline} \& {\em weight\_\-function})}}
\label{classHydrodynamics_22df8569d81c4c363029efac143ddb26}

calculate interaction with updating interaction list

\begin{itemize}
\item initiate change rate of each real particle by calling ZerpChangeRate()\end{itemize}

\begin{itemize}
\item obtain the interaction pairs\end{itemize}

\begin{itemize}
\item iterate the interaction list\end{itemize}

\begin{itemize}
\item calculate for eahc pair the pair forces or change rate\end{itemize}

\begin{itemize}
\item include the gravity effects by calling \hyperlink{classHydrodynamics_abdebac769f07f2500c21689fafa2981}{AddGravity()} \end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=363pt]{classHydrodynamics_22df8569d81c4c363029efac143ddb26_cgraph}
\end{center}
\end{figure}

Here is the caller graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=252pt]{classHydrodynamics_22df8569d81c4c363029efac143ddb26_icgraph}
\end{center}
\end{figure}
\hypertarget{classHydrodynamics_b6cab35a4d7adf70657ef16b9ec4dafd}{
\index{Hydrodynamics@{Hydrodynamics}!UpdateDensity@{UpdateDensity}}
\index{UpdateDensity@{UpdateDensity}!Hydrodynamics@{Hydrodynamics}}
\subsubsection[{UpdateDensity}]{\setlength{\rightskip}{0pt plus 5cm}void Hydrodynamics::UpdateDensity ()}}
\label{classHydrodynamics_b6cab35a4d7adf70657ef16b9ec4dafd}

currently no shear rate calculated without updating interaction list

\begin{itemize}
\item initiate zero density\end{itemize}

\begin{itemize}
\item iterate the interaction list\end{itemize}

\begin{itemize}
\item calculate for each pair the pair forces or change rate\end{itemize}

\begin{itemize}
\item calulate new pressure by calling \hyperlink{classHydrodynamics_1698dbe8ecc0e730319de5d7eca7a891}{UpdateState()} \end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=251pt]{classHydrodynamics_b6cab35a4d7adf70657ef16b9ec4dafd_cgraph}
\end{center}
\end{figure}
\hypertarget{classHydrodynamics_1818eb1f9044ae980e92c70146e51c5b}{
\index{Hydrodynamics@{Hydrodynamics}!UpdateDensity@{UpdateDensity}}
\index{UpdateDensity@{UpdateDensity}!Hydrodynamics@{Hydrodynamics}}
\subsubsection[{UpdateDensity}]{\setlength{\rightskip}{0pt plus 5cm}void Hydrodynamics::UpdateDensity ({\bf ParticleManager} \& {\em particles}, \/  {\bf QuinticSpline} \& {\em weight\_\-function})}}
\label{classHydrodynamics_1818eb1f9044ae980e92c70146e51c5b}

summation for particles density (with updating interaction list)

\begin{itemize}
\item obtain the interaction pairs\end{itemize}

\begin{itemize}
\item initiate by calling Zero\_\-density method\end{itemize}

\begin{itemize}
\item iterate the interaction list\end{itemize}

\begin{itemize}
\item calculate for each pair the pair forces or change rate by calling SummationDensity() method\end{itemize}

\begin{itemize}
\item calulate new pressure by calling \hyperlink{classHydrodynamics_1698dbe8ecc0e730319de5d7eca7a891}{UpdateState()} Method \end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=352pt]{classHydrodynamics_1818eb1f9044ae980e92c70146e51c5b_cgraph}
\end{center}
\end{figure}

Here is the caller graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=242pt]{classHydrodynamics_1818eb1f9044ae980e92c70146e51c5b_icgraph}
\end{center}
\end{figure}
\hypertarget{classHydrodynamics_a755853b07cbbc5b484d78a3e787d1ef}{
\index{Hydrodynamics@{Hydrodynamics}!UpdatePahseMatrix@{UpdatePahseMatrix}}
\index{UpdatePahseMatrix@{UpdatePahseMatrix}!Hydrodynamics@{Hydrodynamics}}
\subsubsection[{UpdatePahseMatrix}]{\setlength{\rightskip}{0pt plus 5cm}void Hydrodynamics::UpdatePahseMatrix ({\bf Boundary} \& {\em boundary})}}
\label{classHydrodynamics_a755853b07cbbc5b484d78a3e787d1ef}

this method currently does {\bf NOTHING}

\begin{itemize}
\item iterate particles on the real particle list\end{itemize}

\begin{itemize}
\item all phase surface stress (currently no action is performed)\end{itemize}

\begin{itemize}
\item iterate particles on the boundary particle list\end{itemize}

\begin{itemize}
\item all phase surface stress (currently no action is perforemd) \end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=164pt]{classHydrodynamics_a755853b07cbbc5b484d78a3e787d1ef_cgraph}
\end{center}
\end{figure}
\hypertarget{classHydrodynamics_cf0d749b11e8b8474dd204c6ea080235}{
\index{Hydrodynamics@{Hydrodynamics}!UpdatePair@{UpdatePair}}
\index{UpdatePair@{UpdatePair}!Hydrodynamics@{Hydrodynamics}}
\subsubsection[{UpdatePair}]{\setlength{\rightskip}{0pt plus 5cm}void Hydrodynamics::UpdatePair ({\bf QuinticSpline} \& {\em weight\_\-function})}}
\label{classHydrodynamics_cf0d749b11e8b8474dd204c6ea080235}

update new parameters in pairs

\begin{itemize}
\item iterate the interaction list\end{itemize}

\begin{itemize}
\item and for each interactionpair call RenewInteraction method \end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=251pt]{classHydrodynamics_cf0d749b11e8b8474dd204c6ea080235_cgraph}
\end{center}
\end{figure}

Here is the caller graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=234pt]{classHydrodynamics_cf0d749b11e8b8474dd204c6ea080235_icgraph}
\end{center}
\end{figure}
\hypertarget{classHydrodynamics_c87ff6437d540454ed7de2f726858f4a}{
\index{Hydrodynamics@{Hydrodynamics}!UpdatePhaseField@{UpdatePhaseField}}
\index{UpdatePhaseField@{UpdatePhaseField}!Hydrodynamics@{Hydrodynamics}}
\subsubsection[{UpdatePhaseField}]{\setlength{\rightskip}{0pt plus 5cm}void Hydrodynamics::UpdatePhaseField ({\bf Boundary} \& {\em boundary})}}
\label{classHydrodynamics_c87ff6437d540454ed7de2f726858f4a}

\begin{itemize}
\item initiate by calling \hyperlink{classHydrodynamics_dedab91b9a62eb7e26931cbbe73f79cc}{Zero\_\-PhaseField()}\end{itemize}

\begin{itemize}
\item iterate the interaction list\end{itemize}

\begin{itemize}
\item calculate for each pair the pair forces or change rate \end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=262pt]{classHydrodynamics_c87ff6437d540454ed7de2f726858f4a_cgraph}
\end{center}
\end{figure}
\hypertarget{classHydrodynamics_939e9b2ec26b3e4a2eaf2683e5f0343d}{
\index{Hydrodynamics@{Hydrodynamics}!UpdatePhaseGradient@{UpdatePhaseGradient}}
\index{UpdatePhaseGradient@{UpdatePhaseGradient}!Hydrodynamics@{Hydrodynamics}}
\subsubsection[{UpdatePhaseGradient}]{\setlength{\rightskip}{0pt plus 5cm}void Hydrodynamics::UpdatePhaseGradient ({\bf Boundary} \& {\em boundary})}}
\label{classHydrodynamics_939e9b2ec26b3e4a2eaf2683e5f0343d}

not independant with UpdateDensity

\begin{itemize}
\item initiate by callingZer\_\-PhaseGradient()\end{itemize}

\begin{itemize}
\item iterate the interaction list\end{itemize}

\begin{itemize}
\item calculate for each pair the pair forces or change rate \end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=278pt]{classHydrodynamics_939e9b2ec26b3e4a2eaf2683e5f0343d_cgraph}
\end{center}
\end{figure}

Here is the caller graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=258pt]{classHydrodynamics_939e9b2ec26b3e4a2eaf2683e5f0343d_icgraph}
\end{center}
\end{figure}
\hypertarget{classHydrodynamics_dbbe86fd4eac49912a9ffb5e284883ee}{
\index{Hydrodynamics@{Hydrodynamics}!UpdatePhaseLaplacian@{UpdatePhaseLaplacian}}
\index{UpdatePhaseLaplacian@{UpdatePhaseLaplacian}!Hydrodynamics@{Hydrodynamics}}
\subsubsection[{UpdatePhaseLaplacian}]{\setlength{\rightskip}{0pt plus 5cm}void Hydrodynamics::UpdatePhaseLaplacian ({\bf Boundary} \& {\em boundary})}}
\label{classHydrodynamics_dbbe86fd4eac49912a9ffb5e284883ee}

\begin{itemize}
\item initiate zero shear rateby calling \hyperlink{classHydrodynamics_d1c5fd5b500eacb886a72f2dab13e91b}{Zero\_\-PhaseLaplacian()}\end{itemize}

\begin{itemize}
\item iterate the interaction list\end{itemize}

\begin{itemize}
\item calculate for each pair the pair forces or change rate \end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=281pt]{classHydrodynamics_dbbe86fd4eac49912a9ffb5e284883ee_cgraph}
\end{center}
\end{figure}
\hypertarget{classHydrodynamics_49570aaeedea32d7a2baf0532880b2e8}{
\index{Hydrodynamics@{Hydrodynamics}!UpdateShearRate@{UpdateShearRate}}
\index{UpdateShearRate@{UpdateShearRate}!Hydrodynamics@{Hydrodynamics}}
\subsubsection[{UpdateShearRate}]{\setlength{\rightskip}{0pt plus 5cm}void Hydrodynamics::UpdateShearRate ()}}
\label{classHydrodynamics_49570aaeedea32d7a2baf0532880b2e8}

???

\begin{itemize}
\item initiate by calling Zero\_\-ShearRate(\end{itemize}

\begin{itemize}
\item iterate the interaction list\end{itemize}

\begin{itemize}
\item calculate for each pair the pair forces or change rate \end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=258pt]{classHydrodynamics_49570aaeedea32d7a2baf0532880b2e8_cgraph}
\end{center}
\end{figure}
\hypertarget{classHydrodynamics_7ae4f1005699f13414ec6c487bac0ee0}{
\index{Hydrodynamics@{Hydrodynamics}!UpdateShearRate@{UpdateShearRate}}
\index{UpdateShearRate@{UpdateShearRate}!Hydrodynamics@{Hydrodynamics}}
\subsubsection[{UpdateShearRate}]{\setlength{\rightskip}{0pt plus 5cm}void Hydrodynamics::UpdateShearRate ({\bf ParticleManager} \& {\em particles}, \/  {\bf QuinticSpline} \& {\em weight\_\-function})}}
\label{classHydrodynamics_7ae4f1005699f13414ec6c487bac0ee0}

summation for shear rates (with updating interaction list)

\begin{itemize}
\item obtain the interaction pairs\end{itemize}

\begin{itemize}
\item initiate by calling Zero\_\-ShearRate Method\end{itemize}

\begin{itemize}
\item iterate the interaction list\end{itemize}

\begin{itemize}
\item calculate for each pair the pair forces or change rate \end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=359pt]{classHydrodynamics_7ae4f1005699f13414ec6c487bac0ee0_cgraph}
\end{center}
\end{figure}
\hypertarget{classHydrodynamics_1698dbe8ecc0e730319de5d7eca7a891}{
\index{Hydrodynamics@{Hydrodynamics}!UpdateState@{UpdateState}}
\index{UpdateState@{UpdateState}!Hydrodynamics@{Hydrodynamics}}
\subsubsection[{UpdateState}]{\setlength{\rightskip}{0pt plus 5cm}void Hydrodynamics::UpdateState ()}}
\label{classHydrodynamics_1698dbe8ecc0e730319de5d7eca7a891}

calculate states from conservatives

\begin{itemize}
\item iterate particles on the real particle list\end{itemize}

\begin{itemize}
\item calculate pressure for each particle \end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=153pt]{classHydrodynamics_1698dbe8ecc0e730319de5d7eca7a891_cgraph}
\end{center}
\end{figure}

Here is the caller graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=332pt]{classHydrodynamics_1698dbe8ecc0e730319de5d7eca7a891_icgraph}
\end{center}
\end{figure}
\hypertarget{classHydrodynamics_f5b2c4ad14824e4b6192faa6f60f4014}{
\index{Hydrodynamics@{Hydrodynamics}!UpdateSurfaceStress@{UpdateSurfaceStress}}
\index{UpdateSurfaceStress@{UpdateSurfaceStress}!Hydrodynamics@{Hydrodynamics}}
\subsubsection[{UpdateSurfaceStress}]{\setlength{\rightskip}{0pt plus 5cm}void Hydrodynamics::UpdateSurfaceStress ({\bf Boundary} \& {\em boundary})}}
\label{classHydrodynamics_f5b2c4ad14824e4b6192faa6f60f4014}

\begin{itemize}
\item iterate particles on the real particle list\end{itemize}

\begin{itemize}
\item update phase surface stress for all particles on this list\end{itemize}

\begin{itemize}
\item iterate particles on the boundary particle list\end{itemize}

\begin{itemize}
\item update phase surface stress for all particles on this list \end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=167pt]{classHydrodynamics_f5b2c4ad14824e4b6192faa6f60f4014_cgraph}
\end{center}
\end{figure}

Here is the caller graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=256pt]{classHydrodynamics_f5b2c4ad14824e4b6192faa6f60f4014_icgraph}
\end{center}
\end{figure}
\hypertarget{classHydrodynamics_09a3f5a5c055e0efc822fdf80f40cd7d}{
\index{Hydrodynamics@{Hydrodynamics}!UpdateVolume@{UpdateVolume}}
\index{UpdateVolume@{UpdateVolume}!Hydrodynamics@{Hydrodynamics}}
\subsubsection[{UpdateVolume}]{\setlength{\rightskip}{0pt plus 5cm}void Hydrodynamics::UpdateVolume ({\bf ParticleManager} \& {\em particles}, \/  {\bf QuinticSpline} \& {\em weight\_\-function})}}
\label{classHydrodynamics_09a3f5a5c055e0efc822fdf80f40cd7d}

calculate partilce volume

\begin{itemize}
\item iterate particles on the particle list

\begin{itemize}
\item take origin particle

\begin{itemize}
\item sum the weights for all of these particles (because they are the inverse of a volume!?!)\end{itemize}

\item clear the NNP\_\-list\end{itemize}
\end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=246pt]{classHydrodynamics_09a3f5a5c055e0efc822fdf80f40cd7d_cgraph}
\end{center}
\end{figure}
\hypertarget{classHydrodynamics_13810b3b4110a2b931bf7426af2f9bcf}{
\index{Hydrodynamics@{Hydrodynamics}!Zero\_\-density@{Zero\_\-density}}
\index{Zero\_\-density@{Zero\_\-density}!Hydrodynamics@{Hydrodynamics}}
\subsubsection[{Zero\_\-density}]{\setlength{\rightskip}{0pt plus 5cm}void Hydrodynamics::Zero\_\-density ()}}
\label{classHydrodynamics_13810b3b4110a2b931bf7426af2f9bcf}

initiate particle density to zero

\begin{itemize}
\item iterate particles on the real particle list\end{itemize}

\begin{itemize}
\item set for each particle density to zero \end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=149pt]{classHydrodynamics_13810b3b4110a2b931bf7426af2f9bcf_cgraph}
\end{center}
\end{figure}

Here is the caller graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=333pt]{classHydrodynamics_13810b3b4110a2b931bf7426af2f9bcf_icgraph}
\end{center}
\end{figure}
\hypertarget{classHydrodynamics_dedab91b9a62eb7e26931cbbe73f79cc}{
\index{Hydrodynamics@{Hydrodynamics}!Zero\_\-PhaseField@{Zero\_\-PhaseField}}
\index{Zero\_\-PhaseField@{Zero\_\-PhaseField}!Hydrodynamics@{Hydrodynamics}}
\subsubsection[{Zero\_\-PhaseField}]{\setlength{\rightskip}{0pt plus 5cm}void Hydrodynamics::Zero\_\-PhaseField ({\bf Boundary} \& {\em boundary})}}
\label{classHydrodynamics_dedab91b9a62eb7e26931cbbe73f79cc}

\begin{itemize}
\item iterate particles on the real particle list\end{itemize}

\begin{itemize}
\item set all phases for all of these particles to zero\end{itemize}

\begin{itemize}
\item iterate particles on the boundary particle list\end{itemize}

\begin{itemize}
\item set all phases for all of these particles to zero \end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=157pt]{classHydrodynamics_dedab91b9a62eb7e26931cbbe73f79cc_cgraph}
\end{center}
\end{figure}

Here is the caller graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=206pt]{classHydrodynamics_dedab91b9a62eb7e26931cbbe73f79cc_icgraph}
\end{center}
\end{figure}
\hypertarget{classHydrodynamics_3c9019f19ccac5370b7b0ccda343b5d2}{
\index{Hydrodynamics@{Hydrodynamics}!Zero\_\-PhaseGradient@{Zero\_\-PhaseGradient}}
\index{Zero\_\-PhaseGradient@{Zero\_\-PhaseGradient}!Hydrodynamics@{Hydrodynamics}}
\subsubsection[{Zero\_\-PhaseGradient}]{\setlength{\rightskip}{0pt plus 5cm}void Hydrodynamics::Zero\_\-PhaseGradient ({\bf Boundary} \& {\em boundary})}}
\label{classHydrodynamics_3c9019f19ccac5370b7b0ccda343b5d2}

\begin{itemize}
\item iterate particles on the real particle list\end{itemize}

\begin{itemize}
\item set phase gradient to zeo for each of these particles\end{itemize}

\begin{itemize}
\item iterate particles on the boundary particle list\end{itemize}

\begin{itemize}
\item set phase gradient to zeo for each of these particle \end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=165pt]{classHydrodynamics_3c9019f19ccac5370b7b0ccda343b5d2_cgraph}
\end{center}
\end{figure}

Here is the caller graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=365pt]{classHydrodynamics_3c9019f19ccac5370b7b0ccda343b5d2_icgraph}
\end{center}
\end{figure}
\hypertarget{classHydrodynamics_d1c5fd5b500eacb886a72f2dab13e91b}{
\index{Hydrodynamics@{Hydrodynamics}!Zero\_\-PhaseLaplacian@{Zero\_\-PhaseLaplacian}}
\index{Zero\_\-PhaseLaplacian@{Zero\_\-PhaseLaplacian}!Hydrodynamics@{Hydrodynamics}}
\subsubsection[{Zero\_\-PhaseLaplacian}]{\setlength{\rightskip}{0pt plus 5cm}void Hydrodynamics::Zero\_\-PhaseLaplacian ({\bf Boundary} \& {\em boundary})}}
\label{classHydrodynamics_d1c5fd5b500eacb886a72f2dab13e91b}

\begin{itemize}
\item iterate particles on the real particle list\end{itemize}

\begin{itemize}
\item set to zero all phase Laplacians for these particles\end{itemize}

\begin{itemize}
\item iterate particles on the boundary particle list\end{itemize}

\begin{itemize}
\item set to zero all phase Laplacians for these particles \end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=168pt]{classHydrodynamics_d1c5fd5b500eacb886a72f2dab13e91b_cgraph}
\end{center}
\end{figure}

Here is the caller graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=227pt]{classHydrodynamics_d1c5fd5b500eacb886a72f2dab13e91b_icgraph}
\end{center}
\end{figure}
\hypertarget{classHydrodynamics_f67579320cbdccaf49bbdf3dfb13f23c}{
\index{Hydrodynamics@{Hydrodynamics}!Zero\_\-Random@{Zero\_\-Random}}
\index{Zero\_\-Random@{Zero\_\-Random}!Hydrodynamics@{Hydrodynamics}}
\subsubsection[{Zero\_\-Random}]{\setlength{\rightskip}{0pt plus 5cm}void Hydrodynamics::Zero\_\-Random ()}}
\label{classHydrodynamics_f67579320cbdccaf49bbdf3dfb13f23c}

initiate random force (DPD simulation)

\begin{itemize}
\item iterate particles on the real particle list\end{itemize}

\begin{itemize}
\item all random values to zero (so, \_\-dU is random value???) \end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=151pt]{classHydrodynamics_f67579320cbdccaf49bbdf3dfb13f23c_cgraph}
\end{center}
\end{figure}

Here is the caller graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=272pt]{classHydrodynamics_f67579320cbdccaf49bbdf3dfb13f23c_icgraph}
\end{center}
\end{figure}
\hypertarget{classHydrodynamics_049f88c2f97863edae4489618dc9e853}{
\index{Hydrodynamics@{Hydrodynamics}!Zero\_\-ShearRate@{Zero\_\-ShearRate}}
\index{Zero\_\-ShearRate@{Zero\_\-ShearRate}!Hydrodynamics@{Hydrodynamics}}
\subsubsection[{Zero\_\-ShearRate}]{\setlength{\rightskip}{0pt plus 5cm}void Hydrodynamics::Zero\_\-ShearRate ()}}
\label{classHydrodynamics_049f88c2f97863edae4489618dc9e853}

\begin{itemize}
\item iterate particles on the real particle list\end{itemize}

\begin{itemize}
\item set ShearRate to zero for each particle \end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=156pt]{classHydrodynamics_049f88c2f97863edae4489618dc9e853_cgraph}
\end{center}
\end{figure}

Here is the caller graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=203pt]{classHydrodynamics_049f88c2f97863edae4489618dc9e853_icgraph}
\end{center}
\end{figure}
\hypertarget{classHydrodynamics_4f65196eb3002238371d0e04dd76c02a}{
\index{Hydrodynamics@{Hydrodynamics}!Zero\_\-Velocity@{Zero\_\-Velocity}}
\index{Zero\_\-Velocity@{Zero\_\-Velocity}!Hydrodynamics@{Hydrodynamics}}
\subsubsection[{Zero\_\-Velocity}]{\setlength{\rightskip}{0pt plus 5cm}void Hydrodynamics::Zero\_\-Velocity ()}}
\label{classHydrodynamics_4f65196eb3002238371d0e04dd76c02a}

special uitilities

\begin{itemize}
\item iterate particles on the real particle list\end{itemize}

\begin{itemize}
\item all velocities to zero \end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=151pt]{classHydrodynamics_4f65196eb3002238371d0e04dd76c02a_cgraph}
\end{center}
\end{figure}
\hypertarget{classHydrodynamics_67e2d72d69156a086d43168f5cd58bf7}{
\index{Hydrodynamics@{Hydrodynamics}!ZeroChangeRate@{ZeroChangeRate}}
\index{ZeroChangeRate@{ZeroChangeRate}!Hydrodynamics@{Hydrodynamics}}
\subsubsection[{ZeroChangeRate}]{\setlength{\rightskip}{0pt plus 5cm}void Hydrodynamics::ZeroChangeRate ()}}
\label{classHydrodynamics_67e2d72d69156a086d43168f5cd58bf7}

initiate particle change rate

\begin{itemize}
\item iterate particles on the real particle list\end{itemize}

\begin{itemize}
\item set for each particle change rates to zero \end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=157pt]{classHydrodynamics_67e2d72d69156a086d43168f5cd58bf7_cgraph}
\end{center}
\end{figure}

Here is the caller graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=351pt]{classHydrodynamics_67e2d72d69156a086d43168f5cd58bf7_icgraph}
\end{center}
\end{figure}

The documentation for this class was generated from the following files:\begin{CompactItemize}
\item 
\hyperlink{hydrodynamics_8h}{hydrodynamics.h}\item 
\hyperlink{hydrodynamics_8cpp}{hydrodynamics.cpp}\end{CompactItemize}