### 2010-06-17 src/latex/classInitiation.tex

\hypertarget{classInitiation}{
\section{Initiation Class Reference}
\label{classInitiation}\index{Initiation@{Initiation}}
}
Initiates the simulation.

{\tt \#include $<$initiation.h$>$}

\subsection*{Public Member Functions}
\begin{CompactItemize}
\item 
\hyperlink{classInitiation_4db3b8c45d38cf4e188de6aa672e8c0d}{Initiation} (const char $\ast$project\_\-name)
\item 
void \hyperlink{classInitiation_916e5b3e316dad1f524a1327b9e4aeb4}{non\_\-dimensionalize} ()
\begin{CompactList}\small\item\em non-dimensionalize initial condition and parameters \item\end{CompactList}\item 
void \hyperlink{classInitiation_cdbf1374b9b0f0b50a0bb1d9abf76293}{show\_\-information} ()
\begin{CompactList}\small\item\em show information on screen \item\end{CompactList}\item 
void \hyperlink{classInitiation_e80adc7d2105ad67b766fd1127966df3}{VolumeMass} (\hyperlink{classHydrodynamics}{Hydrodynamics} \&hydro, \hyperlink{classParticleManager}{ParticleManager} \&particles, \hyperlink{classQuinticSpline}{QuinticSpline} \&weight\_\-function)
\begin{CompactList}\small\item\em predict the particle volume and mass \item\end{CompactList}\item 
\hypertarget{classInitiation_689c73dd7149dc11207eca5c3109a7cd}{
double \hyperlink{classInitiation_689c73dd7149dc11207eca5c3109a7cd}{non\_\-dms\_\-p} (double p)}
\label{classInitiation_689c73dd7149dc11207eca5c3109a7cd}

\begin{CompactList}\small\item\em a non dimensionalize method \item\end{CompactList}\item 
\hypertarget{classInitiation_deddac317c4755623e4fcef187b3f07a}{
double \hyperlink{classInitiation_deddac317c4755623e4fcef187b3f07a}{non\_\-dms\_\-T} (double T)}
\label{classInitiation_deddac317c4755623e4fcef187b3f07a}

\begin{CompactList}\small\item\em a non dimensionalize method \item\end{CompactList}\item 
\hypertarget{classInitiation_29245385c400c4f0de40707cb07dc0dd}{
double \hyperlink{classInitiation_29245385c400c4f0de40707cb07dc0dd}{non\_\-dms\_\-rho} (double rho)}
\label{classInitiation_29245385c400c4f0de40707cb07dc0dd}

\begin{CompactList}\small\item\em a non dimensionalize method \item\end{CompactList}\item 
\hypertarget{classInitiation_e39448af30a53dd8f4354492be9883fc}{
double \hyperlink{classInitiation_e39448af30a53dd8f4354492be9883fc}{non\_\-dms\_\-mass} (double mass)}
\label{classInitiation_e39448af30a53dd8f4354492be9883fc}

\begin{CompactList}\small\item\em a non dimensionalize method \item\end{CompactList}\item 
\hypertarget{classInitiation_2ec28f27a197bfaeb3937733dbb47b03}{
double \hyperlink{classInitiation_2ec28f27a197bfaeb3937733dbb47b03}{non\_\-dms\_\-time} (double time)}
\label{classInitiation_2ec28f27a197bfaeb3937733dbb47b03}

\begin{CompactList}\small\item\em a non dimensionalize method \item\end{CompactList}\item 
\hypertarget{classInitiation_921fe74dff7653520c5e72cc5c11b353}{
double \hyperlink{classInitiation_921fe74dff7653520c5e72cc5c11b353}{non\_\-dms\_\-length} (double length)}
\label{classInitiation_921fe74dff7653520c5e72cc5c11b353}

\begin{CompactList}\small\item\em a non dimensionalize method \item\end{CompactList}\item 
\hypertarget{classInitiation_1b13c381a36c5d841bf9b4dc108a03f6}{
Vec2d \hyperlink{classInitiation_1b13c381a36c5d841bf9b4dc108a03f6}{non\_\-dms\_\-box\_\-size} (Vec2d \hyperlink{classInitiation_c4ab38f327f64513f4c21cd5781f508a}{box\_\-size})}
\label{classInitiation_1b13c381a36c5d841bf9b4dc108a03f6}

\begin{CompactList}\small\item\em a non dimensionalize method \item\end{CompactList}\item 
\hypertarget{classInitiation_7e4fbcccd66df50f14970074038c2988}{
double \hyperlink{classInitiation_7e4fbcccd66df50f14970074038c2988}{non\_\-dms\_\-velocity} (double velocity)}
\label{classInitiation_7e4fbcccd66df50f14970074038c2988}

\begin{CompactList}\small\item\em a non dimensionalize method \item\end{CompactList}\item 
\hypertarget{classInitiation_f48f16aceb256dd582f8697f21408322}{
double \hyperlink{classInitiation_f48f16aceb256dd582f8697f21408322}{non\_\-dms\_\-kinetic\_\-viscosity} (double nu)}
\label{classInitiation_f48f16aceb256dd582f8697f21408322}

\begin{CompactList}\small\item\em a non dimensionalize method \item\end{CompactList}\item 
\hypertarget{classInitiation_32f485a6951b34ce75bac6422419ebce}{
Vec2d \hyperlink{classInitiation_32f485a6951b34ce75bac6422419ebce}{non\_\-dms\_\-velocity} (Vec2d velocity)}
\label{classInitiation_32f485a6951b34ce75bac6422419ebce}

\begin{CompactList}\small\item\em a non dimensionalize method \item\end{CompactList}\item 
\hypertarget{classInitiation_cc810254f426d6f2ed61ef6edee27d89}{
double \hyperlink{classInitiation_cc810254f426d6f2ed61ef6edee27d89}{non\_\-dms\_\-acceleration} (double acceleration)}
\label{classInitiation_cc810254f426d6f2ed61ef6edee27d89}

\begin{CompactList}\small\item\em a non dimensionalize method \item\end{CompactList}\item 
\hypertarget{classInitiation_68c764e4253f4cc5b66c8e978ef37fc8}{
Vec2d \hyperlink{classInitiation_68c764e4253f4cc5b66c8e978ef37fc8}{non\_\-dms\_\-acceleration} (Vec2d acceleration)}
\label{classInitiation_68c764e4253f4cc5b66c8e978ef37fc8}

\begin{CompactList}\small\item\em a non dimensionalize method \item\end{CompactList}\item 
\hypertarget{classInitiation_0d433208c495509df2e2270804b9c2ef}{
double \hyperlink{classInitiation_0d433208c495509df2e2270804b9c2ef}{non\_\-dms\_\-viscosity} (double mu)}
\label{classInitiation_0d433208c495509df2e2270804b9c2ef}

\begin{CompactList}\small\item\em a non dimensionalize method \item\end{CompactList}\item 
\hypertarget{classInitiation_98ded96b747a9ee7512fcc95702b2580}{
double \hyperlink{classInitiation_98ded96b747a9ee7512fcc95702b2580}{non\_\-dms\_\-heat\_\-ratio} (double cv)}
\label{classInitiation_98ded96b747a9ee7512fcc95702b2580}

\begin{CompactList}\small\item\em a non dimensionalize method \item\end{CompactList}\item 
\hypertarget{classInitiation_7d54bcdd903f0d9a0b602008bf9cd10d}{
double \hyperlink{classInitiation_7d54bcdd903f0d9a0b602008bf9cd10d}{non\_\-dms\_\-heat\_\-conduction} (double kappa)}
\label{classInitiation_7d54bcdd903f0d9a0b602008bf9cd10d}

\begin{CompactList}\small\item\em a non dimensionalize method \item\end{CompactList}\item 
\hypertarget{classInitiation_5e09a2250d66d54ef63def3cb9419eae}{
double \hyperlink{classInitiation_5e09a2250d66d54ef63def3cb9419eae}{non\_\-dms\_\-Boltzmann} (double \hyperlink{sph_8cpp_a50583a118c0a3e801d3e48a45b6a259}{k\_\-bltz})}
\label{classInitiation_5e09a2250d66d54ef63def3cb9419eae}

\begin{CompactList}\small\item\em a non dimensionalize method \item\end{CompactList}\item 
\hypertarget{classInitiation_190a357cdd44440885bd5081cc4236a5}{
double \hyperlink{classInitiation_190a357cdd44440885bd5081cc4236a5}{non\_\-dms\_\-surface} (double sigma)}
\label{classInitiation_190a357cdd44440885bd5081cc4236a5}

\begin{CompactList}\small\item\em a non dimensionalize method \item\end{CompactList}\item 
\hypertarget{classInitiation_282fa06b6cb9769ceb01a2b22976126f}{
double \hyperlink{classInitiation_282fa06b6cb9769ceb01a2b22976126f}{dms\_\-p} (double p\_\-non)}
\label{classInitiation_282fa06b6cb9769ceb01a2b22976126f}

\begin{CompactList}\small\item\em a re-dimensionalize method \item\end{CompactList}\item 
\hypertarget{classInitiation_f769934489a9293b6059072bf0e26ae7}{
double \hyperlink{classInitiation_f769934489a9293b6059072bf0e26ae7}{dms\_\-T} (double T\_\-non)}
\label{classInitiation_f769934489a9293b6059072bf0e26ae7}

\begin{CompactList}\small\item\em a re-dimensionalize method \item\end{CompactList}\item 
\hypertarget{classInitiation_4d77129052a63bf33ad36638aaa47972}{
double \hyperlink{classInitiation_4d77129052a63bf33ad36638aaa47972}{dms\_\-rho} (double rho\_\-non)}
\label{classInitiation_4d77129052a63bf33ad36638aaa47972}

\begin{CompactList}\small\item\em a re-dimensionalize method \item\end{CompactList}\item 
\hypertarget{classInitiation_e28be0f6a53920bdab379301eb414990}{
double \hyperlink{classInitiation_e28be0f6a53920bdab379301eb414990}{dms\_\-mass} (double mass\_\-non)}
\label{classInitiation_e28be0f6a53920bdab379301eb414990}

\begin{CompactList}\small\item\em a re-dimensionalize method \item\end{CompactList}\item 
\hypertarget{classInitiation_0e76e3a9e585398854689d151b4af2d8}{
double \hyperlink{classInitiation_0e76e3a9e585398854689d151b4af2d8}{dms\_\-time} (double time\_\-non)}
\label{classInitiation_0e76e3a9e585398854689d151b4af2d8}

\begin{CompactList}\small\item\em a re-dimensionalize method \item\end{CompactList}\item 
\hypertarget{classInitiation_d6747d9cbd12e6327249606b021a87f1}{
double \hyperlink{classInitiation_d6747d9cbd12e6327249606b021a87f1}{dms\_\-length} (double length\_\-non)}
\label{classInitiation_d6747d9cbd12e6327249606b021a87f1}

\begin{CompactList}\small\item\em a re-dimensionalize method \item\end{CompactList}\item 
\hypertarget{classInitiation_8b5815911256da330c686c9e84e3ccbc}{
Vec2d \hyperlink{classInitiation_8b5815911256da330c686c9e84e3ccbc}{dms\_\-box\_\-size} (Vec2d box\_\-size\_\-non)}
\label{classInitiation_8b5815911256da330c686c9e84e3ccbc}

\begin{CompactList}\small\item\em a re-dimensionalize method \item\end{CompactList}\item 
\hypertarget{classInitiation_1bbabaf118b739c082d9db6e5a0d1021}{
double \hyperlink{classInitiation_1bbabaf118b739c082d9db6e5a0d1021}{dms\_\-velocity} (double velocity\_\-non)}
\label{classInitiation_1bbabaf118b739c082d9db6e5a0d1021}

\begin{CompactList}\small\item\em a re-dimensionalize method \item\end{CompactList}\item 
\hypertarget{classInitiation_ff3a9cd4918e5d823b4a1413b805c056}{
Vec2d \hyperlink{classInitiation_ff3a9cd4918e5d823b4a1413b805c056}{dms\_\-velocity} (Vec2d velocity\_\-non)}
\label{classInitiation_ff3a9cd4918e5d823b4a1413b805c056}

\begin{CompactList}\small\item\em a re-dimensionalize method \item\end{CompactList}\item 
\hypertarget{classInitiation_d08c20a915f7c4878a8e382d54874fb1}{
double \hyperlink{classInitiation_d08c20a915f7c4878a8e382d54874fb1}{dms\_\-energy} (double energy\_\-non)}
\label{classInitiation_d08c20a915f7c4878a8e382d54874fb1}

\begin{CompactList}\small\item\em a re-dimensionalize method \item\end{CompactList}\item 
\hypertarget{classInitiation_7746aa0e130fbfdd4ef72d662eaf4fe5}{
double \hyperlink{classInitiation_7746aa0e130fbfdd4ef72d662eaf4fe5}{dms\_\-acceleration} (double acceleration\_\-non)}
\label{classInitiation_7746aa0e130fbfdd4ef72d662eaf4fe5}

\begin{CompactList}\small\item\em a re-dimensionalize method \item\end{CompactList}\item 
\hypertarget{classInitiation_c82bd7ac7f359587868b80b4498c4220}{
Vec2d \hyperlink{classInitiation_c82bd7ac7f359587868b80b4498c4220}{dms\_\-acceleration} (Vec2d acceleration\_\-non)}
\label{classInitiation_c82bd7ac7f359587868b80b4498c4220}

\begin{CompactList}\small\item\em a re-dimensionalize method \item\end{CompactList}\item 
\hypertarget{classInitiation_f867793b42e943a473d1132d0bbb721a}{
double \hyperlink{classInitiation_f867793b42e943a473d1132d0bbb721a}{dms\_\-viscosity} (double mu\_\-non)}
\label{classInitiation_f867793b42e943a473d1132d0bbb721a}

\begin{CompactList}\small\item\em a re-dimensionalize method \item\end{CompactList}\item 
\hypertarget{classInitiation_cbd8852ba19297b3278e96f6fb9ecb60}{
double \hyperlink{classInitiation_cbd8852ba19297b3278e96f6fb9ecb60}{dms\_\-surface} (double sigma\_\-non)}
\label{classInitiation_cbd8852ba19297b3278e96f6fb9ecb60}

\begin{CompactList}\small\item\em a re-dimensionalize method \item\end{CompactList}\end{CompactItemize}
\subsection*{Public Attributes}
\begin{CompactItemize}
\item 
\hypertarget{classInitiation_d011ce1385a1cd46fb500d9a62de6883}{
char \hyperlink{classInitiation_d011ce1385a1cd46fb500d9a62de6883}{Project\_\-name} \mbox{[}25\mbox{]}}
\label{classInitiation_d011ce1385a1cd46fb500d9a62de6883}

\begin{CompactList}\small\item\em the project name \item\end{CompactList}\item 
int \hyperlink{classInitiation_ac12d75613f51d4a10a3f28887be0821}{number\_\-of\_\-materials}
\begin{CompactList}\small\item\em number of materials \item\end{CompactList}\item 
\hypertarget{classInitiation_3f9f020a489c9179885c02c867e77ef0}{
char \hyperlink{classInitiation_3f9f020a489c9179885c02c867e77ef0}{inputfile} \mbox{[}25\mbox{]}}
\label{classInitiation_3f9f020a489c9179885c02c867e77ef0}

\begin{CompactList}\small\item\em the global inputfile name: a $\ast$.cfg file \item\end{CompactList}\item 
int \hyperlink{classInitiation_1bbced7f2136569c065e0f5c874c6cff}{initial\_\-condition}
\begin{CompactList}\small\item\em initial condition marker: \item\end{CompactList}\item 
\hypertarget{classInitiation_b9979da1d25141fa58f79393232b0a9c}{
int \hyperlink{classInitiation_b9979da1d25141fa58f79393232b0a9c}{diagnose}}
\label{classInitiation_b9979da1d25141fa58f79393232b0a9c}

\begin{CompactList}\small\item\em diagnose information maker: 1 output diagnose information \item\end{CompactList}\item 
\hypertarget{classInitiation_115596cc83b34b6f75031c5e5b0c3fe7}{
double \hyperlink{classInitiation_115596cc83b34b6f75031c5e5b0c3fe7}{art\_\-vis}}
\label{classInitiation_115596cc83b34b6f75031c5e5b0c3fe7}

\begin{CompactList}\small\item\em artificial viscosity \item\end{CompactList}\item 
\hypertarget{classInitiation_db9ebb73d2d6e17bb8e0179d29abaf7b}{
int \hyperlink{classInitiation_db9ebb73d2d6e17bb8e0179d29abaf7b}{simu\_\-mode}}
\label{classInitiation_db9ebb73d2d6e17bb8e0179d29abaf7b}

\begin{CompactList}\small\item\em simulation mode (1:liquids, 2: gas dynamics) \item\end{CompactList}\item 
\hypertarget{classInitiation_7053a11a810568d6beb69ed19bc7fa59}{
double \hyperlink{classInitiation_7053a11a810568d6beb69ed19bc7fa59}{smoothinglength}}
\label{classInitiation_7053a11a810568d6beb69ed19bc7fa59}

\begin{CompactList}\small\item\em smoothinglength \item\end{CompactList}\item 
\hypertarget{classInitiation_c4ab38f327f64513f4c21cd5781f508a}{
Vec2d \hyperlink{classInitiation_c4ab38f327f64513f4c21cd5781f508a}{box\_\-size}}
\label{classInitiation_c4ab38f327f64513f4c21cd5781f508a}

\begin{CompactList}\small\item\em the compuational domain size \item\end{CompactList}\item 
\hypertarget{classInitiation_ed9ecf2c5296af4ac243b1a089966c86}{
double \hyperlink{classInitiation_ed9ecf2c5296af4ac243b1a089966c86}{cell\_\-size}}
\label{classInitiation_ed9ecf2c5296af4ac243b1a089966c86}

\begin{CompactList}\small\item\em cell size \item\end{CompactList}\item 
\hypertarget{classInitiation_300dc1f911e95df0ecf34179bf1931fa}{
double \hyperlink{classInitiation_300dc1f911e95df0ecf34179bf1931fa}{delta}}
\label{classInitiation_300dc1f911e95df0ecf34179bf1931fa}

\begin{CompactList}\small\item\em the inital particle distance \item\end{CompactList}\item 
\hypertarget{classInitiation_c0d2d3c34bb4831258ce646b05a35981}{
int \hyperlink{classInitiation_c0d2d3c34bb4831258ce646b05a35981}{hdelta}}
\label{classInitiation_c0d2d3c34bb4831258ce646b05a35981}

\begin{CompactList}\small\item\em the ration between smoothing length and inital particle distance \item\end{CompactList}\item 
\hypertarget{classInitiation_8bc245d0b81987fc9fb8443713ebe6df}{
int \hyperlink{classInitiation_8bc245d0b81987fc9fb8443713ebe6df}{x\_\-cells}}
\label{classInitiation_8bc245d0b81987fc9fb8443713ebe6df}

\begin{CompactList}\small\item\em cells matrix for real particles \item\end{CompactList}\item 
\hypertarget{classInitiation_4741b9c5dfa95aadd6b103f091456f72}{
int \textbf{y\_\-cells}}
\label{classInitiation_4741b9c5dfa95aadd6b103f091456f72}

\item 
\hypertarget{classInitiation_cf2bd144db5215805966ccd40a5abaed}{
Vec2d \hyperlink{classInitiation_cf2bd144db5215805966ccd40a5abaed}{g\_\-force}}
\label{classInitiation_cf2bd144db5215805966ccd40a5abaed}

\begin{CompactList}\small\item\em g force on particles \item\end{CompactList}\item 
\hypertarget{classInitiation_802c22e6b71f5d2cfe68309685bf423b}{
double \hyperlink{classInitiation_802c22e6b71f5d2cfe68309685bf423b}{Start\_\-time}}
\label{classInitiation_802c22e6b71f5d2cfe68309685bf423b}

\begin{CompactList}\small\item\em Simulation start time. \item\end{CompactList}\item 
\hypertarget{classInitiation_0d3e0f13a25c1bbbc7b319ef59838012}{
double \hyperlink{classInitiation_0d3e0f13a25c1bbbc7b319ef59838012}{End\_\-time}}
\label{classInitiation_0d3e0f13a25c1bbbc7b319ef59838012}

\begin{CompactList}\small\item\em Simulation end time. \item\end{CompactList}\item 
\hypertarget{classInitiation_e6599efb2ef3de1c552d39ba6ce5fe5e}{
double \hyperlink{classInitiation_e6599efb2ef3de1c552d39ba6ce5fe5e}{D\_\-time}}
\label{classInitiation_e6599efb2ef3de1c552d39ba6ce5fe5e}

\begin{CompactList}\small\item\em time interval for output (every D\_\-time: output) \item\end{CompactList}\item 
\hypertarget{classInitiation_8f3749288efc593efb8c6b2e5da2dce2}{
Vec2d \hyperlink{classInitiation_8f3749288efc593efb8c6b2e5da2dce2}{U0}}
\label{classInitiation_8f3749288efc593efb8c6b2e5da2dce2}

\begin{CompactList}\small\item\em inital flow speed (if initial condition is defined here)w \item\end{CompactList}\item 
\hypertarget{classInitiation_a426f5569c652979800af92d6d4dd033}{
double \hyperlink{classInitiation_a426f5569c652979800af92d6d4dd033}{rho0}}
\label{classInitiation_a426f5569c652979800af92d6d4dd033}

\begin{CompactList}\small\item\em initial particle density(if initial condition is defined here) \item\end{CompactList}\item 
\hypertarget{classInitiation_d108934b865ac8267f61d1e04339fea0}{
double \hyperlink{classInitiation_d108934b865ac8267f61d1e04339fea0}{p0}}
\label{classInitiation_d108934b865ac8267f61d1e04339fea0}

\begin{CompactList}\small\item\em initial pressure(if initial condition is defined here) \item\end{CompactList}\item 
\hypertarget{classInitiation_b8db429b0356e54c88d8769191e6f550}{
double \hyperlink{classInitiation_b8db429b0356e54c88d8769191e6f550}{T0}}
\label{classInitiation_b8db429b0356e54c88d8769191e6f550}

\begin{CompactList}\small\item\em initial temperature(if initial condition is defined here) \item\end{CompactList}\item 
\hypertarget{classInitiation_4a7f19654a3c5fc2c703796150ab7de5}{
int \hyperlink{classInitiation_4a7f19654a3c5fc2c703796150ab7de5}{MLS\_\-MAX}}
\label{classInitiation_4a7f19654a3c5fc2c703796150ab7de5}

\begin{CompactList}\small\item\em Moving Least Squarr. \item\end{CompactList}\end{CompactItemize}
\subsection*{Private Attributes}
\begin{CompactItemize}
\item 
\hypertarget{classInitiation_18480401f4b40c2830a764cc56085167}{
double \hyperlink{classInitiation_18480401f4b40c2830a764cc56085167}{\_\-length}}
\label{classInitiation_18480401f4b40c2830a764cc56085167}

\begin{CompactList}\small\item\em reference length for non dimensional value \item\end{CompactList}\item 
\hypertarget{classInitiation_0e7baf9e9f15d03c957db7bc1df9e5f0}{
double \hyperlink{classInitiation_0e7baf9e9f15d03c957db7bc1df9e5f0}{\_\-v}}
\label{classInitiation_0e7baf9e9f15d03c957db7bc1df9e5f0}

\begin{CompactList}\small\item\em reference speed for non dimensional value \item\end{CompactList}\item 
\hypertarget{classInitiation_a20ec3ff1ccfd3aaf9505a86c2d629a1}{
double \hyperlink{classInitiation_a20ec3ff1ccfd3aaf9505a86c2d629a1}{\_\-rho}}
\label{classInitiation_a20ec3ff1ccfd3aaf9505a86c2d629a1}

\begin{CompactList}\small\item\em reference density for non dimensional value \item\end{CompactList}\item 
\hypertarget{classInitiation_8dc68dc47e54d0e3bf9ab4a222513c51}{
double \hyperlink{classInitiation_8dc68dc47e54d0e3bf9ab4a222513c51}{\_\-T}}
\label{classInitiation_8dc68dc47e54d0e3bf9ab4a222513c51}

\begin{CompactList}\small\item\em reference temperature for non dimensional value \item\end{CompactList}\end{CompactItemize}
\subsection*{Friends}
\begin{CompactItemize}
\item 
\hypertarget{classInitiation_a1212b6e372a0f45d2c01f3cd203af77}{
class \hyperlink{classInitiation_a1212b6e372a0f45d2c01f3cd203af77}{Material}}
\label{classInitiation_a1212b6e372a0f45d2c01f3cd203af77}

\item 
\hypertarget{classInitiation_652e6d35f07e2091db6c1d01e86e2286}{
class \hyperlink{classInitiation_652e6d35f07e2091db6c1d01e86e2286}{Hydrodynamics}}
\label{classInitiation_652e6d35f07e2091db6c1d01e86e2286}

\end{CompactItemize}

\subsection{Detailed Description}
Initiates the simulation.

\subsection{Constructor \& Destructor Documentation}
\hypertarget{classInitiation_4db3b8c45d38cf4e188de6aa672e8c0d}{
\index{Initiation@{Initiation}!Initiation@{Initiation}}
\index{Initiation@{Initiation}!Initiation@{Initiation}}
\subsubsection[{Initiation}]{\setlength{\rightskip}{0pt plus 5cm}Initiation::Initiation (const char $\ast$ {\em project\_\-name})\hspace{0.3cm}{\tt  \mbox{[}explicit\mbox{]}}}}
\label{classInitiation_4db3b8c45d38cf4e188de6aa672e8c0d}

constructor \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em $\ast$project\_\-name,:}]a pointer to the project name array (the project name is defined as a main-function argument and therefore given as an additional command line element when running the progam) \end{description}
\end{Desc}

\begin{itemize}
\item check if inputfile exists (if not exit the program)

\item reading key words and configuration data from configuration file and assign them to the appropriate variable

\item create outdata directory

\item process the data {\bf !!!Question!!!}

(line 104) this is only true if h=cell\_\-size (which is not necessarily given, as h, cell\_\-size can be initiated independently in configuration file)

\item output information to screen

\item non-dimensionalize \par
\par
 \end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=255pt]{classInitiation_4db3b8c45d38cf4e188de6aa672e8c0d_cgraph}
\end{center}
\end{figure}

\subsection{Member Function Documentation}
\hypertarget{classInitiation_916e5b3e316dad1f524a1327b9e4aeb4}{
\index{Initiation@{Initiation}!non\_\-dimensionalize@{non\_\-dimensionalize}}
\index{non\_\-dimensionalize@{non\_\-dimensionalize}!Initiation@{Initiation}}
\subsubsection[{non\_\-dimensionalize}]{\setlength{\rightskip}{0pt plus 5cm}void Initiation::non\_\-dimensionalize ()}}
\label{classInitiation_916e5b3e316dad1f524a1327b9e4aeb4}

non-dimensionalize initial condition and parameters

method calls the individual non\_\-dms\_\-methods vor each variable\par
 remark:to avoid confusion: the non dimensional variables have the same identifiers like the dimensional ones!!!\par
\par

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=190pt]{classInitiation_916e5b3e316dad1f524a1327b9e4aeb4_cgraph}
\end{center}
\end{figure}

Here is the caller graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=159pt]{classInitiation_916e5b3e316dad1f524a1327b9e4aeb4_icgraph}
\end{center}
\end{figure}
\hypertarget{classInitiation_cdbf1374b9b0f0b50a0bb1d9abf76293}{
\index{Initiation@{Initiation}!show\_\-information@{show\_\-information}}
\index{show\_\-information@{show\_\-information}!Initiation@{Initiation}}
\subsubsection[{show\_\-information}]{\setlength{\rightskip}{0pt plus 5cm}void Initiation::show\_\-information ()}}
\label{classInitiation_cdbf1374b9b0f0b50a0bb1d9abf76293}

show information on screen

\begin{itemize}
\item output general information on screen\end{itemize}

\begin{itemize}
\item output the timing information on screen\end{itemize}

\begin{itemize}
\item output iniformation on initialization mode (.cfg file or .rst file) \end{itemize}

Here is the caller graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=154pt]{classInitiation_cdbf1374b9b0f0b50a0bb1d9abf76293_icgraph}
\end{center}
\end{figure}
\hypertarget{classInitiation_e80adc7d2105ad67b766fd1127966df3}{
\index{Initiation@{Initiation}!VolumeMass@{VolumeMass}}
\index{VolumeMass@{VolumeMass}!Initiation@{Initiation}}
\subsubsection[{VolumeMass}]{\setlength{\rightskip}{0pt plus 5cm}void Initiation::VolumeMass ({\bf Hydrodynamics} \& {\em hydro}, \/  {\bf ParticleManager} \& {\em particles}, \/  {\bf QuinticSpline} \& {\em weight\_\-function})}}
\label{classInitiation_e80adc7d2105ad67b766fd1127966df3}

predict the particle volume and mass

\begin{itemize}
\item iterate particles on the particle list

\begin{itemize}
\item pick an origin particle

\item build the nearest neighbor particle list for chosen origin particle

\item iterate this Nearest Neighbor \hyperlink{classParticle}{Particle} list

\begin{itemize}
\item get a particle

\item calculate distance (origin$<$-$>$ neighbor)

\item calculate weight function for given distance (w=0, if dist$>$supportlengtg) an summ it up \end{itemize}

\item calculate volume as reciprocal value of weight function

\item save volume and mass in the respective particle list node (whih is each a \hyperlink{classParticle}{Particle} object with all the particle properties)

\item clear the NNP\_\-list\end{itemize}
\end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=227pt]{classInitiation_e80adc7d2105ad67b766fd1127966df3_cgraph}
\end{center}
\end{figure}

Here is the caller graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=117pt]{classInitiation_e80adc7d2105ad67b766fd1127966df3_icgraph}
\end{center}
\end{figure}

\subsection{Member Data Documentation}
\hypertarget{classInitiation_1bbced7f2136569c065e0f5c874c6cff}{
\index{Initiation@{Initiation}!initial\_\-condition@{initial\_\-condition}}
\index{initial\_\-condition@{initial\_\-condition}!Initiation@{Initiation}}
\subsubsection[{initial\_\-condition}]{\setlength{\rightskip}{0pt plus 5cm}int {\bf Initiation::initial\_\-condition}}}
\label{classInitiation_1bbced7f2136569c065e0f5c874c6cff}

initial condition marker:

\begin{itemize}
\item 0 initialize from the .cfg file;\item 1 read from the .rst file particle by particle with non-dimensional data \end{itemize}
\hypertarget{classInitiation_ac12d75613f51d4a10a3f28887be0821}{
\index{Initiation@{Initiation}!number\_\-of\_\-materials@{number\_\-of\_\-materials}}
\index{number\_\-of\_\-materials@{number\_\-of\_\-materials}!Initiation@{Initiation}}
\subsubsection[{number\_\-of\_\-materials}]{\setlength{\rightskip}{0pt plus 5cm}int {\bf Initiation::number\_\-of\_\-materials}}}
\label{classInitiation_ac12d75613f51d4a10a3f28887be0821}

number of materials

\begin{itemize}
\item the no. zero material is always the wall\item therefore at least 2 materials should be included \end{itemize}

The documentation for this class was generated from the following files:\begin{CompactItemize}
\item 
\hyperlink{initiation_8h}{initiation.h}\item 
\hyperlink{initiation_8cpp}{initiation.cpp}\end{CompactItemize}