### 2010-06-17 src/latex/classKernel.tex

\hypertarget{classKernel}{
\section{Kernel Class Reference}
\label{classKernel}\index{Kernel@{Kernel}}
}
\hyperlink{classKernel}{Kernel} abstract base.

{\tt \#include $<$kernel.h$>$}

Inheritance diagram for Kernel:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=198pt]{classKernel__inherit__graph}
\end{center}
\end{figure}
\subsection*{Public Member Functions}
\begin{CompactItemize}
\item 
\hypertarget{classKernel_7ff93426a5df552fdb40f0600dbfe314}{
\hyperlink{classKernel_7ff93426a5df552fdb40f0600dbfe314}{Kernel} (double \hyperlink{classKernel_f088821dba4a00d1d320b27ebcb71258}{smoothingLength})}
\label{classKernel_7ff93426a5df552fdb40f0600dbfe314}

\begin{CompactList}\small\item\em Constructor to initialize data members. \item\end{CompactList}\item 
virtual \hyperlink{classKernel_9c7407e3a0b1cb9b2f96e9a030187064}{$\sim$Kernel} ()
\begin{CompactList}\small\item\em Base classes with virtual member functions should have a virtual destructor. \item\end{CompactList}\item 
\hypertarget{classKernel_6b4d26ba99457a4acbbbca1454dbdc8a}{
virtual double \hyperlink{classKernel_6b4d26ba99457a4acbbbca1454dbdc8a}{w} (double distance) const =0}
\label{classKernel_6b4d26ba99457a4acbbbca1454dbdc8a}

\begin{CompactList}\small\item\em Calculates the kernel value for the given distance of two particles. \item\end{CompactList}\item 
\hypertarget{classKernel_545e61b98db05db6dd02b88e2ad3da23}{
virtual Vec2d \hyperlink{classKernel_545e61b98db05db6dd02b88e2ad3da23}{gradW} (double distance, const Vec2d \&distanceVector) const =0}
\label{classKernel_545e61b98db05db6dd02b88e2ad3da23}

\begin{CompactList}\small\item\em Calculates the kernel derivation for the given distance of two particles. \item\end{CompactList}\end{CompactItemize}
\subsection*{Protected Attributes}
\begin{CompactItemize}
\item 
\hypertarget{classKernel_f088821dba4a00d1d320b27ebcb71258}{
const double \hyperlink{classKernel_f088821dba4a00d1d320b27ebcb71258}{smoothingLength}}
\label{classKernel_f088821dba4a00d1d320b27ebcb71258}

\begin{CompactList}\small\item\em A local copy of the smoothing length. \item\end{CompactList}\end{CompactItemize}

\subsection{Detailed Description}
\hyperlink{classKernel}{Kernel} abstract base.

\subsection{Constructor \& Destructor Documentation}
\hypertarget{classKernel_9c7407e3a0b1cb9b2f96e9a030187064}{
\index{Kernel@{Kernel}!$\sim$Kernel@{$\sim$Kernel}}
\index{$\sim$Kernel@{$\sim$Kernel}!Kernel@{Kernel}}
\subsubsection[{$\sim$Kernel}]{\setlength{\rightskip}{0pt plus 5cm}Kernel::$\sim$Kernel ()\hspace{0.3cm}{\tt  \mbox{[}virtual\mbox{]}}}}
\label{classKernel_9c7407e3a0b1cb9b2f96e9a030187064}

Base classes with virtual member functions should have a virtual destructor.

deconstructor

The documentation for this class was generated from the following files:\begin{CompactItemize}
\item 
\hyperlink{kernel_8h}{kernel.h}\item 
\hyperlink{kernel_8cpp}{kernel.cpp}\end{CompactItemize}