### 2010-06-17 src/latex/classLlistNode.tex

\hypertarget{classLlistNode}{
\section{LlistNode$<$ Ldata $>$ Class Template Reference}
\label{classLlistNode}\index{LlistNode@{LlistNode}}
}
A template node in the linked list.

{\tt \#include $<$dllist.h$>$}

Collaboration diagram for LlistNode$<$ Ldata $>$:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=142pt]{classLlistNode__coll__graph}
\end{center}
\end{figure}
\subsection*{Private Member Functions}
\begin{CompactItemize}
\item 
\hyperlink{classLlistNode_9e537419faf945c1be64fb8a994773f3}{LlistNode} ()
\begin{CompactList}\small\item\em constructors \item\end{CompactList}\item 
\hyperlink{classLlistNode_0257863d0b35acbdac9e297cadf90a04}{LlistNode} (Ldata $\ast$d, \hyperlink{classLlistNode}{LlistNode}$<$ Ldata $>$ $\ast$n)
\begin{CompactList}\small\item\em Creates a new node and inserts it before n. \item\end{CompactList}\end{CompactItemize}
\subsection*{Private Attributes}
\begin{CompactItemize}
\item 
\hypertarget{classLlistNode_f6147c63bdee0e7d19323ee2c8390755}{
Ldata $\ast$ \hyperlink{classLlistNode_f6147c63bdee0e7d19323ee2c8390755}{data}}
\label{classLlistNode_f6147c63bdee0e7d19323ee2c8390755}

\begin{CompactList}\small\item\em The data of this node. \item\end{CompactList}\item 
\hypertarget{classLlistNode_06cabe26723b810e801708ef9660f556}{
\hyperlink{classLlistNode}{LlistNode}$<$ Ldata $>$ $\ast$ \hyperlink{classLlistNode_06cabe26723b810e801708ef9660f556}{next}}
\label{classLlistNode_06cabe26723b810e801708ef9660f556}

\begin{CompactList}\small\item\em The next and the previous nodes. \item\end{CompactList}\item 
\hypertarget{classLlistNode_445540ddfb7fe610b13ebe59312b8545}{
\hyperlink{classLlistNode}{LlistNode}$<$ Ldata $>$ $\ast$ \textbf{prev}}
\label{classLlistNode_445540ddfb7fe610b13ebe59312b8545}

\end{CompactItemize}
\subsection*{Friends}
\begin{CompactItemize}
\item 
\hypertarget{classLlistNode_59c8480ff68562e8cfa2bc1285d79b77}{
class \hyperlink{classLlistNode_59c8480ff68562e8cfa2bc1285d79b77}{Llist$<$ Ldata $>$}}
\label{classLlistNode_59c8480ff68562e8cfa2bc1285d79b77}

\begin{CompactList}\small\item\em the list class which manipulates the nodes \item\end{CompactList}\end{CompactItemize}

\subsection{Detailed Description}
\subsubsection*{template$<$class Ldata$>$ class LlistNode$<$ Ldata $>$}

A template node in the linked list.

----------------------------------------------------------------------- a template node on the list ----------------------------------------------------------------------- This class represents a node of the generic double linked list \hyperlink{classLlist}{Llist}

\subsection{Constructor \& Destructor Documentation}
\hypertarget{classLlistNode_9e537419faf945c1be64fb8a994773f3}{
\index{LlistNode@{LlistNode}!LlistNode@{LlistNode}}
\index{LlistNode@{LlistNode}!LlistNode@{LlistNode}}
\subsubsection[{LlistNode}]{\setlength{\rightskip}{0pt plus 5cm}template$<$class Ldata$>$ {\bf LlistNode}$<$ Ldata $>$::{\bf LlistNode} ()\hspace{0.3cm}{\tt  \mbox{[}inline, private\mbox{]}}}}
\label{classLlistNode_9e537419faf945c1be64fb8a994773f3}

constructors

An empty node, used for the head node \hypertarget{classLlistNode_0257863d0b35acbdac9e297cadf90a04}{
\index{LlistNode@{LlistNode}!LlistNode@{LlistNode}}
\index{LlistNode@{LlistNode}!LlistNode@{LlistNode}}
\subsubsection[{LlistNode}]{\setlength{\rightskip}{0pt plus 5cm}template$<$class Ldata$>$ {\bf LlistNode}$<$ Ldata $>$::{\bf LlistNode} (Ldata $\ast$ {\em d}, \/  {\bf LlistNode}$<$ Ldata $>$ $\ast$ {\em n})\hspace{0.3cm}{\tt  \mbox{[}inline, private\mbox{]}}}}
\label{classLlistNode_0257863d0b35acbdac9e297cadf90a04}

Creates a new node and inserts it before n.

\begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em d}]The data for the new node \item[{\em n}]The node before which the new node is to be inserted \end{description}
\end{Desc}

The documentation for this class was generated from the following file:\begin{CompactItemize}
\item 
\hyperlink{dllist_8h}{dllist.h}\end{CompactItemize}