### 2010-06-17 src/latex/classMLS.tex

\hypertarget{classMLS}{
\section{MLS Class Reference}
\label{classMLS}\index{MLS@{MLS}}
}
Moving Least Squares (\hyperlink{classMLS}{MLS}) class.

{\tt \#include $<$mls.h$>$}

\subsection*{Public Member Functions}
\begin{CompactItemize}
\item 
\hyperlink{classMLS_69c81b7b81631bb306259deaf925cac3}{MLS} (\hyperlink{classInitiation}{Initiation} \&ini)
\begin{CompactList}\small\item\em constructor \item\end{CompactList}\item 
void \hyperlink{classMLS_70f00e273459e8fe02baa5a0c06d6f24}{MLSMapping} (Vec2d \&point, \hyperlink{classLlist}{Llist}$<$ \hyperlink{classParticle}{Particle} $>$ \&NNP\_\-list, \hyperlink{classQuinticSpline}{QuinticSpline} \&weight\_\-function, int order)
\begin{CompactList}\small\item\em \hyperlink{classMLS}{MLS} Mapping to grid points. \item\end{CompactList}\end{CompactItemize}
\subsection*{Public Attributes}
\begin{CompactItemize}
\item 
\hypertarget{classMLS_8e02a81746ff6002015a7325b871282c}{
double $\ast$ \hyperlink{classMLS_8e02a81746ff6002015a7325b871282c}{phi}}
\label{classMLS_8e02a81746ff6002015a7325b871282c}

\begin{CompactList}\small\item\em shape function (pointer to) \item\end{CompactList}\item 
\hypertarget{classMLS_328f8049ce8b237bdc42f89a8b5dde07}{
Vec2d $\ast$ \hyperlink{classMLS_328f8049ce8b237bdc42f89a8b5dde07}{dphi}}
\label{classMLS_328f8049ce8b237bdc42f89a8b5dde07}

\begin{CompactList}\small\item\em derivatives (pointer to): have not been implemented! \item\end{CompactList}\end{CompactItemize}
\subsection*{Private Member Functions}
\begin{CompactItemize}
\item 
\hypertarget{classMLS_7c7baa105805cd17714d35e351283bf7}{
void \hyperlink{classMLS_7c7baa105805cd17714d35e351283bf7}{MLS\_\-Solver} (int order)}
\label{classMLS_7c7baa105805cd17714d35e351283bf7}

\begin{CompactList}\small\item\em calculate the shape function and its derivative \item\end{CompactList}\end{CompactItemize}
\subsection*{Private Attributes}
\begin{CompactItemize}
\item 
\hypertarget{classMLS_2a4087a4494f44c3876144093a355b5f}{
int \hyperlink{classMLS_2a4087a4494f44c3876144093a355b5f}{N}}
\label{classMLS_2a4087a4494f44c3876144093a355b5f}

\begin{CompactList}\small\item\em numbers of points \item\end{CompactList}\item 
\hypertarget{classMLS_b610a9eb7b7df89432454f08ad98c9b0}{
int \hyperlink{classMLS_b610a9eb7b7df89432454f08ad98c9b0}{MLS\_\-MAX}}
\label{classMLS_b610a9eb7b7df89432454f08ad98c9b0}

\begin{CompactList}\small\item\em maximum number of reference points \item\end{CompactList}\item 
\hypertarget{classMLS_9499afde579c88688a27973bde8b4183}{
double \textbf{A} \mbox{[}3\mbox{]}\mbox{[}3\mbox{]}}
\label{classMLS_9499afde579c88688a27973bde8b4183}

\item 
\hypertarget{classMLS_ecdcd478c6683a0a865ef4db7385d77a}{
double $\ast$$\ast$ \hyperlink{classMLS_ecdcd478c6683a0a865ef4db7385d77a}{B}}
\label{classMLS_ecdcd478c6683a0a865ef4db7385d77a}

\begin{CompactList}\small\item\em weighted moment matrix (pointer to pointer) \item\end{CompactList}\item 
\hypertarget{classMLS_871ec34ef81919130fa552b982324a40}{
double \textbf{p} \mbox{[}3\mbox{]}}
\label{classMLS_871ec34ef81919130fa552b982324a40}

\item 
\hypertarget{classMLS_71607b0dededb152f2c75a767b84ffb8}{
double $\ast$$\ast$ \hyperlink{classMLS_71607b0dededb152f2c75a767b84ffb8}{pi}}
\label{classMLS_71607b0dededb152f2c75a767b84ffb8}

\begin{CompactList}\small\item\em ploynomial basis matrix (pointer to pointer) \item\end{CompactList}\item 
\hypertarget{classMLS_78d2e05f618cac0c7a8ca0ddef8c21c8}{
double $\ast$ \hyperlink{classMLS_78d2e05f618cac0c7a8ca0ddef8c21c8}{Wi}}
\label{classMLS_78d2e05f618cac0c7a8ca0ddef8c21c8}

\begin{CompactList}\small\item\em weights (pointer to) \item\end{CompactList}\end{CompactItemize}

\subsection{Detailed Description}
Moving Least Squares (\hyperlink{classMLS}{MLS}) class.

Using Moving Least Squares Approximation 2D and linear aproximation based on (Liu GR, Mesh Free Methods-$>$Moving beyond the finite element method, CRC press, London, 2003, p79-87)

\subsection{Constructor \& Destructor Documentation}
\hypertarget{classMLS_69c81b7b81631bb306259deaf925cac3}{
\index{MLS@{MLS}!MLS@{MLS}}
\index{MLS@{MLS}!MLS@{MLS}}
\subsubsection[{MLS}]{\setlength{\rightskip}{0pt plus 5cm}MLS::MLS ({\bf Initiation} \& {\em ini})}}
\label{classMLS_69c81b7b81631bb306259deaf925cac3}

constructor

\begin{itemize}
\item declares the weighted moment matrix and assigns it to corresponding pointer\end{itemize}

\begin{itemize}
\item declares the ploynomial basis matrix and assigns it to corresponding pointer\end{itemize}

\begin{itemize}
\item declares the weights and assigns it to corresponding pointer\end{itemize}

\begin{itemize}
\item declares shape function matrix and assigns it to corresponding pointer\end{itemize}

\begin{itemize}
\item declares derivatives matrix (HAS NOT BEEN IMPLEMENTED!!!) and assigns it to corresponding pointer \end{itemize}

\subsection{Member Function Documentation}
\hypertarget{classMLS_70f00e273459e8fe02baa5a0c06d6f24}{
\index{MLS@{MLS}!MLSMapping@{MLSMapping}}
\index{MLSMapping@{MLSMapping}!MLS@{MLS}}
\subsubsection[{MLSMapping}]{\setlength{\rightskip}{0pt plus 5cm}void MLS::MLSMapping (Vec2d \& {\em point}, \/  {\bf Llist}$<$ {\bf Particle} $>$ \& {\em NNP\_\-list}, \/  {\bf QuinticSpline} \& {\em weight\_\-function}, \/  int {\em order})}}
\label{classMLS_70f00e273459e8fe02baa5a0c06d6f24}

\hyperlink{classMLS}{MLS} Mapping to grid points.

set up the enviroment for \hyperlink{classMLS}{MLS} calculation and run \hyperlink{classMLS}{MLS} solver

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=246pt]{classMLS_70f00e273459e8fe02baa5a0c06d6f24_cgraph}
\end{center}
\end{figure}

Here is the caller graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=284pt]{classMLS_70f00e273459e8fe02baa5a0c06d6f24_icgraph}
\end{center}
\end{figure}

The documentation for this class was generated from the following files:\begin{CompactItemize}
\item 
\hyperlink{mls_8h}{mls.h}\item 
\hyperlink{mls_8cpp}{mls.cpp}\end{CompactItemize}