### 2010-06-17 src/latex/classMaterial.tex

\hypertarget{classMaterial}{
\section{Material Class Reference}
\label{classMaterial}\index{Material@{Material}}
}
\hyperlink{classMaterial}{Material}.

{\tt \#include $<$material.h$>$}

\subsection*{Public Member Functions}
\begin{CompactItemize}
\item 
\hypertarget{classMaterial_137e987401b63eb7c6c27c3e38bc74b5}{
\hyperlink{classMaterial_137e987401b63eb7c6c27c3e38bc74b5}{Material} ()}
\label{classMaterial_137e987401b63eb7c6c27c3e38bc74b5}

\begin{CompactList}\small\item\em constructor \item\end{CompactList}\item 
\hypertarget{classMaterial_8232db4572229d207a6438cf37b0c548}{
\hyperlink{classMaterial_8232db4572229d207a6438cf37b0c548}{Material} (\hyperlink{classInitiation}{Initiation} \&ini)}
\label{classMaterial_8232db4572229d207a6438cf37b0c548}

\begin{CompactList}\small\item\em constructor \item\end{CompactList}\item 
\hyperlink{classMaterial_16e68aae88c0ecd98a88654e4648aa75}{Material} (char $\ast$\hyperlink{classMaterial_d7d10964971b23171efeac3c801371cf}{material\_\-name}, \hyperlink{classInitiation}{Initiation} \&ini)
\begin{CompactList}\small\item\em constructor \item\end{CompactList}\item 
\hypertarget{classMaterial_7319953e9448604807d6e9301eea6986}{
void \hyperlink{classMaterial_7319953e9448604807d6e9301eea6986}{non\_\-dimensionalize} (\hyperlink{classInitiation}{Initiation} \&ini)}
\label{classMaterial_7319953e9448604807d6e9301eea6986}

\begin{CompactList}\small\item\em non-dimensionalize \item\end{CompactList}\item 
\hypertarget{classMaterial_1c79d7b457fd73587612d34c687f4286}{
void \hyperlink{classMaterial_1c79d7b457fd73587612d34c687f4286}{show\_\-properties} ()}
\label{classMaterial_1c79d7b457fd73587612d34c687f4286}

\begin{CompactList}\small\item\em show properties \item\end{CompactList}\item 
\hypertarget{classMaterial_fb5f537d2e8476debf2228ce8db4e8cf}{
void \hyperlink{classMaterial_fb5f537d2e8476debf2228ce8db4e8cf}{Get\_\-b0} (double sound)}
\label{classMaterial_fb5f537d2e8476debf2228ce8db4e8cf}

\begin{CompactList}\small\item\em obtain parameter b0 \item\end{CompactList}\item 
\hypertarget{classMaterial_be327be4e21bff9a227d0152c0db3f97}{
double \hyperlink{classMaterial_be327be4e21bff9a227d0152c0db3f97}{get\_\-p} (double rho)}
\label{classMaterial_be327be4e21bff9a227d0152c0db3f97}

\begin{CompactList}\small\item\em get pressure \item\end{CompactList}\item 
\hypertarget{classMaterial_6d61301ba97502c9acb2987afc0a8f5e}{
double \hyperlink{classMaterial_6d61301ba97502c9acb2987afc0a8f5e}{get\_\-rho} (double p)}
\label{classMaterial_6d61301ba97502c9acb2987afc0a8f5e}

\begin{CompactList}\small\item\em get rho from pressure \item\end{CompactList}\item 
\hypertarget{classMaterial_e3c05673e9e43b1d16340b95ebaf64b0}{
double \hyperlink{classMaterial_e3c05673e9e43b1d16340b95ebaf64b0}{get\_\-e} (double T)}
\label{classMaterial_e3c05673e9e43b1d16340b95ebaf64b0}

\begin{CompactList}\small\item\em get interal energy \item\end{CompactList}\item 
\hypertarget{classMaterial_8f02587e88f4e307bd7952f0d47a6737}{
double \hyperlink{classMaterial_8f02587e88f4e307bd7952f0d47a6737}{get\_\-T} (double e)}
\label{classMaterial_8f02587e88f4e307bd7952f0d47a6737}

\begin{CompactList}\small\item\em get temperature (function is defined nowhere) \item\end{CompactList}\item 
\hypertarget{classMaterial_ad32f778be3001bd9d359beb8c071106}{
double \hyperlink{classMaterial_ad32f778be3001bd9d359beb8c071106}{get\_\-Cs} (double p, double rho)}
\label{classMaterial_ad32f778be3001bd9d359beb8c071106}

\begin{CompactList}\small\item\em get sound speed \item\end{CompactList}\end{CompactItemize}
\subsection*{Public Attributes}
\begin{CompactItemize}
\item 
\hypertarget{classMaterial_d7d10964971b23171efeac3c801371cf}{
char \hyperlink{classMaterial_d7d10964971b23171efeac3c801371cf}{material\_\-name} \mbox{[}25\mbox{]}}
\label{classMaterial_d7d10964971b23171efeac3c801371cf}

\begin{CompactList}\small\item\em material name string \item\end{CompactList}\item 
\hypertarget{classMaterial_e773333eded6b2af56382fb5c405310f}{
int \hyperlink{classMaterial_e773333eded6b2af56382fb5c405310f}{number}}
\label{classMaterial_e773333eded6b2af56382fb5c405310f}

\begin{CompactList}\small\item\em the material NO. \item\end{CompactList}\item 
int \hyperlink{classMaterial_4eaab2babdbe6e6338cf094bfefb6b2e}{material\_\-type}
\begin{CompactList}\small\item\em material type \item\end{CompactList}\item 
\hypertarget{classMaterial_994575fad2f67950eb32dad7fbd63e5e}{
double \hyperlink{classMaterial_994575fad2f67950eb32dad7fbd63e5e}{cv}}
\label{classMaterial_994575fad2f67950eb32dad7fbd63e5e}

\begin{CompactList}\small\item\em heat capacity \item\end{CompactList}\item 
\hypertarget{classMaterial_62e8368c76c975790315b16d97558c26}{
double \hyperlink{classMaterial_62e8368c76c975790315b16d97558c26}{eta}}
\label{classMaterial_62e8368c76c975790315b16d97558c26}

\begin{CompactList}\small\item\em viscosity \item\end{CompactList}\item 
\hypertarget{classMaterial_a9a05f29e651c0ca772a50f5dbbfda2a}{
double \hyperlink{classMaterial_a9a05f29e651c0ca772a50f5dbbfda2a}{zeta}}
\label{classMaterial_a9a05f29e651c0ca772a50f5dbbfda2a}

\begin{CompactList}\small\item\em heat conduction rate \item\end{CompactList}\item 
\hypertarget{classMaterial_2444df069ed1e68eea1aead044e47847}{
double \hyperlink{classMaterial_2444df069ed1e68eea1aead044e47847}{kappa}}
\label{classMaterial_2444df069ed1e68eea1aead044e47847}

\begin{CompactList}\small\item\em ??? \item\end{CompactList}\item 
\hypertarget{classMaterial_fcbe6b9bc1cffe3105bdfbf593834e37}{
double \hyperlink{classMaterial_fcbe6b9bc1cffe3105bdfbf593834e37}{gamma}}
\label{classMaterial_fcbe6b9bc1cffe3105bdfbf593834e37}

\begin{CompactList}\small\item\em ??? \item\end{CompactList}\item 
\hypertarget{classMaterial_3da7d0218fcd6369d0bdbb7f91280664}{
double \hyperlink{classMaterial_3da7d0218fcd6369d0bdbb7f91280664}{nu}}
\label{classMaterial_3da7d0218fcd6369d0bdbb7f91280664}

\begin{CompactList}\small\item\em kinetice viscosity \item\end{CompactList}\item 
\hypertarget{classMaterial_9ead67ba9a2882dabf61b6a5e501c8a9}{
double \hyperlink{classMaterial_9ead67ba9a2882dabf61b6a5e501c8a9}{b0}}
\label{classMaterial_9ead67ba9a2882dabf61b6a5e501c8a9}

\begin{CompactList}\small\item\em reference pressure \item\end{CompactList}\item 
\hypertarget{classMaterial_80c74d0a402e1150466fb3d350c81e61}{
double \hyperlink{classMaterial_80c74d0a402e1150466fb3d350c81e61}{rho0}}
\label{classMaterial_80c74d0a402e1150466fb3d350c81e61}

\begin{CompactList}\small\item\em reference density \item\end{CompactList}\item 
\hypertarget{classMaterial_c6109a2fded5aa9c518b2d5271fc7d76}{
double \hyperlink{classMaterial_c6109a2fded5aa9c518b2d5271fc7d76}{a0}}
\label{classMaterial_c6109a2fded5aa9c518b2d5271fc7d76}

\begin{CompactList}\small\item\em reference {\bf sound speed??? was not commented} \item\end{CompactList}\end{CompactItemize}
\subsection*{Static Private Attributes}
\begin{CompactItemize}
\item 
\hypertarget{classMaterial_d8ce370248325165c1c2505895cc9658}{
static int \hyperlink{classMaterial_d8ce370248325165c1c2505895cc9658}{number\_\-of\_\-materials} = 0}
\label{classMaterial_d8ce370248325165c1c2505895cc9658}

\begin{CompactList}\small\item\em total number of materials \item\end{CompactList}\item 
\hypertarget{classMaterial_623d1eecd8b673ffbfb5934df209c5a7}{
static double \hyperlink{classMaterial_623d1eecd8b673ffbfb5934df209c5a7}{smoothinglength} = 0.0}
\label{classMaterial_623d1eecd8b673ffbfb5934df209c5a7}

\begin{CompactList}\small\item\em smoothinglenth \item\end{CompactList}\item 
\hypertarget{classMaterial_31f7c32c96eae2d91b5988d1945e9c18}{
static double \textbf{delta} = 0.0}
\label{classMaterial_31f7c32c96eae2d91b5988d1945e9c18}

\end{CompactItemize}

\subsection{Detailed Description}
\hyperlink{classMaterial}{Material}.

\subsection{Constructor \& Destructor Documentation}
\hypertarget{classMaterial_16e68aae88c0ecd98a88654e4648aa75}{
\index{Material@{Material}!Material@{Material}}
\index{Material@{Material}!Material@{Material}}
\subsubsection[{Material}]{\setlength{\rightskip}{0pt plus 5cm}Material::Material (char $\ast$ {\em material\_\-name}, \/  {\bf Initiation} \& {\em ini})}}
\label{classMaterial_16e68aae88c0ecd98a88654e4648aa75}

constructor

-check if inputfile exist

-read material properties from .cfg file

\begin{itemize}
\item output the property parameters to the screen \end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=262pt]{classMaterial_16e68aae88c0ecd98a88654e4648aa75_cgraph}
\end{center}
\end{figure}

\subsection{Member Data Documentation}
\hypertarget{classMaterial_4eaab2babdbe6e6338cf094bfefb6b2e}{
\index{Material@{Material}!material\_\-type@{material\_\-type}}
\index{material\_\-type@{material\_\-type}!Material@{Material}}
\subsubsection[{material\_\-type}]{\setlength{\rightskip}{0pt plus 5cm}int {\bf Material::material\_\-type}}}
\label{classMaterial_4eaab2babdbe6e6338cf094bfefb6b2e}

material type

\begin{itemize}
\item 0: solid;\item 1: liquid;\item 2: gas \end{itemize}

The documentation for this class was generated from the following files:\begin{CompactItemize}
\item 
\hyperlink{material_8h}{material.h}\item 
\hyperlink{material_8cpp}{material.cpp}\end{CompactItemize}