### 2010-06-17 src/latex/classOutput.tex

\hypertarget{classOutput}{
\section{Output Class Reference}
\label{classOutput}\index{Output@{Output}}
}
\hyperlink{classOutput}{Output} class.

{\tt \#include $<$output.h$>$}

\subsection*{Public Member Functions}
\begin{CompactItemize}
\item 
\hyperlink{classOutput_2bd84635e13bedd1b8417fd39062a767}{Output} (\hyperlink{classInitiation}{Initiation} \&ini)
\begin{CompactList}\small\item\em constructor \item\end{CompactList}\item 
void \hyperlink{classOutput_f9e8b1a0aa456c01c69f7b3c7ee2649a}{OutputParticles} (\hyperlink{classHydrodynamics}{Hydrodynamics} \&hydro, \hyperlink{classBoundary}{Boundary} \&boundary, double Time, \hyperlink{classInitiation}{Initiation} \&ini)
\begin{CompactList}\small\item\em output particle positions respected different materials \item\end{CompactList}\item 
void \hyperlink{classOutput_2b41cacb61eecf10cbaca6bebb3e6316}{OutputStates} (\hyperlink{classParticleManager}{ParticleManager} \&particles, \hyperlink{classMLS}{MLS} \&mls, \hyperlink{classQuinticSpline}{QuinticSpline} \&weight\_\-function, double Time, \hyperlink{classInitiation}{Initiation} \&ini)
\begin{CompactList}\small\item\em output material states on uniform grid \item\end{CompactList}\item 
void \hyperlink{classOutput_a96aad43762590fc7779d24050c0a151}{OutRestart} (\hyperlink{classHydrodynamics}{Hydrodynamics} \&hydro, double Time, \hyperlink{classInitiation}{Initiation} \&ini)
\begin{CompactList}\small\item\em \hyperlink{classOutput}{Output} real particle data for restart the computation. \item\end{CompactList}\item 
\hypertarget{classOutput_16c4f8bae53fec90a812e5d479ae822d}{
void \hyperlink{classOutput_16c4f8bae53fec90a812e5d479ae822d}{CreatParticleMovie} ()}
\label{classOutput_16c4f8bae53fec90a812e5d479ae822d}

\begin{CompactList}\small\item\em create a head file for a movie of particle motion \item\end{CompactList}\item 
void \hyperlink{classOutput_3ff4c5c04d9f7e89eb3f36b299256f10}{WriteParticleMovie} (\hyperlink{classHydrodynamics}{Hydrodynamics} \&hydro, double Time, \hyperlink{classInitiation}{Initiation} \&ini)
\begin{CompactList}\small\item\em write the data for a movie of particle motion (output real and wall particles) \item\end{CompactList}\item 
void \hyperlink{classOutput_13b9be49d851806b4ab40d60624095ad}{OutAverage} (\hyperlink{classParticleManager}{ParticleManager} \&particles, \hyperlink{classMLS}{MLS} \&mls, \hyperlink{classQuinticSpline}{QuinticSpline} \&weight\_\-function, double Time, \hyperlink{classInitiation}{Initiation} \&ini)
\begin{CompactList}\small\item\em avergaed values for Poiseuille and Couette flows (average is taken in horizontal direction) \item\end{CompactList}\end{CompactItemize}
\subsection*{Private Attributes}
\begin{CompactItemize}
\item 
\hypertarget{classOutput_b52dd1cb165a9270cc45610b0bc8cc43}{
char \hyperlink{classOutput_b52dd1cb165a9270cc45610b0bc8cc43}{Project\_\-name} \mbox{[}25\mbox{]}}
\label{classOutput_b52dd1cb165a9270cc45610b0bc8cc43}

\begin{CompactList}\small\item\em the project name \item\end{CompactList}\item 
\hypertarget{classOutput_c2696f9c30e4840eab5b7ab3901bbf6b}{
int \textbf{number\_\-of\_\-materials}}
\label{classOutput_c2696f9c30e4840eab5b7ab3901bbf6b}

\item 
\hypertarget{classOutput_b3283271fd00c93268f9af19879a2fc6}{
double \hyperlink{classOutput_b3283271fd00c93268f9af19879a2fc6}{delta}}
\label{classOutput_b3283271fd00c93268f9af19879a2fc6}

\begin{CompactList}\small\item\em the inital particle distance \item\end{CompactList}\item 
\hypertarget{classOutput_8222be56a91be9ffd805ae2f87d68ec5}{
int \hyperlink{classOutput_8222be56a91be9ffd805ae2f87d68ec5}{x\_\-cells}}
\label{classOutput_8222be56a91be9ffd805ae2f87d68ec5}

\begin{CompactList}\small\item\em cells matrix for real particles \item\end{CompactList}\item 
\hypertarget{classOutput_d9bb2a18d020f12343b75eb5844fa79b}{
int \textbf{y\_\-cells}}
\label{classOutput_d9bb2a18d020f12343b75eb5844fa79b}

\item 
\hypertarget{classOutput_b2ba6673b6ca7b7ad7d952015c6dbfb7}{
int \hyperlink{classOutput_b2ba6673b6ca7b7ad7d952015c6dbfb7}{hdelta}}
\label{classOutput_b2ba6673b6ca7b7ad7d952015c6dbfb7}

\begin{CompactList}\small\item\em the ratio between smoothing length and inital particle distance \item\end{CompactList}\end{CompactItemize}

\subsection{Detailed Description}
\hyperlink{classOutput}{Output} class.

\subsection{Constructor \& Destructor Documentation}
\hypertarget{classOutput_2bd84635e13bedd1b8417fd39062a767}{
\index{Output@{Output}!Output@{Output}}
\index{Output@{Output}!Output@{Output}}
\subsubsection[{Output}]{\setlength{\rightskip}{0pt plus 5cm}Output::Output ({\bf Initiation} \& {\em ini})}}
\label{classOutput_2bd84635e13bedd1b8417fd39062a767}

constructor

copy boundary properties from initiation class

\subsection{Member Function Documentation}
\hypertarget{classOutput_13b9be49d851806b4ab40d60624095ad}{
\index{Output@{Output}!OutAverage@{OutAverage}}
\index{OutAverage@{OutAverage}!Output@{Output}}
\subsubsection[{OutAverage}]{\setlength{\rightskip}{0pt plus 5cm}void Output::OutAverage ({\bf ParticleManager} \& {\em particles}, \/  {\bf MLS} \& {\em mls}, \/  {\bf QuinticSpline} \& {\em weight\_\-function}, \/  double {\em Time}, \/  {\bf Initiation} \& {\em ini})}}
\label{classOutput_13b9be49d851806b4ab40d60624095ad}

avergaed values for Poiseuille and Couette flows (average is taken in horizontal direction)

\begin{itemize}
\item produce output file name\end{itemize}

\begin{itemize}
\item defining header for tecplot(plot software)\end{itemize}

\begin{itemize}
\item loop the grid points \end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=333pt]{classOutput_13b9be49d851806b4ab40d60624095ad_cgraph}
\end{center}
\end{figure}
\hypertarget{classOutput_f9e8b1a0aa456c01c69f7b3c7ee2649a}{
\index{Output@{Output}!OutputParticles@{OutputParticles}}
\index{OutputParticles@{OutputParticles}!Output@{Output}}
\subsubsection[{OutputParticles}]{\setlength{\rightskip}{0pt plus 5cm}void Output::OutputParticles ({\bf Hydrodynamics} \& {\em hydro}, \/  {\bf Boundary} \& {\em boundary}, \/  double {\em Time}, \/  {\bf Initiation} \& {\em ini})}}
\label{classOutput_f9e8b1a0aa456c01c69f7b3c7ee2649a}

output particle positions respected different materials

\begin{itemize}
\item produce output file name

\item defining header for tecplot(plot software)

\item output real and soild particles

\begin{itemize}
\item iterate the real partilce list

\item iterate the boundary partilce list \end{itemize}
\end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=157pt]{classOutput_f9e8b1a0aa456c01c69f7b3c7ee2649a_cgraph}
\end{center}
\end{figure}

Here is the caller graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=119pt]{classOutput_f9e8b1a0aa456c01c69f7b3c7ee2649a_icgraph}
\end{center}
\end{figure}
\hypertarget{classOutput_2b41cacb61eecf10cbaca6bebb3e6316}{
\index{Output@{Output}!OutputStates@{OutputStates}}
\index{OutputStates@{OutputStates}!Output@{Output}}
\subsubsection[{OutputStates}]{\setlength{\rightskip}{0pt plus 5cm}void Output::OutputStates ({\bf ParticleManager} \& {\em particles}, \/  {\bf MLS} \& {\em mls}, \/  {\bf QuinticSpline} \& {\em weight\_\-function}, \/  double {\em Time}, \/  {\bf Initiation} \& {\em ini})}}
\label{classOutput_2b41cacb61eecf10cbaca6bebb3e6316}

output material states on uniform grid

\begin{itemize}
\item produce output file name\end{itemize}

\begin{itemize}
\item defining header for tecplot(plot software)\end{itemize}

\begin{itemize}
\item loop the grid points \end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=337pt]{classOutput_2b41cacb61eecf10cbaca6bebb3e6316_cgraph}
\end{center}
\end{figure}

Here is the caller graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=114pt]{classOutput_2b41cacb61eecf10cbaca6bebb3e6316_icgraph}
\end{center}
\end{figure}
\hypertarget{classOutput_a96aad43762590fc7779d24050c0a151}{
\index{Output@{Output}!OutRestart@{OutRestart}}
\index{OutRestart@{OutRestart}!Output@{Output}}
\subsubsection[{OutRestart}]{\setlength{\rightskip}{0pt plus 5cm}void Output::OutRestart ({\bf Hydrodynamics} \& {\em hydro}, \/  double {\em Time}, \/  {\bf Initiation} \& {\em ini})}}
\label{classOutput_a96aad43762590fc7779d24050c0a151}

\hyperlink{classOutput}{Output} real particle data for restart the computation.

\begin{itemize}
\item output non-dimensional data\end{itemize}

\begin{itemize}
\item out reinitiation Time\end{itemize}

\begin{itemize}
\item output real particles (by iterating the particle list) \end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=126pt]{classOutput_a96aad43762590fc7779d24050c0a151_cgraph}
\end{center}
\end{figure}

Here is the caller graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=109pt]{classOutput_a96aad43762590fc7779d24050c0a151_icgraph}
\end{center}
\end{figure}
\hypertarget{classOutput_3ff4c5c04d9f7e89eb3f36b299256f10}{
\index{Output@{Output}!WriteParticleMovie@{WriteParticleMovie}}
\index{WriteParticleMovie@{WriteParticleMovie}!Output@{Output}}
\subsubsection[{WriteParticleMovie}]{\setlength{\rightskip}{0pt plus 5cm}void Output::WriteParticleMovie ({\bf Hydrodynamics} \& {\em hydro}, \/  double {\em Time}, \/  {\bf Initiation} \& {\em ini})}}
\label{classOutput_3ff4c5c04d9f7e89eb3f36b299256f10}

write the data for a movie of particle motion (output real and wall particles)

\begin{itemize}
\item produce output file name\end{itemize}

\begin{itemize}
\item iterate the partilce list \end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=160pt]{classOutput_3ff4c5c04d9f7e89eb3f36b299256f10_cgraph}
\end{center}
\end{figure}

Here is the caller graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=125pt]{classOutput_3ff4c5c04d9f7e89eb3f36b299256f10_icgraph}
\end{center}
\end{figure}

The documentation for this class was generated from the following files:\begin{CompactItemize}
\item 
\hyperlink{output_8h}{output.h}\item 
\hyperlink{output_8cpp}{output.cpp}\end{CompactItemize}