### 2010-06-17 src/latex/classParticleManager.tex

\hypertarget{classParticleManager}{
\section{ParticleManager Class Reference}
\label{classParticleManager}\index{ParticleManager@{ParticleManager}}
}
\hyperlink{classParticle}{Particle} manager class.

{\tt \#include $<$particlemanager.h$>$}

Collaboration diagram for ParticleManager:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=189pt]{classParticleManager__coll__graph}
\end{center}
\end{figure}
\subsection*{Public Member Functions}
\begin{CompactItemize}
\item 
\hypertarget{classParticleManager_36f2371b06ef7bd96fc1e0cb1fcebd66}{
\hyperlink{classParticleManager_36f2371b06ef7bd96fc1e0cb1fcebd66}{ParticleManager} ()}
\label{classParticleManager_36f2371b06ef7bd96fc1e0cb1fcebd66}

\begin{CompactList}\small\item\em \char`\"{}empty\char`\"{} constructor \item\end{CompactList}\item 
\hyperlink{classParticleManager_11a8f14c20f087713172d230de468166}{ParticleManager} (\hyperlink{classInitiation}{Initiation} \&ini)
\begin{CompactList}\small\item\em constructor \item\end{CompactList}\item 
\hypertarget{classParticleManager_7e835509b041635805a127cbedc194bd}{
\textbf{ParticleManager} (double cell\_\-size, int x\_\-cells, int y\_\-cells)}
\label{classParticleManager_7e835509b041635805a127cbedc194bd}

\item 
void \hyperlink{classParticleManager_6ee12c3925f371782800a50b17ac0941}{BiuldRealParticles} (\hyperlink{classHydrodynamics}{Hydrodynamics} \&hydro, \hyperlink{classInitiation}{Initiation} \&ini)
\begin{CompactList}\small\item\em buid the initial particles and the linked lists \item\end{CompactList}\item 
\hypertarget{classParticleManager_04ed35d52cf918b98006a0def713dc89}{
void \textbf{BiuldRealParticles} (\hyperlink{classHydrodynamics}{Hydrodynamics} \&hydro)}
\label{classParticleManager_04ed35d52cf918b98006a0def713dc89}

\item 
void \hyperlink{classParticleManager_740244439672f753735abc5a6a7cc3bc}{UpdateCellLinkedLists} ()
\begin{CompactList}\small\item\em update the cell linked lists for real particles \item\end{CompactList}\item 
void \hyperlink{classParticleManager_4e0f9e5cb222739a83f0f80ee4ba265b}{BuildNNP} (Vec2d \&point)
\begin{CompactList}\small\item\em do NNP search around a point and build the NNP list \item\end{CompactList}\item 
void \hyperlink{classParticleManager_b68ab55a9429ac8d32368b531487dcec}{BuildNNP\_\-MLSMapping} (Vec2d \&point)
\begin{CompactList}\small\item\em do NNP search around a point and build the NNP list for \hyperlink{classMLS}{MLS} approximation \item\end{CompactList}\item 
void \hyperlink{classParticleManager_8cc6658324694ad28d3c50bfc0cbd0f7}{BuildInteraction} (\hyperlink{classLlist}{Llist}$<$ \hyperlink{classInteraction}{Interaction} $>$ \&interactions, \hyperlink{classLlist}{Llist}$<$ \hyperlink{classParticle}{Particle} $>$ \&particle\_\-list, \hyperlink{classForce}{Force} $\ast$$\ast$forces, \hyperlink{classQuinticSpline}{QuinticSpline} \&weight\_\-function)
\begin{CompactList}\small\item\em build the interaction (particle pair) list \item\end{CompactList}\end{CompactItemize}
\subsection*{Public Attributes}
\begin{CompactItemize}
\item 
\hypertarget{classParticleManager_26e6109376eb2a1dc37f7fe623aaf528}{
int \hyperlink{classParticleManager_26e6109376eb2a1dc37f7fe623aaf528}{x\_\-clls}}
\label{classParticleManager_26e6109376eb2a1dc37f7fe623aaf528}

\begin{CompactList}\small\item\em linked cell matrix size x-direction \item\end{CompactList}\item 
\hypertarget{classParticleManager_09eae89d0d9e56d988e40cdebf13825b}{
int \hyperlink{classParticleManager_09eae89d0d9e56d988e40cdebf13825b}{y\_\-clls}}
\label{classParticleManager_09eae89d0d9e56d988e40cdebf13825b}

\begin{CompactList}\small\item\em linked cell matrix size y-direction \item\end{CompactList}\item 
\hypertarget{classParticleManager_0b7d522a4ecc14073c1adbb42c23fc84}{
\hyperlink{classLlist}{Llist}$<$ \hyperlink{classParticle}{Particle} $>$ $\ast$$\ast$ \hyperlink{classParticleManager_0b7d522a4ecc14073c1adbb42c23fc84}{cell\_\-lists}}
\label{classParticleManager_0b7d522a4ecc14073c1adbb42c23fc84}

\begin{CompactList}\small\item\em cell linked list in 2-d array \item\end{CompactList}\item 
\hypertarget{classParticleManager_8f47c372d7ea5ff68657defd7a492e85}{
\hyperlink{classLlist}{Llist}$<$ \hyperlink{classParticle}{Particle} $>$ \hyperlink{classParticleManager_8f47c372d7ea5ff68657defd7a492e85}{NNP\_\-list}}
\label{classParticleManager_8f47c372d7ea5ff68657defd7a492e85}

\begin{CompactList}\small\item\em list for the nearest neighbor particles \item\end{CompactList}\end{CompactItemize}
\subsection*{Private Member Functions}
\begin{CompactItemize}
\item 
void \hyperlink{classParticleManager_edebb3e37b06426f6509b35c0926bd59}{BiuldWallParticles} (\hyperlink{classHydrodynamics}{Hydrodynamics} \&hydro, \hyperlink{classInitiation}{Initiation} \&ini, \hyperlink{classBoundary}{Boundary} \&boundary)
\begin{CompactList}\small\item\em buid the initial wall particles and the linked lists \item\end{CompactList}\end{CompactItemize}
\subsection*{Private Attributes}
\begin{CompactItemize}
\item 
\hypertarget{classParticleManager_85dd6a84aa88a5e873ea6c23f65f9358}{
char \hyperlink{classParticleManager_85dd6a84aa88a5e873ea6c23f65f9358}{Project\_\-name} \mbox{[}25\mbox{]}}
\label{classParticleManager_85dd6a84aa88a5e873ea6c23f65f9358}

\begin{CompactList}\small\item\em the project name \item\end{CompactList}\item 
\hypertarget{classParticleManager_4585baa6023df69c39f277ee05cd21c8}{
int \textbf{number\_\-of\_\-materials}}
\label{classParticleManager_4585baa6023df69c39f277ee05cd21c8}

\item 
\hypertarget{classParticleManager_ed08b546912db6b921a172fe93dec0ab}{
double \hyperlink{classParticleManager_ed08b546912db6b921a172fe93dec0ab}{smoothinglength}}
\label{classParticleManager_ed08b546912db6b921a172fe93dec0ab}

\begin{CompactList}\small\item\em smoothinglenth \item\end{CompactList}\item 
\hypertarget{classParticleManager_43a0ada795094f76ae26db96c10e7955}{
double \hyperlink{classParticleManager_43a0ada795094f76ae26db96c10e7955}{smoothinglengthsquare}}
\label{classParticleManager_43a0ada795094f76ae26db96c10e7955}

\begin{CompactList}\small\item\em smoothinglenth \item\end{CompactList}\item 
\hypertarget{classParticleManager_585655101c829ec78e9abe6d849210fe}{
double \hyperlink{classParticleManager_585655101c829ec78e9abe6d849210fe}{cll\_\-sz}}
\label{classParticleManager_585655101c829ec78e9abe6d849210fe}

\begin{CompactList}\small\item\em cell size \item\end{CompactList}\item 
\hypertarget{classParticleManager_5e9cc4e55e0af75b59849f83a63bbec6}{
Vec2d \hyperlink{classParticleManager_5e9cc4e55e0af75b59849f83a63bbec6}{box\_\-size}}
\label{classParticleManager_5e9cc4e55e0af75b59849f83a63bbec6}

\begin{CompactList}\small\item\em computational domain size \item\end{CompactList}\item 
\hypertarget{classParticleManager_f91681a876131de64a0e2f6209d9bd89}{
int \hyperlink{classParticleManager_f91681a876131de64a0e2f6209d9bd89}{initial\_\-condition}}
\label{classParticleManager_f91681a876131de64a0e2f6209d9bd89}

\begin{CompactList}\small\item\em initial condition marker \item\end{CompactList}\item 
\hypertarget{classParticleManager_186eaba52a3b1a3ed7f098d1f2a56d9c}{
double \hyperlink{classParticleManager_186eaba52a3b1a3ed7f098d1f2a56d9c}{delta}}
\label{classParticleManager_186eaba52a3b1a3ed7f098d1f2a56d9c}

\begin{CompactList}\small\item\em the inital particle distance \item\end{CompactList}\item 
\hypertarget{classParticleManager_d77df8bc47beefcd049c4277c53d095a}{
int \hyperlink{classParticleManager_d77df8bc47beefcd049c4277c53d095a}{hdelta}}
\label{classParticleManager_d77df8bc47beefcd049c4277c53d095a}

\begin{CompactList}\small\item\em the ration between smoothing length and inital particle distance \item\end{CompactList}\item 
\hypertarget{classParticleManager_cb03c536574d333e0d733158aae652a1}{
Vec2d \hyperlink{classParticleManager_cb03c536574d333e0d733158aae652a1}{U0}}
\label{classParticleManager_cb03c536574d333e0d733158aae652a1}

\begin{CompactList}\small\item\em inital flow speed \item\end{CompactList}\item 
\hypertarget{classParticleManager_f19707700ab5014b13c1f623438d6b2f}{
double \textbf{rho0}}
\label{classParticleManager_f19707700ab5014b13c1f623438d6b2f}

\item 
\hypertarget{classParticleManager_729e474fe73f5d59f5448023798aee66}{
double \textbf{p0}}
\label{classParticleManager_729e474fe73f5d59f5448023798aee66}

\item 
\hypertarget{classParticleManager_edabda5cb1b485cd2ddaf10dad18f896}{
double \hyperlink{classParticleManager_edabda5cb1b485cd2ddaf10dad18f896}{T0}}
\label{classParticleManager_edabda5cb1b485cd2ddaf10dad18f896}

\begin{CompactList}\small\item\em initial particle mass and density, pressure and temperature \item\end{CompactList}\end{CompactItemize}

\subsection{Detailed Description}
\hyperlink{classParticle}{Particle} manager class.

\subsection{Constructor \& Destructor Documentation}
\hypertarget{classParticleManager_11a8f14c20f087713172d230de468166}{
\index{ParticleManager@{ParticleManager}!ParticleManager@{ParticleManager}}
\index{ParticleManager@{ParticleManager}!ParticleManager@{ParticleManager}}
\subsubsection[{ParticleManager}]{\setlength{\rightskip}{0pt plus 5cm}ParticleManager::ParticleManager ({\bf Initiation} \& {\em ini})}}
\label{classParticleManager_11a8f14c20f087713172d230de468166}

constructor

\begin{itemize}
\item copy properties from class \hyperlink{classInitiation}{Initiation}\end{itemize}

\begin{itemize}
\item strore the cell linked lists in a 2-d array \end{itemize}

\subsection{Member Function Documentation}
\hypertarget{classParticleManager_6ee12c3925f371782800a50b17ac0941}{
\index{ParticleManager@{ParticleManager}!BiuldRealParticles@{BiuldRealParticles}}
\index{BiuldRealParticles@{BiuldRealParticles}!ParticleManager@{ParticleManager}}
\subsubsection[{BiuldRealParticles}]{\setlength{\rightskip}{0pt plus 5cm}void ParticleManager::BiuldRealParticles ({\bf Hydrodynamics} \& {\em hydro}, \/  {\bf Initiation} \& {\em ini})}}
\label{classParticleManager_6ee12c3925f371782800a50b17ac0941}

buid the initial particles and the linked lists

initial particles either from .cfg file or from .rst file

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=171pt]{classParticleManager_6ee12c3925f371782800a50b17ac0941_cgraph}
\end{center}
\end{figure}

Here is the caller graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=205pt]{classParticleManager_6ee12c3925f371782800a50b17ac0941_icgraph}
\end{center}
\end{figure}
\hypertarget{classParticleManager_edebb3e37b06426f6509b35c0926bd59}{
\index{ParticleManager@{ParticleManager}!BiuldWallParticles@{BiuldWallParticles}}
\index{BiuldWallParticles@{BiuldWallParticles}!ParticleManager@{ParticleManager}}
\subsubsection[{BiuldWallParticles}]{\setlength{\rightskip}{0pt plus 5cm}void ParticleManager::BiuldWallParticles ({\bf Hydrodynamics} \& {\em hydro}, \/  {\bf Initiation} \& {\em ini}, \/  {\bf Boundary} \& {\em boundary})\hspace{0.3cm}{\tt  \mbox{[}private\mbox{]}}}}
\label{classParticleManager_edebb3e37b06426f6509b35c0926bd59}

buid the initial wall particles and the linked lists

left hand border (corresponds to the first column)

\begin{itemize}
\item iterate over the first colums of all cells that constitute this boundary

\begin{itemize}
\item create a new wall particle

\item insert its position on the particle list

\item insert the position into corresponding cell list\end{itemize}
\end{itemize}

right hand border(corresponds to the last column)

\begin{itemize}
\item iterate over the last colums of all cells that constitute this boundary

\begin{itemize}
\item create a new wall particle

\item insert its poistion on the particle list

\item insert the position into corresponding cell list\end{itemize}
\end{itemize}

lower border(corresponds to the lowest row)

\begin{itemize}
\item iterate over the lowest row of all cells that constitute this boundary

\begin{itemize}
\item creat a new wall particle

\item insert its poistion on the particle list

\item insert the position into corresponding cell list\end{itemize}
\end{itemize}

upper border (corresponds to the last row)

\begin{itemize}
\item iterate over the last rows of all cells that constitute this boundary

\begin{itemize}
\item create a new wall particle

\item insert its poistion on the particle list

\item insert the position into corresponding cell list\end{itemize}
\end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=158pt]{classParticleManager_edebb3e37b06426f6509b35c0926bd59_cgraph}
\end{center}
\end{figure}
\hypertarget{classParticleManager_8cc6658324694ad28d3c50bfc0cbd0f7}{
\index{ParticleManager@{ParticleManager}!BuildInteraction@{BuildInteraction}}
\index{BuildInteraction@{BuildInteraction}!ParticleManager@{ParticleManager}}
\subsubsection[{BuildInteraction}]{\setlength{\rightskip}{0pt plus 5cm}void ParticleManager::BuildInteraction ({\bf Llist}$<$ {\bf Interaction} $>$ \& {\em interactions}, \/  {\bf Llist}$<$ {\bf Particle} $>$ \& {\em particle\_\-list}, \/  {\bf Force} $\ast$$\ast$ {\em forces}, \/  {\bf QuinticSpline} \& {\em weight\_\-function})}}
\label{classParticleManager_8cc6658324694ad28d3c50bfc0cbd0f7}

build the interaction (particle pair) list

\begin{itemize}
\item iterate particles on the particle list

\begin{itemize}
\item choose origin particle

\item find out where(in which cell) the particle is

\item loop on this and all surrounding cells

\begin{itemize}
\item iterate this cell list

\begin{itemize}
\item calculate distance between particle in question and destination particle (which is iterated)and if interaction takes place: add pair to inetraction list ({\bf question: why is dst compared to h$^\wedge$2 and not support length to determine if there is interaction or not??} \end{itemize}
\end{itemize}
\end{itemize}
\end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=257pt]{classParticleManager_8cc6658324694ad28d3c50bfc0cbd0f7_cgraph}
\end{center}
\end{figure}

Here is the caller graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=350pt]{classParticleManager_8cc6658324694ad28d3c50bfc0cbd0f7_icgraph}
\end{center}
\end{figure}
\hypertarget{classParticleManager_4e0f9e5cb222739a83f0f80ee4ba265b}{
\index{ParticleManager@{ParticleManager}!BuildNNP@{BuildNNP}}
\index{BuildNNP@{BuildNNP}!ParticleManager@{ParticleManager}}
\subsubsection[{BuildNNP}]{\setlength{\rightskip}{0pt plus 5cm}void ParticleManager::BuildNNP (Vec2d \& {\em point})}}
\label{classParticleManager_4e0f9e5cb222739a83f0f80ee4ba265b}

do NNP search around a point and build the NNP list

\begin{itemize}
\item clear the list first

\item loop on this and all surrounding cells

\begin{itemize}
\item iterate this cell list

\begin{itemize}
\item check the position of the particle and (if particle is NNP) insert the nearest particle to the list

\item (line 137){\bf Question: WHY SMOOTHINGLENGTH AND NOT SUPPORT LENGT???} \end{itemize}
\end{itemize}
\end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=151pt]{classParticleManager_4e0f9e5cb222739a83f0f80ee4ba265b_cgraph}
\end{center}
\end{figure}

Here is the caller graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=328pt]{classParticleManager_4e0f9e5cb222739a83f0f80ee4ba265b_icgraph}
\end{center}
\end{figure}
\hypertarget{classParticleManager_b68ab55a9429ac8d32368b531487dcec}{
\index{ParticleManager@{ParticleManager}!BuildNNP\_\-MLSMapping@{BuildNNP\_\-MLSMapping}}
\index{BuildNNP\_\-MLSMapping@{BuildNNP\_\-MLSMapping}!ParticleManager@{ParticleManager}}
\subsubsection[{BuildNNP\_\-MLSMapping}]{\setlength{\rightskip}{0pt plus 5cm}void ParticleManager::BuildNNP\_\-MLSMapping (Vec2d \& {\em point})}}
\label{classParticleManager_b68ab55a9429ac8d32368b531487dcec}

do NNP search around a point and build the NNP list for \hyperlink{classMLS}{MLS} approximation

\begin{itemize}
\item loop on this and all surrounding cells

\begin{itemize}
\item iterate this cell list

\begin{itemize}
\item check the position of the real particle and (if particle is NNP) insert it to the list \end{itemize}
\end{itemize}
\end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=175pt]{classParticleManager_b68ab55a9429ac8d32368b531487dcec_cgraph}
\end{center}
\end{figure}
\hypertarget{classParticleManager_740244439672f753735abc5a6a7cc3bc}{
\index{ParticleManager@{ParticleManager}!UpdateCellLinkedLists@{UpdateCellLinkedLists}}
\index{UpdateCellLinkedLists@{UpdateCellLinkedLists}!ParticleManager@{ParticleManager}}
\subsubsection[{UpdateCellLinkedLists}]{\setlength{\rightskip}{0pt plus 5cm}void ParticleManager::UpdateCellLinkedLists ()}}
\label{classParticleManager_740244439672f753735abc5a6a7cc3bc}

update the cell linked lists for real particles

\begin{itemize}
\item (double) loop (as indices i, j) on all cells

\begin{itemize}
\item iterate this cell list

\item if the list is empty or the node position is at the end {\bf !!!Question!!! is this comment right? would it not rather be...if list NOT empty and NOT at the end }

\begin{itemize}
\item check the position of the real particle

\begin{itemize}
\item if the partilce runs out of the current cell

\begin{itemize}
\item delete the current node

\item insert it to the new cell linked list\end{itemize}
\end{itemize}
\end{itemize}
\end{itemize}
\end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=171pt]{classParticleManager_740244439672f753735abc5a6a7cc3bc_cgraph}
\end{center}
\end{figure}

Here is the caller graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=260pt]{classParticleManager_740244439672f753735abc5a6a7cc3bc_icgraph}
\end{center}
\end{figure}

The documentation for this class was generated from the following files:\begin{CompactItemize}
\item 
\hyperlink{particlemanager_8h}{particlemanager.h}\item 
particlemanager.cpp\end{CompactItemize}