### 2010-06-17 src/latex/classQuinticSpline.tex

\hypertarget{classQuinticSpline}{
\section{QuinticSpline Class Reference}
\label{classQuinticSpline}\index{QuinticSpline@{QuinticSpline}}
}
Quintic spline class.

{\tt \#include $<$quinticspline.h$>$}

Inheritance diagram for QuinticSpline:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=118pt]{classQuinticSpline__inherit__graph}
\end{center}
\end{figure}
Collaboration diagram for QuinticSpline:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=118pt]{classQuinticSpline__coll__graph}
\end{center}
\end{figure}
\subsection*{Public Member Functions}
\begin{CompactItemize}
\item 
\hypertarget{classQuinticSpline_e2d0a39985327a020d2794516f895f27}{
\hyperlink{classQuinticSpline_e2d0a39985327a020d2794516f895f27}{QuinticSpline} (const double \hyperlink{classKernel_f088821dba4a00d1d320b27ebcb71258}{smoothingLength})}
\label{classQuinticSpline_e2d0a39985327a020d2794516f895f27}

\begin{CompactList}\small\item\em constructor to initialize the data members (auxiliary factors for kernel calculation) \item\end{CompactList}\item 
virtual double \hyperlink{classQuinticSpline_021acaeb55230bf10cf3a2c634381088}{w} (const double distance) const 
\begin{CompactList}\small\item\em Calculates the kernel value for the given distance of two particles. \item\end{CompactList}\item 
virtual Vec2d \hyperlink{classQuinticSpline_00a31074e58e60a4721144e4935c56b2}{gradW} (const double distance, const Vec2d \&distanceVector) const 
\begin{CompactList}\small\item\em Calculates the kernel derivation for the given distance of two particles. \item\end{CompactList}\item 
\hypertarget{classQuinticSpline_67fa50759531bbbbef728d28282788e7}{
double \hyperlink{classQuinticSpline_67fa50759531bbbbef728d28282788e7}{F} (const double distance) const }
\label{classQuinticSpline_67fa50759531bbbbef728d28282788e7}

\begin{CompactList}\small\item\em Calculates the kernel derivation (a double, not a vector) to distance. \item\end{CompactList}\item 
\hypertarget{classQuinticSpline_adc23e95cd346c434b423b3276156e8c}{
double \hyperlink{classQuinticSpline_adc23e95cd346c434b423b3276156e8c}{LapW} (const double distance) const }
\label{classQuinticSpline_adc23e95cd346c434b423b3276156e8c}

\begin{CompactList}\small\item\em Calculates the kernel Laplacian. \item\end{CompactList}\end{CompactItemize}
\subsection*{Private Attributes}
\begin{CompactItemize}
\item 
\hypertarget{classQuinticSpline_bd65ef2de82ce850ec80393582726532}{
const double \hyperlink{classQuinticSpline_bd65ef2de82ce850ec80393582726532}{norm}}
\label{classQuinticSpline_bd65ef2de82ce850ec80393582726532}

\begin{CompactList}\small\item\em Normalization factor. \item\end{CompactList}\item 
\hypertarget{classQuinticSpline_743bd6fe1beb6049f5bdcf76beba04e2}{
double \hyperlink{classQuinticSpline_743bd6fe1beb6049f5bdcf76beba04e2}{reciprocH}}
\label{classQuinticSpline_743bd6fe1beb6049f5bdcf76beba04e2}

\begin{CompactList}\small\item\em Auxiliary factors for intermediate results: The inverse smoothing length. \item\end{CompactList}\item 
\hypertarget{classQuinticSpline_e260195cb042866d86a822f469a1277f}{
double \hyperlink{classQuinticSpline_e260195cb042866d86a822f469a1277f}{factorW}}
\label{classQuinticSpline_e260195cb042866d86a822f469a1277f}

\begin{CompactList}\small\item\em Auxiliary factors for intermediate results: A pre-factor for w. \item\end{CompactList}\item 
\hypertarget{classQuinticSpline_fda98a6d7b9feda4e30871111ece88b3}{
double \hyperlink{classQuinticSpline_fda98a6d7b9feda4e30871111ece88b3}{factorGradW}}
\label{classQuinticSpline_fda98a6d7b9feda4e30871111ece88b3}

\begin{CompactList}\small\item\em Auxiliary factors for intermediate results: A pre-factor for grad w. \item\end{CompactList}\item 
\hypertarget{classQuinticSpline_5cf5d05e9069c965ffc315770a1206ec}{
double \hyperlink{classQuinticSpline_5cf5d05e9069c965ffc315770a1206ec}{factorLapW}}
\label{classQuinticSpline_5cf5d05e9069c965ffc315770a1206ec}

\begin{CompactList}\small\item\em Auxiliary factors for Laplacian. \item\end{CompactList}\end{CompactItemize}

\subsection{Detailed Description}
Quintic spline class.

\subsection{Member Function Documentation}
\hypertarget{classQuinticSpline_00a31074e58e60a4721144e4935c56b2}{
\index{QuinticSpline@{QuinticSpline}!gradW@{gradW}}
\index{gradW@{gradW}!QuinticSpline@{QuinticSpline}}
\subsubsection[{gradW}]{\setlength{\rightskip}{0pt plus 5cm}Vec2d QuinticSpline::gradW (const double {\em distance}, \/  const Vec2d \& {\em distanceVector}) const\hspace{0.3cm}{\tt  \mbox{[}virtual\mbox{]}}}}
\label{classQuinticSpline_00a31074e58e60a4721144e4935c56b2}

Calculates the kernel derivation for the given distance of two particles.

We take this from Morris, Fox and Zhu (1997) but used a tripled smoothing length for the definition of the interaction radius.

Implements \hyperlink{classKernel_545e61b98db05db6dd02b88e2ad3da23}{Kernel}.\hypertarget{classQuinticSpline_021acaeb55230bf10cf3a2c634381088}{
\index{QuinticSpline@{QuinticSpline}!w@{w}}
\index{w@{w}!QuinticSpline@{QuinticSpline}}
\subsubsection[{w}]{\setlength{\rightskip}{0pt plus 5cm}double QuinticSpline::w (const double {\em distance}) const\hspace{0.3cm}{\tt  \mbox{[}virtual\mbox{]}}}}
\label{classQuinticSpline_021acaeb55230bf10cf3a2c634381088}

Calculates the kernel value for the given distance of two particles.

We take this from Morris, Fox and Zhu (1997) but used a tripled smoothing length for the definition of the interaction radius.

Implements \hyperlink{classKernel_6b4d26ba99457a4acbbbca1454dbdc8a}{Kernel}.

Here is the caller graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=420pt]{classQuinticSpline_021acaeb55230bf10cf3a2c634381088_icgraph}
\end{center}
\end{figure}

The documentation for this class was generated from the following files:\begin{CompactItemize}
\item 
\hyperlink{quinticspline_8h}{quinticspline.h}\item 
\hyperlink{quinticspline_8cpp}{quinticspline.cpp}\end{CompactItemize}