### 2010-06-17 src/latex/classTimeSolver.tex

\hypertarget{classTimeSolver}{
\section{TimeSolver Class Reference}
\label{classTimeSolver}\index{TimeSolver@{TimeSolver}}
}
Time solver class.

{\tt \#include $<$timesolver.h$>$}

\subsection*{Public Member Functions}
\begin{CompactItemize}
\item 
\hyperlink{classTimeSolver_c5ed4d99be532dd9c3e77424b3e7a60a}{TimeSolver} (\hyperlink{classInitiation}{Initiation} \&ini)
\begin{CompactList}\small\item\em constructor \item\end{CompactList}\item 
void \hyperlink{classTimeSolver_edd74b3f1e292277b972b938787f4a3c}{TimeIntegral} (\hyperlink{classHydrodynamics}{Hydrodynamics} \&hydro, \hyperlink{classParticleManager}{ParticleManager} \&particles, \hyperlink{classBoundary}{Boundary} \&boundary, double \&Time, double D\_\-time, \hyperlink{classDiagnose}{Diagnose} \&diagnose, \hyperlink{classInitiation}{Initiation} \&ini, \hyperlink{classQuinticSpline}{QuinticSpline} \&weight\_\-function, \hyperlink{classMLS}{MLS} \&mls)
\begin{CompactList}\small\item\em advance time interval D\_\-time (=output time interval) \item\end{CompactList}\item 
void \hyperlink{classTimeSolver_2cbbdd8eecbd3aae12c153f8b9da1402}{TimeIntegral\_\-summation} (\hyperlink{classHydrodynamics}{Hydrodynamics} \&hydro, \hyperlink{classParticleManager}{ParticleManager} \&particles, \hyperlink{classBoundary}{Boundary} \&boundary, double \&Time, double D\_\-time, \hyperlink{classDiagnose}{Diagnose} \&diagnose, \hyperlink{classInitiation}{Initiation} \&ini, \hyperlink{classQuinticSpline}{QuinticSpline} \&weight\_\-function, \hyperlink{classMLS}{MLS} \&mls)
\begin{CompactList}\small\item\em advance time interval D\_\-time with summation for density \item\end{CompactList}\end{CompactItemize}
\subsection*{Private Attributes}
\begin{CompactItemize}
\item 
\hypertarget{classTimeSolver_6650fefd2ce17fe4b84c21d2cf84b3c0}{
double \hyperlink{classTimeSolver_6650fefd2ce17fe4b84c21d2cf84b3c0}{cell\_\-size}}
\label{classTimeSolver_6650fefd2ce17fe4b84c21d2cf84b3c0}

\begin{CompactList}\small\item\em will be copied from initiation \item\end{CompactList}\item 
\hypertarget{classTimeSolver_091ff25ed4144106f86ec1b0043a2b00}{
Vec2d \hyperlink{classTimeSolver_091ff25ed4144106f86ec1b0043a2b00}{box\_\-size}}
\label{classTimeSolver_091ff25ed4144106f86ec1b0043a2b00}

\begin{CompactList}\small\item\em will be copied from initiation \item\end{CompactList}\item 
\hypertarget{classTimeSolver_8a435c6de2503a8cd74868e19a9ed099}{
double \hyperlink{classTimeSolver_8a435c6de2503a8cd74868e19a9ed099}{smoothinglength}}
\label{classTimeSolver_8a435c6de2503a8cd74868e19a9ed099}

\begin{CompactList}\small\item\em will be copied from initiation \item\end{CompactList}\item 
\hypertarget{classTimeSolver_059eae3a0771b4d3201aea0aef1aaee4}{
int \hyperlink{classTimeSolver_059eae3a0771b4d3201aea0aef1aaee4}{ite}}
\label{classTimeSolver_059eae3a0771b4d3201aea0aef1aaee4}

\begin{CompactList}\small\item\em number of iteration \item\end{CompactList}\item 
\hypertarget{classTimeSolver_66c0a9afc7f8fbc205e4ce4de9435aed}{
double \hyperlink{classTimeSolver_66c0a9afc7f8fbc205e4ce4de9435aed}{dt}}
\label{classTimeSolver_66c0a9afc7f8fbc205e4ce4de9435aed}

\begin{CompactList}\small\item\em time step \item\end{CompactList}\end{CompactItemize}

\subsection{Detailed Description}
Time solver class.

\subsection{Constructor \& Destructor Documentation}
\hypertarget{classTimeSolver_c5ed4d99be532dd9c3e77424b3e7a60a}{
\index{TimeSolver@{TimeSolver}!TimeSolver@{TimeSolver}}
\index{TimeSolver@{TimeSolver}!TimeSolver@{TimeSolver}}
\subsubsection[{TimeSolver}]{\setlength{\rightskip}{0pt plus 5cm}TimeSolver::TimeSolver ({\bf Initiation} \& {\em ini})}}
\label{classTimeSolver_c5ed4d99be532dd9c3e77424b3e7a60a}

constructor

\begin{itemize}
\item copy properties from class \hyperlink{classInitiation}{Initiation}\end{itemize}

\begin{itemize}
\item initialize the iteration \end{itemize}

\subsection{Member Function Documentation}
\hypertarget{classTimeSolver_edd74b3f1e292277b972b938787f4a3c}{
\index{TimeSolver@{TimeSolver}!TimeIntegral@{TimeIntegral}}
\index{TimeIntegral@{TimeIntegral}!TimeSolver@{TimeSolver}}
\subsubsection[{TimeIntegral}]{\setlength{\rightskip}{0pt plus 5cm}void TimeSolver::TimeIntegral ({\bf Hydrodynamics} \& {\em hydro}, \/  {\bf ParticleManager} \& {\em particles}, \/  {\bf Boundary} \& {\em boundary}, \/  double \& {\em Time}, \/  double {\em D\_\-time}, \/  {\bf Diagnose} \& {\em diagnose}, \/  {\bf Initiation} \& {\em ini}, \/  {\bf QuinticSpline} \& {\em weight\_\-function}, \/  {\bf MLS} \& {\em mls})}}
\label{classTimeSolver_edd74b3f1e292277b972b938787f4a3c}

advance time interval D\_\-time (=output time interval)

\begin{itemize}
\item print out screen information for the iteration

\item calculating diagonse information

\item the prediction step

\begin{enumerate}
\item hydro.BuildPair

\item hydro.UpdateChangeRate

\item hydro.Predictor

\item hydro.UpdateState\end{enumerate}

\item the correction step without update the interaction list

\begin{enumerate}
\item hydro.UpdatePair

\item boundary.BoundaryCondition

\item hydro.UpdateChangeRate

\item hydro.UpdateRandom

\item hydro.Corrector

\item hydro.RandomEffects

\item hydro.UpdateState\end{enumerate}

\item renew boundary particles

\begin{enumerate}
\item boundary.RunAwayCheck

\item particles.UpdateCellLinkedLists

\item boundary.BuildBoundaryParticles\end{enumerate}
\end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=420pt]{classTimeSolver_edd74b3f1e292277b972b938787f4a3c_cgraph}
\end{center}
\end{figure}
\hypertarget{classTimeSolver_2cbbdd8eecbd3aae12c153f8b9da1402}{
\index{TimeSolver@{TimeSolver}!TimeIntegral\_\-summation@{TimeIntegral\_\-summation}}
\index{TimeIntegral\_\-summation@{TimeIntegral\_\-summation}!TimeSolver@{TimeSolver}}
\subsubsection[{TimeIntegral\_\-summation}]{\setlength{\rightskip}{0pt plus 5cm}void TimeSolver::TimeIntegral\_\-summation ({\bf Hydrodynamics} \& {\em hydro}, \/  {\bf ParticleManager} \& {\em particles}, \/  {\bf Boundary} \& {\em boundary}, \/  double \& {\em Time}, \/  double {\em D\_\-time}, \/  {\bf Diagnose} \& {\em diagnose}, \/  {\bf Initiation} \& {\em ini}, \/  {\bf QuinticSpline} \& {\em weight\_\-function}, \/  {\bf MLS} \& {\em mls})}}
\label{classTimeSolver_2cbbdd8eecbd3aae12c153f8b9da1402}

advance time interval D\_\-time with summation for density

\begin{itemize}
\item screen information for the iteration

\item calculating diagnose information

\item the prediction step

\begin{enumerate}
\item hydro.buildPair

\item hydro.UpdateDensity

\item boundary.BoundaryCondition

\item hydro.UpdatePhaseGradient

\item boundary.BoundaryCondition

\item hydro.UpdateSurfaceStress

\item hydro.UpdateChangeRate

\item hydro.Predictor\_\-summation\end{enumerate}

\item the correction step without update the interaction list

\begin{enumerate}
\item boundary.BoundaryCondition

\item hydro.UpdatePair

\item hydro.UpdateDensity

\item boundary.BoundaryCondition

\item hydro.UpdatePhaseGradient

\item boundary.BoundaryCondition

\item hydro.UpdateSurfaceStress

\item hydro.UpdateChangeRate

\item hydro.Corrector\_\-summation\end{enumerate}

\item renew boundary particles

\begin{enumerate}
\item boundary.RunAwayCheck

\item particles.UpdateCellLinkedLists

\item boundary.BuildBoundaryParticles\end{enumerate}
\end{itemize}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=409pt]{classTimeSolver_2cbbdd8eecbd3aae12c153f8b9da1402_cgraph}
\end{center}
\end{figure}

Here is the caller graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=147pt]{classTimeSolver_2cbbdd8eecbd3aae12c153f8b9da1402_icgraph}
\end{center}
\end{figure}

The documentation for this class was generated from the following files:\begin{CompactItemize}
\item 
\hyperlink{timesolver_8h}{timesolver.h}\item 
\hyperlink{timesolver_8cpp}{timesolver.cpp}\end{CompactItemize}