### 2010-06-17 src/latex/classWiener.tex

\hypertarget{classWiener}{
\section{Wiener Class Reference}
\label{classWiener}\index{Wiener@{Wiener}}
}
\hyperlink{classWiener}{Wiener} process.

{\tt \#include $<$wiener.h$>$}

\subsection*{Public Member Functions}
\begin{CompactItemize}
\item 
\hypertarget{classWiener_9b2daa0e357c11b963463e6c2a510ad1}{
\hyperlink{classWiener_9b2daa0e357c11b963463e6c2a510ad1}{Wiener} ()}
\label{classWiener_9b2daa0e357c11b963463e6c2a510ad1}

\begin{CompactList}\small\item\em costructure \item\end{CompactList}\item 
\hypertarget{classWiener_454f6ab52cf050daf682fa6f374504ea}{
\textbf{Wiener} (const int dimension)}
\label{classWiener_454f6ab52cf050daf682fa6f374504ea}

\item 
\hypertarget{classWiener_9bf2a13de68b15f52e6b3ff5a9c485b6}{
void \hyperlink{classWiener_9bf2a13de68b15f52e6b3ff5a9c485b6}{get\_\-wiener\_\-Espanol} (const double sqrtdt)}
\label{classWiener_9bf2a13de68b15f52e6b3ff5a9c485b6}

\begin{CompactList}\small\item\em produce wiener vlaue matrix with symmetric tracless part, and the trace \item\end{CompactList}\item 
\hypertarget{classWiener_be754cfa9c89ec2e405126f26997e0f0}{
void \hyperlink{classWiener_be754cfa9c89ec2e405126f26997e0f0}{get\_\-wiener} (const double sqrtdt)}
\label{classWiener_be754cfa9c89ec2e405126f26997e0f0}

\begin{CompactList}\small\item\em wiener process with Flekkoy's form \item\end{CompactList}\item 
\hypertarget{classWiener_2fb89a7b43c2d6378ff6698614907098}{
void \hyperlink{classWiener_2fb89a7b43c2d6378ff6698614907098}{Ranils} ()}
\label{classWiener_2fb89a7b43c2d6378ff6698614907098}

\begin{CompactList}\small\item\em set the random seed \item\end{CompactList}\end{CompactItemize}
\subsection*{Public Attributes}
\begin{CompactItemize}
\item 
\hypertarget{classWiener_cd79f000952f14b816c3cdd008c34a2a}{
double $\ast$$\ast$ \textbf{sym\_\-trclss}}
\label{classWiener_cd79f000952f14b816c3cdd008c34a2a}

\item 
\hypertarget{classWiener_c24cf9f8fd8385325584c190f93f0abe}{
double \textbf{trace\_\-d}}
\label{classWiener_c24cf9f8fd8385325584c190f93f0abe}

\item 
\hypertarget{classWiener_f47c4ccdb5cc42751beaf9023e64d426}{
double \textbf{Random\_\-p}}
\label{classWiener_f47c4ccdb5cc42751beaf9023e64d426}

\item 
\hypertarget{classWiener_411516a928b7d38aa170c2311af6c26a}{
double \textbf{Random\_\-v}}
\label{classWiener_411516a928b7d38aa170c2311af6c26a}

\end{CompactItemize}
\subsection*{Private Member Functions}
\begin{CompactItemize}
\item 
\hypertarget{classWiener_238e64201f993e2e4fff5c6615b1cf85}{
double \hyperlink{classWiener_238e64201f993e2e4fff5c6615b1cf85}{Ranuls} ()}
\label{classWiener_238e64201f993e2e4fff5c6615b1cf85}

\begin{CompactList}\small\item\em get the random number uniform distributed in \mbox{[}0, 1\mbox{]} \item\end{CompactList}\item 
void \hyperlink{classWiener_afdd7d1aa80e1609edd7bcc7532d56d6}{Gaussian} (double \&y1, double \&y2)
\end{CompactItemize}
\subsection*{Private Attributes}
\begin{CompactItemize}
\item 
\hypertarget{classWiener_4bf535cc91f737b904552f51ae1f0ac6}{
int \textbf{dimension}}
\label{classWiener_4bf535cc91f737b904552f51ae1f0ac6}

\item 
\hypertarget{classWiener_09f6069684c20e3eecdddd17602fcf05}{
double $\ast$$\ast$ \textbf{randoms}}
\label{classWiener_09f6069684c20e3eecdddd17602fcf05}

\item 
long int \hyperlink{classWiener_eb32502c4439ae23e7a6922c5d8cd3ba}{idum}
\item 
\hypertarget{classWiener_b06d8f8b4ff3e2c1c099570a66c1ce84}{
long int \textbf{idum2}}
\label{classWiener_b06d8f8b4ff3e2c1c099570a66c1ce84}

\item 
\hypertarget{classWiener_040aaffc7d7758d597582660a09656d3}{
long int \textbf{iy}}
\label{classWiener_040aaffc7d7758d597582660a09656d3}

\item 
\hypertarget{classWiener_f6aa4a6e172025e1ab3f9150e69f865a}{
long int \textbf{ntab}}
\label{classWiener_f6aa4a6e172025e1ab3f9150e69f865a}

\item 
\hypertarget{classWiener_56f50cb31e458ec78b843809ac4a0101}{
long int \textbf{iseed}}
\label{classWiener_56f50cb31e458ec78b843809ac4a0101}

\item 
\hypertarget{classWiener_353476d32254b68fdedaee29f3ae137d}{
long int $\ast$ \textbf{iv}}
\label{classWiener_353476d32254b68fdedaee29f3ae137d}

\end{CompactItemize}

\subsection{Detailed Description}
\hyperlink{classWiener}{Wiener} process.

----------------------------------------------------------------------- The class defining \hyperlink{classWiener}{Wiener} process in two dimensions -----------------------------------------------------------------------

\subsection{Member Function Documentation}
\hypertarget{classWiener_afdd7d1aa80e1609edd7bcc7532d56d6}{
\index{Wiener@{Wiener}!Gaussian@{Gaussian}}
\index{Gaussian@{Gaussian}!Wiener@{Wiener}}
\subsubsection[{Gaussian}]{\setlength{\rightskip}{0pt plus 5cm}void Wiener::Gaussian (double \& {\em y1}, \/  double \& {\em y2})\hspace{0.3cm}{\tt  \mbox{[}private\mbox{]}}}}
\label{classWiener_afdd7d1aa80e1609edd7bcc7532d56d6}

get two random numbers y1, y2 with guassian distribution with zero mean and variance one from random numbers uniform distributed in \mbox{[}0, 1\mbox{]}

Here is the call graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=129pt]{classWiener_afdd7d1aa80e1609edd7bcc7532d56d6_cgraph}
\end{center}
\end{figure}

Here is the caller graph for this function:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=420pt]{classWiener_afdd7d1aa80e1609edd7bcc7532d56d6_icgraph}
\end{center}
\end{figure}

\subsection{Member Data Documentation}
\hypertarget{classWiener_eb32502c4439ae23e7a6922c5d8cd3ba}{
\index{Wiener@{Wiener}!idum@{idum}}
\index{idum@{idum}!Wiener@{Wiener}}
\subsubsection[{idum}]{\setlength{\rightskip}{0pt plus 5cm}long int {\bf Wiener::idum}\hspace{0.3cm}{\tt  \mbox{[}private\mbox{]}}}}
\label{classWiener_eb32502c4439ae23e7a6922c5d8cd3ba}

for random number generator variables

The documentation for this class was generated from the following files:\begin{CompactItemize}
\item 
\hyperlink{wiener_8h}{wiener.h}\item 
wiener.cpp\end{CompactItemize}