### 2010-06-17 src/latex/dllist_8h.tex

\hypertarget{dllist_8h}{
\section{dllist.h File Reference}
\label{dllist_8h}\index{dllist.h@{dllist.h}}
}
Double linked list implementation.

This graph shows which files directly or indirectly include this file:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=420pt]{dllist_8h__dep__incl}
\end{center}
\end{figure}
\subsection*{Classes}
\begin{CompactItemize}
\item 
class \hyperlink{classLlistNode}{LlistNode$<$ Ldata $>$}
\begin{CompactList}\small\item\em A template node in the linked list. \item\end{CompactList}\item 
class \hyperlink{classLlist}{Llist$<$ Ldata $>$}
\begin{CompactList}\small\item\em Double linked list implementation. \item\end{CompactList}\end{CompactItemize}

\label{_details}
\hypertarget{_details}{}
\subsection{Detailed Description}
Double linked list implementation.

\begin{Desc}
\item[Author:]Xiangyu Hu $<$\href{mailto:Xiangyu.Hu@aer.mw.tum.de}{\tt Xiangyu.Hu@aer.mw.tum.de}$>$

changes by: Martin Bernreuther $<$\href{mailto:Martin.Bernreuther@ipvs.uni-stuttgart.de}{\tt Martin.Bernreuther@ipvs.uni-stuttgart.de}$>$

changes by: Andreas Mattes Declares and defines a generic double linked list template \end{Desc}