### 2010-06-17 src/latex/glbfunc_8h.tex

\hypertarget{glbfunc_8h}{
\section{glbfunc.h File Reference}
\label{glbfunc_8h}\index{glbfunc.h@{glbfunc.h}}
}

This graph shows which files directly or indirectly include this file:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=420pt]{glbfunc_8h__dep__incl}
\end{center}
\end{figure}
\subsection*{Functions}
\begin{CompactItemize}
\item 
\hypertarget{glbfunc_8h_78a6c97f8b8f5e0fbb4e52d0aa6110aa}{
double \hyperlink{glbfunc_8h_78a6c97f8b8f5e0fbb4e52d0aa6110aa}{AMAX1} (double a, double b)}
\label{glbfunc_8h_78a6c97f8b8f5e0fbb4e52d0aa6110aa}

\begin{CompactList}\small\item\em a. Get the maximum \item\end{CompactList}\item 
\hypertarget{glbfunc_8h_70d74bfbfb9d5e299e3effddb147b5ce}{
double \hyperlink{glbfunc_8h_70d74bfbfb9d5e299e3effddb147b5ce}{AMIN1} (double a, double b)}
\label{glbfunc_8h_70d74bfbfb9d5e299e3effddb147b5ce}

\begin{CompactList}\small\item\em b. Get the minimum \item\end{CompactList}\item 
\hypertarget{glbfunc_8h_ecdd3901ee6e4ecdafd81998cef6cafe}{
double \hyperlink{glbfunc_8h_ecdd3901ee6e4ecdafd81998cef6cafe}{ABS} (double a)}
\label{glbfunc_8h_ecdd3901ee6e4ecdafd81998cef6cafe}

\begin{CompactList}\small\item\em c. Get the absolute \item\end{CompactList}\item 
\hypertarget{glbfunc_8h_4285ca4ceabec63382f74718de72daf0}{
double \hyperlink{glbfunc_8h_4285ca4ceabec63382f74718de72daf0}{SIGN} (double a, double b)}
\label{glbfunc_8h_4285ca4ceabec63382f74718de72daf0}

\begin{CompactList}\small\item\em d. Sign of the first value is determined by the secomax\_\-species\_\-number value's sign \item\end{CompactList}\item 
\hypertarget{glbfunc_8h_0b49a61653d985585465515587b087ee}{
double \hyperlink{glbfunc_8h_0b49a61653d985585465515587b087ee}{AMAX4} (double a, double b, double c, double d)}
\label{glbfunc_8h_0b49a61653d985585465515587b087ee}

\begin{CompactList}\small\item\em e. Get the maximum from four \item\end{CompactList}\item 
\hypertarget{glbfunc_8h_7bf26e1cff694aad68df394cb63c1a31}{
int \hyperlink{glbfunc_8h_7bf26e1cff694aad68df394cb63c1a31}{SymmetricInverse3x3} (double M\mbox{[}3\mbox{]}\mbox{[}3\mbox{]})}
\label{glbfunc_8h_7bf26e1cff694aad68df394cb63c1a31}

\begin{CompactList}\small\item\em f. Get the inverse of symmetric array \item\end{CompactList}\item 
\hypertarget{glbfunc_8h_810e8e13a2409b115ead17fdf00760d1}{
double \hyperlink{glbfunc_8h_810e8e13a2409b115ead17fdf00760d1}{Kronecker} (int a, int b)}
\label{glbfunc_8h_810e8e13a2409b115ead17fdf00760d1}

\begin{CompactList}\small\item\em g. Kronecker delta \item\end{CompactList}\item 
\hypertarget{glbfunc_8h_3cab9f4eb994d85cb307ef4ca698b923}{
double \hyperlink{glbfunc_8h_3cab9f4eb994d85cb307ef4ca698b923}{K\_\-sign} (int a, int b)}
\label{glbfunc_8h_3cab9f4eb994d85cb307ef4ca698b923}

\begin{CompactList}\small\item\em h. Sign function \item\end{CompactList}\item 
\hypertarget{glbfunc_8h_3b0843a41bc641e5553d442b94b1f3e1}{
double \hyperlink{glbfunc_8h_3b0843a41bc641e5553d442b94b1f3e1}{compare} (int a, int b)}
\label{glbfunc_8h_3b0843a41bc641e5553d442b94b1f3e1}

\begin{CompactList}\small\item\em h. comparing \item\end{CompactList}\item 
\hypertarget{glbfunc_8h_21546457e67986c8772ab1ce75ab8e7a}{
double \hyperlink{glbfunc_8h_21546457e67986c8772ab1ce75ab8e7a}{sq} (double a)}
\label{glbfunc_8h_21546457e67986c8772ab1ce75ab8e7a}

\begin{CompactList}\small\item\em i. square \item\end{CompactList}\end{CompactItemize}
\subsection*{Variables}
\begin{CompactItemize}
\item 
const double \hyperlink{glbfunc_8h_43016d873124d39034edb8cd164794db}{pi} = 3.141592653589793
\begin{CompactList}\small\item\em Pi. \item\end{CompactList}\end{CompactItemize}

\label{_details}
\hypertarget{_details}{}
\subsection{Detailed Description}
\begin{Desc}
\item[Author:]Xiangyu Hu $<$\href{mailto:Xiangyu.Hu@aer.mw.tum.de}{\tt Xiangyu.Hu@aer.mw.tum.de}$>$  by: Martin Bernreuther $<$\href{mailto:Martin.Bernreuther@ipvs.uni-stuttgart.de}{\tt Martin.Bernreuther@ipvs.uni-stuttgart.de}$>$, \end{Desc}

\subsection{Variable Documentation}
\hypertarget{glbfunc_8h_43016d873124d39034edb8cd164794db}{
\index{glbfunc.h@{glbfunc.h}!pi@{pi}}
\index{pi@{pi}!glbfunc.h@{glbfunc.h}}
\subsubsection[{pi}]{\setlength{\rightskip}{0pt plus 5cm}const double {\bf pi} = 3.141592653589793}}
\label{glbfunc_8h_43016d873124d39034edb8cd164794db}

Pi.

---------------------------------------------------------------------------------------- Define globle paremeters and functions \hyperlink{glbfunc_8h}{glbfunc.h} ----------------------------------------------------------------------------------------