### 2010-06-17 src/latex/index.tex

\hypertarget{index_intro_sec}{}\section{Introduction}\label{index_intro_sec}
SPH and SDPD 2D code\hypertarget{index_install_sec}{}\section{Installation}\label{index_install_sec}
Check the source code from github and compile

\begin{footnotesize}\begin{verbatim}
   git clone git://github.com/slitvinov/sph-blitz
   cd sph-blitz
   ./local-install.sh  \end{verbatim}
\end{footnotesize}
\hypertarget{index_restart_file}{}\section{Restart file format}\label{index_restart_file}
\hypertarget{index_input_file}{}\section{Input file format}\label{index_input_file}
\hypertarget{index_sim}{}\section{Runnig simulations}\label{index_sim}

\begin{footnotesize}\begin{verbatim}
   cd src
   ./sph ../cases/couette \end{verbatim}
\end{footnotesize}
 \hypertarget{index_post}{}\section{Postprocessing}\label{index_post}
To combine all time snapshots in one file

\begin{footnotesize}\begin{verbatim}
   cd outdata/
   ../../scripts/dat2punto.sh > punto.dat
 \end{verbatim}
\end{footnotesize}
\hypertarget{index_vis}{}\section{Visuzalization}\label{index_vis}
\hypertarget{index_vis_gnuplot}{}\subsection{gnuplot}\label{index_vis_gnuplot}
 \hypertarget{index_vis_punto}{}\subsection{punto}\label{index_vis_punto}

\begin{footnotesize}\begin{verbatim}
    punto -D 2 -c 4 -B 0:0:0.04:0.04 -G -0.6:0.6 -s 8 -lc black -bg white  punto.dat \end{verbatim}
\end{footnotesize}