### 2010-06-17 src/latex/initiation_8cpp.tex

\hypertarget{initiation_8cpp}{
\section{initiation.cpp File Reference}
\label{initiation_8cpp}\index{initiation.cpp@{initiation.cpp}}
}
{\tt \#include $<$iostream$>$}\par
{\tt \#include $<$fstream$>$}\par
{\tt \#include $<$string$>$}\par
{\tt \#include $<$cstdio$>$}\par
{\tt \#include $<$cstdlib$>$}\par
{\tt \#include \char`\"{}glbcls.h\char`\"{}}\par
{\tt \#include \char`\"{}glbfunc.h\char`\"{}}\par

Include dependency graph for initiation.cpp:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=420pt]{initiation_8cpp__incl}
\end{center}
\end{figure}

\label{_details}
\hypertarget{_details}{}
\subsection{Detailed Description}
\begin{Desc}
\item[Author:]author Xiangyu Hu $<$\href{mailto:Xiangyu.Hu@aer.mw.tum.de}{\tt Xiangyu.Hu@aer.mw.tum.de}$>$ \end{Desc}