### 2010-06-17 src/latex/initiation_8h.tex

\hypertarget{initiation_8h}{
\section{initiation.h File Reference}
\label{initiation_8h}\index{initiation.h@{initiation.h}}
}
Initiates the simulation.

This graph shows which files directly or indirectly include this file:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=420pt]{initiation_8h__dep__incl}
\end{center}
\end{figure}
\subsection*{Classes}
\begin{CompactItemize}
\item 
class \hyperlink{classInitiation}{Initiation}
\begin{CompactList}\small\item\em Initiates the simulation. \item\end{CompactList}\end{CompactItemize}

\label{_details}
\hypertarget{_details}{}
\subsection{Detailed Description}
Initiates the simulation.