### 2010-06-17 src/latex/interaction_8h.tex

\hypertarget{interaction_8h}{
\section{interaction.h File Reference}
\label{interaction_8h}\index{interaction.h@{interaction.h}}
}
Defines interaction between particles.

This graph shows which files directly or indirectly include this file:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=420pt]{interaction_8h__dep__incl}
\end{center}
\end{figure}
\subsection*{Classes}
\begin{CompactItemize}
\item 
class \hyperlink{classInteraction}{Interaction}
\begin{CompactList}\small\item\em Defines interaction between particles. \item\end{CompactList}\end{CompactItemize}

\label{_details}
\hypertarget{_details}{}
\subsection{Detailed Description}
Defines interaction between particles.