### 2010-06-17 src/latex/kernel_8cpp.tex

\hypertarget{kernel_8cpp}{
\section{kernel.cpp File Reference}
\label{kernel_8cpp}\index{kernel.cpp@{kernel.cpp}}
}
the abstract kernel base class

{\tt \#include $<$iostream$>$}\par
{\tt \#include $<$cstdlib$>$}\par
{\tt \#include \char`\"{}glbcls.h\char`\"{}}\par
{\tt \#include \char`\"{}glbfunc.h\char`\"{}}\par

Include dependency graph for kernel.cpp:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=420pt]{kernel_8cpp__incl}
\end{center}
\end{figure}

\label{_details}
\hypertarget{_details}{}
\subsection{Detailed Description}
the abstract kernel base class

\begin{Desc}
\item[Author:]Xiangyu Hu $<$\href{mailto:Xiangyu.Hu@aer.mw.tum.de}{\tt Xiangyu.Hu@aer.mw.tum.de}$>$

changes by: Martin Bernreuther $<$\href{mailto:Martin.Bernreuther@ipvs.uni-stuttgart.de}{\tt Martin.Bernreuther@ipvs.uni-stuttgart.de}$>$

changes by: Andreas Mattes \end{Desc}