### 2010-06-17 src/latex/kernel_8h.tex

\hypertarget{kernel_8h}{
\section{kernel.h File Reference}
\label{kernel_8h}\index{kernel.h@{kernel.h}}
}
\hyperlink{classKernel}{Kernel} abstract base.

This graph shows which files directly or indirectly include this file:\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=420pt]{kernel_8h__dep__incl}
\end{center}
\end{figure}
\subsection*{Classes}
\begin{CompactItemize}
\item 
class \hyperlink{classKernel}{Kernel}
\begin{CompactList}\small\item\em \hyperlink{classKernel}{Kernel} abstract base. \item\end{CompactList}\end{CompactItemize}

\label{_details}
\hypertarget{_details}{}
\subsection{Detailed Description}
\hyperlink{classKernel}{Kernel} abstract base.