### 2010-11-17 ms/titelblattNEW.tex
added files

% \documentclass[12pt]{article}

% \usepackage{ngerman}
% \selectlanguage{ngerman}
% \usepackage[latin1]{inputenc}

% \usepackage{graphicx}

% \begin{document}

\pagestyle{empty}

\begin{center}

\vspace*{-2.8cm}
\begin{minipage}[c]{.30\textwidth}
  \begin{flushleft}
    \includegraphics[width=3cm,clip=]{LOGO/caltech_logo}%
  \end{flushleft}
\end{minipage}
\begin{minipage}[c]{.43\textwidth}
\vspace*{1em}
    {GALCIT (Caltech)\\Prof. Hans G. Hornung \\ \\ Institute of Aerodynamics (TUM) \\ Prof.~Dr.-Ing. N.~A.~Adams}%
\end{minipage}
\begin{minipage}[c]{.25\textwidth}
  \begin{flushright}
    \vspace*{1em}
    \includegraphics[width=3cm,clip=]{LOGO/TUMLogo_oZ_Vollfl_blau_RGB}%
  \end{flushright}
\end{minipage}

\vspace*{4.3cm}
\begin{minipage}[c]{11cm}
{\LARGE\bf 
Simulation of high enthalpy flow in porosities using SPH}
\end{minipage}

\vspace*{1.0cm}

Oliver Oberinger\\
\vspace*{1.0cm}
{\bfseries Master's Thesis}\\
\vspace*{0.5cm}

\vspace*{3cm}

\vspace*{1.2cm}
%\large
\normalsize
\vfill
\begin{tabular}{ll}
Advisors: & ~Prof. Hans G. Hornung (Caltech)\\
\ & \ Dipl.-Ing. Sergey Litvinov (TUM)\\
\ & \ Prof. Dr.-Ing. Nikolaus A. Adams (TUM)\\
 \\
\end{tabular}

\vspace*{1.6cm}
%\large
Institute of Aerodynamics Technische Universit\"{a}t M\"{u}nchen\\
Graduate Aerospace Laboratories California Institute of Technology\\
2010	
\end{center}

\pagebreak
\pagestyle{plain}

%\end{document}